### 2020-04-30 Paper/paper.tex
paper

\documentclass[a4paper,11pt]{article}
\usepackage{ifpdf} 
\ifpdf
\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
             % images in pdf, png or jpg format)
\fi

\usepackage{jcappub}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{amssymb,amsmath,bm}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
%\usepackage[utf8]{inputenc}
\usepackage{xfrac}
\usepackage{aas_macros}
\usepackage{mathrsfs}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{chngcntr}
\newcommand{\nv}{\vec{\theta}}
%\newcommand{\pasj}{Publications of the ASJ}
%\newcommand{\aap}{Astronomy and Astrophysics}
%\newcommand{\apj}{Astrophysical Journal}
%\newcommand{\apjs}{Astrophysical Journal, Supplement}
%\newcommand{\mnras}{Monthly Notices of the RAS}
\newcommand{\todo}[1]{{\bf TODO: #1}}

\newcommand{\as}[1]{{\textcolor{blue}{[AS: #1]}}}
\newcommand{\an}[1]{{\textcolor{magenta}{[AN: #1]}}}
\newcommand{\da}[1]{{\textcolor{red}{[DA: #1]}}}

\newcommand{\vd}{\mathbf{d}}
\newcommand{\vt}{\mathbf{t}}
\newcommand{\vN}{\mathbf{N}}

\newcommand\Tstrut{\rule{0pt}{3ex}}

%\usepackage{lineno}
%\linenumbers

\definecolor{internationalkleinblue}{rgb}{0.0, 0.18, 0.65}
\hypersetup{urlcolor=internationalkleinblue, linkcolor=internationalkleinblue, citecolor=internationalkleinblue}

\usepackage[T1]{fontenc} % if needed
\usepackage{natbib}
\bibliographystyle{JHEP}
\title{Marginalization of photometric redshift uncertainties}

\author[a,1]{Boryana Hadzhiyska,}
\author[b]{David Alonso,}
\author[c]{Andrina Nicola,}
\author[d]{An\v{z}e Slosar,}

\affiliation[a]{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138, USA}
\affiliation[b]{Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom}
\affiliation[c]{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton NJ 08544-0010, USA}
\affiliation[d]{Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
\emailAdd{boryana.hadzhiyska@cfa.harvard.edu}

\abstract{We present a new method for marginalization over uncertainties in redshift distributions of tomographic bins applicable to current and coming photometric galaxy surveys. We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing incompleteness in our knowledge of redshift distributions. In principle, this is marginalization over potentially hundreds of new parameters describing potential deviations as a function of redshift and tomographic bin. However, by Taylor expanding theory predictions around a fiducial model, we can perform the marginalization analytically, resulting in a modified data covariance matrix. We demonstrate  this method by applying to the galaxy clustering measurements from the Year One of Hyper SupremeCam data demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods. We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations and show that the only deviations that matter are smooth deviations. 
}

\begin{document}
\maketitle
\flushbottom

\section{Introduction}\label{sec:intro}

Yadi, yadi, yadi. Upcoming photometric galaxy suryves, yadi yadi,
importance of 3x2point, yadi, yadi. Define $N(z)$.

\section{Theory}

In typical cosmological analysis problem, we are performing a Bayesian inference on the posterior $p({\rm theory} | {\rm data} ) \propto p({\rm data}|{\rm} theory) p({\rm theory})$ with the data likelihood given by 
\begin{equation}
p({\rm data}|{\rm theory}) =  \mathcal{L} \propto  \exp\left [-\frac{1}{2} \chi^2 \right], \label{eq:like}
\end{equation}
where
\begin{equation}
  \chi^2  = (\vd-\vt)^T C^{-1} (\vd-\vt),
\end{equation}
where $\vd$ is the data vectory, typically containing all the summary statistic of interest like correlation functions ot power spectra and $\vt$ the theory vector through which parameter dependence enters. In this set-up the covariance matrix $C$ is fixed (and hence the Gaussian normalization term does not enter equation \ref{eq:like}), which is usually a very good approximation in typical problems that we wish to consider \cite{1811.11584}, but note that the method is easily generalizable to the cases where this is not true.

We can make significant progress without specifying explicitly how theory is calculate given the parameters. It suffices to say that given a set of cosmological parameters, which specify the evolution of the universe and thus matter power spectra as a function of scale and redshift, a set of survey properties, such as number densities and redshift distribution of galaxies and weak lensing sources, we can integrate these into theory predictions $\vt=\vt(\theta)$.

The traditional approach is to add to add nuissance parameters that describe our lack of knowledge of galaxy redshift distribution. A typical approach is to use shift and width parameters  \cite{1706.09359,1912.08209} \as{did DES Y1 have shift/width?},
where redshift distributions $N_i(z)$ are shifted according to
    \begin{equation}
      N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right),
      \label{eq:photo-z-model}
    \end{equation} 
where the index $i$ runs over the number of tomographic samples. Here $N_{i}(z)$ is the true redshift distribution with  $\bar{N}_{i}(z)$ out best estimate of it. The parameters $\Delta z_{i}$ and $z_{w,i}$ account for shifts in the means of the distributions, while $z_{c,i} = \left<z\right>_{\bar{N}_i}$ is the mean redshift to act as a sensible pivot for width changes. The shift and width parameters for each tomographic bin are then inferred and marginalized over together with all other parameters.

This approach explicitly marginalizes over the two effects that we think are the two most important systematics stemming from redshift distribution uncertainty. However, it suffers from two problems. First, it is fundamentally ad-hoc. This particular parametric form does not stem from consideration of phometric redshift estimators. There is also no reason to believe why it would be complete in the sense that it captures all relevant effects. The second problem is that it is computationally expensive. For example, in the analysis 
    of \cite{1912.08209}, redshift uncertainty parameters accounted for 8 (2 per 4 tomographic bins) of 14 total parameters in the inference process, i.e. over the half parameters inferred were dealing with redshift.

\subsection{A new approach}

In this paper we advocate a different approach. First, let's consider a binned description of $N_i(z)$ in terms of $N_{ij}$, where $i$ index describes the tomographic bin and $j$ is redshift index, so that $N_{ij} = N_i(z_j)$ at some discrete grid of $z_j$. This entire vector can be packed in a vector $\vN$.

We will describe our uncertainty in $\vN$ as a Gaussian
\begin{equation}
  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
\end{equation}
where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations. We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal). Other possible parameterizations are sub-summed in this approach. For example, a general shift-width is allowed with the probability $p(\vN(\Delta z_i, z_{w,i}))$.

We can now write a likelihood for the data in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
\begin{equation}
\mathcal{L} \propto  \int \exp\left [-\frac{1}{2} (\vd-\vt)^T C^{-1} (\vd-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
  \end{equation}
  Note that in the equation above, $\vt=\vt(\theta,\vN)$, where $\theta$ are the remaining parameters. It is this dependence that prevents us from performing what looks like a trivial integral.

  In principle, we could solve this problem by making every single element of vector $\vN$ a part of MCMC chain and simply run the MCMC with $\theta$ and additional hundreds of $\vN$ parameters. Instead, we pull an Eastern-European trick and Taylor expand theory around $\hat{\vN}$:
  \begin{equation}
    \vt(\theta,\vN) = \vt(\theta,\bar{\vN}) + T \left(\vN - \bar{\vN} \right), \label{eq:taylor}
  \end{equation}
  where
  \begin{equation}
    T_{mn} = \frac{\partial t_m}{\partial N_n} \bigg|_{\theta,\hat{\vN}}
  \end{equation}
  The matrix $T$ is the non-square matrix giving derivatives of theory with respect to each element of $\vN$. In other words it contains a full information about how all the correlation functions or power spectra respond to changing $N(z)$ distribution for one tomographic bin at one redshift.

  Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like} results in an integral that is now quadratic in $\bar{\vN}$ and can be performed analytically. After some bothersome algebra, we find:
  \begin{equation}
    \mathcal{L} \propto \frac{1}{|T C^{-1} T^T +P |^{1/2}} \exp\left [-\frac{1}{2} (\vd-\vt)^T C_M^{-1} (\vd-\vt) \right], \label{eq:main1}
  \end{equation}
  where
  \begin{equation}
    C_M = C + T P^{-1}T^T \label{eq:main2}
  \end{equation}
  This is a fascinatingly simple equations and the main result of this paper. What is happening here is that for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$. These are directions in  space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.  Equation \ref{eq:main2}  increases the variance for those deviations commensurate with how far P allows them to go.

  In principle, at every point in likelihood, one should evaluate Eq. \ref{eq:main1}, which means recalculating $T$. Since this is a very expensive computation, we will further simplify the problem, by assuming $T$ is only weakly deviations from some fiducial cosmology. We will therefore pre-compute $T$ once for the fiducial cosmology, which then means that normalization term in equation \ref{eq:main1} is an irrelevant constant and at fixed data covariance matrix $C$, the Eq. \ref{eq:main2} also needs to be evaluated just once.

  The procedure is thus as follows:
  \begin{itemize}
  \item Run inference problem once with fixed redshift distribution to determine a sensible fiducial model.

  \item Calculate $T$ and $C_M$ at this fiducial model

  \item Re-run inference with the modified covariance matrix $C_M$, which leads to broadened contours due to marginalization of redshift distributions

  \end{itemize}
If needed, one could repeat the inference at a refined fiducial model although in pracice we found this to be unnecessary.

Note that there are two distinct approximation here. The first is that theory can validly be Taylor expanded in $\vN-\bar{\vN}$ and the second is that $T$ is model independent for models of interest. We will examine validity of botho of those approximatons in Section \ref{sec:hsc}

\subsection{Determination of prior matrix P}

\subsubsection{Uncertainty due to estimator biases}

\subsubsection{Cosmic Variance in calibrator sample}

\subsubsection{Smoothness prior}

\section{Appplication to HSC clustering data}
\label{sec:hsc}

\section{Conclusions}

\bibliography{bibliography}

\end{document}

### 2020-04-30 Paper/paper.tex
synch test.
P  vs P-1 fix
woodbury

@@
 
-  Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like} results in an integral that is now quadratic in $\bar{\vN}$ and can be performed analytically. After some bothersome algebra, we find:
+  Substituting Eq.
+\ref{eq:taylor} into Eq.
+\ref{eq:like} results in an integral that is now quadratic in $\bar{\vN}$ and can be performed analytically.
+After collecting terms that are quadratic and linear in $\vN$, performing the Gaussian integral and applying the Woodbury matrix identity, we find:
   \begin{equation}
-    \mathcal{L} \propto \frac{1}{|T C^{-1} T^T +P |^{1/2}} \exp\left [-\frac{1}{2} (\vd-\vt)^T C_M^{-1} (\vd-\vt) \right], \label{eq:main1}
+    \mathcal{L} \propto \frac{1}{|T C^{-1} T^T +P^{-1} |^{1/2}} \exp\left [-\frac{1}{2} (\vd-\vt)^T C_M^{-1} (\vd-\vt) \right], \label{eq:main1}
   \end{equation}
@@
   \begin{equation}
-    C_M = C + T P^{-1}T^T \label{eq:main2}
+    C_M = C + T PT^T \label{eq:main2}
   \end{equation}
-  This is a fascinatingly simple equations and the main result of this paper.
-What is happening here is that for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$.
+  This is a fascinatingly simple equation and the main result of this paper.
+This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$.
 These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.
-Equation \ref{eq:main2}  increases the variance for those deviations commensurate with how far P allows them to go.
+Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far P allows them to go.
 
@@
 The first is that theory can validly be Taylor expanded in $\vN-\bar{\vN}$ and the second is that $T$ is model independent for models of interest.
-We will examine validity of botho of those approximatons in Section \ref{sec:hsc}
+We will examine validity of both of those approximations in Section \ref{sec:hsc}.
 

### 2020-05-02 create_sacc.py
some updates for new covmat

@@ -11,6 +11,8 @@ from calculate_smooth_s_and_prior import get_smooth_s_and_prior
 
 # Note to self and reader -> this should eventually be run with the real data, i.e.
 # read = 'COADDED'; write = 'MARG'
+# or for realistic NG covariance
+# read = 'NEWCOV_COADDED'; write = 'NEWCOV_MARG'
 # Important REMOVE l-cuts for the real run
 # TODO we are assuming cosmic variance of the tomographic bins is uncorrelated
 # For testing I am using read = 'COADD' and write = 'test_2' 
@@ -26,7 +28,8 @@ dir_write = "data/"+write
 
 # l-cuts
 # Remove this for the actual MCMC run as the sampler makes its own cut TODO
-lmax = [2000,2000,2600,3200]
+#lmax = [2000,2000,2600,3200]
+lmax = [2170.58958919, 2515.39193451, 3185.36076391, 4017.39370804]
 
 # choice for noise and smoothing
 A_smooth = 0.25
@@ -55,7 +58,8 @@ hod_sigmas = {'lmmin':0.22, 'lmminp':2.0,
 
 # Here are some choice that we make for the tests we are running
 # TEST #1
-if write == "MARG_1sigma":
+if "1sigma" in write:
+    # Delete the CV covmat and the Tmat
     print("Getting marginalized covariance for the 1-sigma test")
     coin = 1
     for key in hod_sigmas.keys():
@@ -64,20 +68,18 @@ if write == "MARG_1sigma":
         print(key,hod_params[key])
 
 # TEST #2
-if write == "MARG_nosmooth":
+if "nosmooth" in write:
+    # Delete the CV covmat
     print("Getting marginalized covariance for the no-smoothing test")
     A_smooth = 0.
 
 # TEST #3
-if write == "MARG_largenoise":
+if "largenoise" in write:
+    # Delete the CV covmat
     print("Getting marginalized covariance for the large-noise test")
     noi_fac = 42.
 
-# TEST #4
-if write == "MARG_nosmooth_largenoise":
-    print("Getting marginalized covariance for the no-smoothing+large-noise test")
-    A_smooth = 0.
-    noi_fac = 42.
+# TEST #4 there is this test too "MARG_nosmooth_largenoise":
     
 # Implementation of the CV+noise precision matrix from Eq. 7 in the HSC Nz marg overleaf
 # NB: the T matrix here is the transpose of what is written in the equations on overleaf

### 2020-05-02 coadd_saccs.py
adding the covariance plotting code

@@ -1,10 +1,16 @@
 import sacc
 import os
 import sys
+import numpy as np
 
 # Name of directory where sacc files will be saved
-write = sys.argv[1] # I am currently choosing COADDED
+write = sys.argv[1] # I am currently choosing COADDED, NEWCOV_COADDED
 dir_write = "data/"+write
+
+# Want to include the new covariance matrix?
+want_newcovmat = bool(sys.argv[2])
+if want_newcovmat:
+    newcov = np.load("/users/boryanah/repos/WeePeeZee/data/NEW_COVMAT/covmat.npy")
 
 # Loading the saccs
 sacc_names = ['/users/boryanah/HSC_data/HSC/GAMA15H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/GAMA09H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/WIDE12H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/VVDS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/XMMLSS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc']
@@ -23,6 +29,16 @@ for i, s in enumerate(saccs):
 # Coadding the files normalizing by area
 sacc_coadded = sacc.coadd(saccs, mode='area')
 sacc_noise_coadded = sacc.coadd(saccs_noise, mode='area')
+
+# Change the covmat to the new one if requested
+if want_newcovmat:
+    lmin = [0, 0, 0, 0]
+    lmax = [2170.58958919, 2515.39193451, 3185.36076391, 4017.39370804]
+    sacc_coadded.cullLminLmax(lmin, lmax)
+    sacc_noise_coadded.cullLminLmax(lmin, lmax)
+    
+    sacc_coadded.precision = sacc.Precision(newcov, 'dense', is_covariance=True)
[…]

### 2020-05-02 plot_cov.py

@@ -1,0 +1,87 @@
+import sacc
+import numpy as np
+import matplotlib.pyplot as plt
+
+def cov2corr(cov):                                                                                              
+    d = np.diag(cov)
+    stddev = np.sqrt(d)
+    corr = cov.copy()*0.
+    for i in range(cov.shape[0]):
+        for j in range(cov.shape[0]):                                                                           
+            corr[i,j] = cov[i,j]/stddev[i]/stddev[j]
+    return corr
+
+def cov2corr(covmat):
+    corrmat = covmat/np.sqrt(np.diag(covmat)[None,:]*np.diag(covmat)[:,None])-np.diag(np.ones(covmat.shape[0]))
+    return corrmat
+    
+def plot2d(data,which,lim=None,levels=None,clip=0,clbar=True,cm=None,label=None,labsize=14,extent=None,ticksize=12,*args):
+    # origin at upper left
+    Nx = data.shape[0]
+    Ny = data.shape[1]
+    arr = data[clip:Nx-clip,clip:Ny-clip]
+    if type(lim) is list:
+        limmin,limmax = lim
+    elif lim==None:
+        limmin=None
+        limmax = None
+    else:
+        limmin=-lim
+        limmax = lim
+    plt.imshow(arr,interpolation="none",vmin=limmin,vmax=limmax,cmap=cm,extent=extent,*args)
+    plt.colorbar()                                                            
+    plt.savefig(which+".png")
+    plt.close()
+
+def plot_evals(data,which):
+    Nx = data.shape[0]
+    evals,evecs = (np.linalg.eig(data))
+    evals_sorted = np.sort(evals)
+    assert evals_sorted[0] > 0., "The matrix is not positive definite"
+    plt.plot(np.arange(Nx),evals_sorted,label=which)
+    plt.yscale('log')
+    
+# Names
+#dir_read = "data/COADDED"
+dir_read = "NEWCOV_COADDED"
+#dir_write = "data/MARG"
+dir_write = "NEWCOV_MARG"
+
+# Load the original and marginalized data 
+s_d = sacc.SACC.loadFromHDF("data/"+dir_read+"/power_spectra_wdpj.sacc")
+s_m = sacc.SACC.loadFromHDF("data/"+dir_write+"/power_spectra_wdpj.sacc")
+
[…]

### 2020-05-14 create_sacc.py
some problems

@@ -5,9 +5,11 @@ import numpy as np
 import pyccl as ccl
 import copy
 import shutil
+import matplotlib.pyplot as plt
 from modules.halo_mod_corr import HaloModCorrection
 from modules.theory_cls import get_theory
 from calculate_smooth_s_and_prior import get_smooth_s_and_prior
+import scipy.linalg as la
 
 # Note to self and reader -> this should eventually be run with the real data, i.e.
 # read = 'COADDED'; write = 'MARG'
@@ -176,7 +178,6 @@ N_tracers = len(s_d.tracers)
 Nz_total = N_tracers*Nz_per_tracer
 
 # Applying l-cuts; remove for MCMC run
-#s_d.cullLminLmax([0,0,0,0],lmax)
 
 # Get the covariance and precision matrices
 N_data = len(s_d.mean.vector)
@@ -203,6 +204,9 @@ cov_n = sacc.Precision(cov_CVnoismo)
 # Create new sacc file (TODO: ask about "dense")
 s_n = sacc.SACC(s_d.tracers,s_d.binning,s_d.mean.vector,cov_n,meta=s_d.meta)
 
+
+
+
 # Save the new covariance and the power spectrum of the data which shall be run via MCMC
 if not os.path.exists(dir_write):
     os.makedirs(dir_write)
@@ -221,23 +225,63 @@ prec_CVnoismo = s_n.precision.getPrecisionMatrix()
 dof = N_data
 
 # Function to calculate chi2
-def print_chi2(di,precision):
+def get_chi2(di,precision):
     # Report the chi2
     return np.dot(di,np.dot(precision,di))
 
-# Difference of N(z) between smooth and jagged
-dNz = NzVec(s_m)-NzVec(s_d)
+## there are already culled in ell
 
+prec_o = s_d.precision.getPrecisionMatrix()
+prec_n = s_n.precision.getPrecisionMatrix()
 # Computing Cls directly from the data
+
 cl_theory = get_theory(hod_params,cosmo_params, s_d,halo_mod_corrector=HMCorrection)
 
+# xar = np.arange(len(cl_theory))
+# errv = np.sqrt(s_d.precision.getCovarianceMatrix().diagonal())
+# #plt.plot(xar,cl_theory,'b-')
[…]

### 2020-05-14 sanity_chi2.py
sanity

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+import os
+import sacc
+import sys
+import numpy as np
+import pyccl as ccl
+import copy
+import shutil
+import matplotlib.pyplot as plt
+from modules.halo_mod_corr import HaloModCorrection
+from modules.theory_cls import get_theory
+from calculate_smooth_s_and_prior import get_smooth_s_and_prior
+import scipy.linalg as la
+
+
+s_d = sacc.SACC.loadFromHDF("data/NEWCOV_COADDED/power_spectra_wdpj.sacc")
+cosmo_params = {'n_s': 0.9649,
+                'sigma8':0.8111,
+                'h':0.6736,
+                'Omega_c':0.264,
+                'Omega_b':0.0493}
+
+# HOD params: assuming posterior mean from paper
+hod_params = {'zfid':0.65,
+              'lmmin':11.88, 'lmminp':-0.5,##
+              'sigm_0':0.4, 'sigm_1':0,
+              'm0':5.7, 'm0p':2.5,##
+              'm1':13.08, 'm1p':0.9,##
+              'alpha_0':1, 'alpha_1':0,
+              'fc_0':1., 'fc_1':0.}
+
+cosmo = ccl.Cosmology(**cosmo_params)
+HMCorrection = HaloModCorrection(cosmo, k_range=[1e-4, 1e2], nlk=256, z_range=[0., 3.], nz=50)
+cl_theory = get_theory(hod_params,cosmo_params, s_d,halo_mod_corrector=HMCorrection)
+
+di = s_d.mean.vector - cl_theory
+prec = s_d.precision.getPrecisionMatrix()
+chi2 = np.dot(di,np.dot(prec,di))
+print ("chi2 = ",chi2, "dof = ", len(di))

### 2020-05-15 Paper/paper.tex

@@
 \subsubsection{Cosmic Variance in calibrator sample}
+The covariance matrix between two tomographic bins is given by:
+\begin{equation}
+Cov_{ij}=\frac{N_i N_j}{2\pi^2} \int_0^\infty dk_\parallel \cos(k_\parallel(\chi_i-\chi_j))\int_0^\infty dk_\perp k_\perp W_i(k_\parallel,k_\perp)W_j(k_\parallel,k_\perp)P_{\bf gg}({\bf k}),
+\end{equation}
+where we can model the power spectra at each bin centre and for all
+transverse and parallel modes using the Kaiser formula (CITE), which corrects for redshift space distortions through
+the expression:
+\begin{equation}
+P_{gg}({\bf k},z)=\left(b_g(z)+f(z)\frac{k_\parallel^2}{k^2}\right)^2\,P_{mm}(k,z),
+\end{equation}
+where $f(z)$ is
+the logarithmic growth
+rate
+and the
+evolution of the linear galaxy bias is given by $b_g(z) = 0.95/D(z)$
+with $D(z)$ the growth factor.
+
+In this work, we use the galaxy sample in the COSMOS survey as calibrator, which
+covers a total area of
+$A_{\rm sky} = 1.7 \ {\rm deg}^2$ for galaxies with redshifts $z < 2$.
+We model the window function for each bin of the dataset through:
+\begin{equation}
+W(k_\parallel,k_\perp)=j_0(k_\parallel\Delta\chi/2)\,\frac{2 J_1(k_\perp R_{\rm mean})}{k_\perp R_{\rm mean}},
+\end{equation}
+where $R_{mean}$ is the
+physical size of the survey measured at the mean
+distance to the
+tomographic bin, 
+$R_{\rm mean} = 
+\sqrt{\frac{A_{\rm sky}}{\pi}} \chi_{\rm mean}$.
+We assume that the
+cosmic covariance
+between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by
+treating each bin independently. We show the resulting correlation matrix of the COSMOS cosmic variance in Fig. \ref{fig:CV}.
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.\textwidth]{Paper/corr_CV_0.png}
+\caption{Cosmic variance (CV) correlation matrix with subtracted diagonal for the first tomographic bin of the HSC data.} 
+\label{fig:CV}
+\end{figure}
 
@@
 \label{sec:hsc}
+\an{Some citations are currently missing but I will put them in later.}
+We apply the method outlined above to the data obtained in Ref.~\cite{1912.08209}.
+This work measured the angular galaxy clustering power spectrum from HSC DR1 data.
+In the following, we give a very brief summary of the methodology employed in this work and refer the reader to Ref.~\cite{1912.08209} for more details.
+\subsection{Background theory}
+The angular clustering signal for galaxies in redshift bins $i, j$ with photo-$z$ distributions $p^i(z), p^j(z)$ can be modeled using the Limber approximation as \cite{1953ApJ...117..134L, 1992ApJ...388..272K, Kaiser:1998}
+    \begin{equation}\label{eq:cell_gg_limber}
+      C^{ij}_\ell = \int \mathrm{d}z\,\frac{H(z)}{\chi^2(z)} p^i(z)p^j(z)\,P_{gg}\left(z,k=\frac{\ell+1/2}{\chi(z)}\right),
+    \end{equation}
+where $P_{gg}(z,k)$ denotes the underlying 3D galaxy power spectrum, $\chi(z)$ is the comoving distance and $H(z)$ denotes the Hubble parameter at redshift $z$.
+As can be seen from Eq.~\ref{eq:cell_gg_limber}, the spherical harmonic power spectrum $C_{\ell}$ depends linearly on the photometric redshift distributions $p^i(z)$, which facilitates the computation of the $T$ matrix defined above.
+
+We follow Ref.~\cite{1912.08209} and compute theoretical predictions for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}.
+\subsection{The HSC dataset}
+The Hyper-Suprime Cam survey is an on-going photometric galaxy survey survey focused mainly on weak gravitational lensing.
+The analysis in Ref.~\cite{1912.08209} is based on the publicly-available HSC DR1 data, whose so-called wide fields cover approximately 108 square degrees on the sky, subdivided into seven distinct patches.
+Ref.~\cite{1912.08209} used these data to compute spherical harmonic galaxy clustering power spectra for four tomographic redshift bins between $z=0.15$ and $z=1.5$, taking both auto- and cross-correlations into account.
+These power spectra have been corrected for observational and extragalactic systematics by deprojection at the map-level \cite{2019MNRAS.484.4127A}.
+Finally, following Ref.~\cite{2019PASJ...71...43H}, photometric redshift distributions have been estimated by cross-matching HSC galaxies to galaxies in the COSMOS 30-band photometric catalog presented in Ref.~\cite{2016ApJS..224...24L}.
+\subsection{Validating the linear expansion}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.2\textwidth]{Paper/Tmat.png}
+\caption{Visualization of the derivative matrix $T_{mn} = {\partial t_m}/{\partial N_n}$. \todo{Make look nicer: split into tomo bins}} 
+\label{fig:Tmat}
+\end{figure}
+
+\begin{itemize}
+  \item Plot showing $T$
+  \item Plot showing exact $C_\ell$ vs linear prediction. \todo{Make your plots into subplots; verify the Tmatrix}
+\end{itemize}{}
+\subsection{Impact of N(z) uncertainties on final parameters}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{Paper/corr_data.png}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{Paper/corr_diff.png}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{Correlation matrix for the fiducial HSC dataset (\textit{left}) and difference with its marginalized version (\textit{right}).}
+\label{fig:fid_marg_cov}
+\end{figure}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.\textwidth]{Paper/triangle_fid_marg.png}
+\caption{Triangle plot with constraints of the 6 HOD parameters obtained using the covariance matrix from the full HSC dataset (\textit{red}) and from its marginalized version (\textit{blue}).}
+\label{fig:triangle_fid_marg}
+\end{figure}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.\textwidth]{Paper/triangle_marg_tests.png}
+\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green} is shown the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from \todo{CITE Andrina's paper}; in \textit{red} is shown the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally in \textit{blue} is shown the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 0.25$.}
+\label{fig:triangle_marg_tests}
+\end{figure}
+
+Plots showing the effects of different things:
+\begin{itemize}
+\item Compare different choices of prior (smoothness, no smoothing, different degrees of noise etc.)
+\item Triangle plot with T matrix computed from a different fiducial.
+\item Compare CV-only covariance vs. no covariance.
+\item Compare old marginalization vs. new one (maybe same plot).
+\item Compare old marginalization with CV priors (importance sample).
+\end{itemize}
+
+Other results:
+\begin{itemize}
+  \item smoothness does nothing
+  \item quote $\chi^2$ values \todo{Make a table}
+\end{itemize}
+
+\begin{center}
+ \begin{tabular}{c | c c c} 
+ \hline\hline
+ Parameter & HSC covariance, full & HSC covariance, HOD & Marginalized covariance, HOD \\ [0.5ex] 
+ \hline
+ $\chi^2/\nu$ & 87.49/80 & 94.97/88 & 91.06/88 \\
+
+ $\mu_{{\rm min},p}$ & -0.52 & -1.47 & -1.42 \\
+
+ $\mu_{\rm min}$ & 11.88 & 11.73 & 11.73 \\
+
+ $\mu_{0,p}$ & 1.68 & 4.86 & 5.03 \\
+
+ $\mu_{0}$ & 0.02 & 11.53 & 11.41 \\
+
+ $\mu_{1,p}$ & 0.87 & -0.58 & -0.51 \\
+
+ $\mu_{1}$ & 13.073 & 12.866 & 12.873 \\ [1ex] 
+ \hline
+ \hline
+\end{tabular}
+\end{center}
+%[-1.42408101 11.73454764  5.0302023  11.41301207 -0.50980494 12.87323354]
+%[-1.47048791 11.73097558  4.85774675 11.52813388 -0.58053445 12.86613972]
+%[ 2.18173486e-02 -1.39414027e-02 -4.61928378e-03  1.27879857e-02
+% -5.18739083e-01  1.18844059e+01  1.67767553e+00  2.01822301e-02
+%  8.65219859e-01  1.30726966e+01 -1.64307267e-01 -3.19342441e-02
+%  3.97310971e-02  3.88321407e-02]
 

### 2020-05-16 Paper/paper.tex

@@
 \usepackage{graphicx}
+\graphicspath{{Paper/}{./}}
 \usepackage{dcolumn}
@@
 
-Yadi, yadi, yadi. Upcoming photometric galaxy suryves, yadi yadi,
-importance of 3x2point, yadi, yadi. Define $N(z)$.
+Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky.
+These catalogs can be used to constrain cosmological parameters through measurements of clustering and weak lensing as demonstrated by numerous surveys that have completed observations including  Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201}.
+
+it has been long recognized that the accuracy of photometric surveys relies on the accuracy of photometric redshift distributions and that these are likely to continue to be a dominant source of systematic uncertainty.
+The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number density of galaxies as a function of redshift for each tomographic sample.
+It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}.
+Improving photometric redshift techniques and calibrations is an area of active research \cite{2004.09542}\as{add citations}, but for the foreseable future we will need to rely on marginalization to control residual systematic errors.
+
+The traditional approach is to add nuissiance parameters that shift the best guess $N(z)$ or additionally change its width or a relative contribution from a secondary peak in $N(z)$.
+This usually works well, but in additional to adding potentially large number of nuissance parameters suffers from model completeness problem: how do we know that any given parameterization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?
+
+In this paper we propose a new technique that performs a marginalization around all possible functional deviations from the best fit $N(z)$.
+The space of possible deviations in $N(z)$ is described as a Gaussian function and under some controlled approximations, the integrals can be performed analytically.
+This results in a simple change to the data covariance matrix with a simple explanation: directions in the space of predictions that are degenerate with changes in $N(z)$ are given large variance, which means that information in these direction is not used for inferring the parameters of interest.
+This then marginalizes over $N(z)$ uncertainty in a robust manner that does not add new niussance parameters.
+
+We applyt this method to the HSC data as an example and  discuss some of the advantages, limitations and possible extensions of this model.
 
@@
 
-\subsection{Determination of prior matrix P}
+\subsection{Determination of prior matrix $P$}
+
+Determination of the prior matrix $P$ is where the the actual and often partially subjective prior on our knowledge of $N(z)$ enters.
+Note that the standard parameterizations can always be recast in this language at least in the perturbative limits.
+For example, a vector of derivatives $\lambda_i = dN_i/dz$ when added to $\vN$ will produce a shift in the the shape of $N_i(z)$.
+Therefore, adding a contribution $\sigma_{zi}^2 \lambda_i \lambda_i^T$ will allow shifts with variance $\sigma_{zi}$.
+Of course, when shifts become so large that the perturbative expansion breaks, the approach will work only approximately.
+But in this paper we are not interesting in shifts; instead we will try to build $P$ is a more systematic manner.
 
 \subsubsection{Uncertainty due to estimator biases}
+We will build $P$ from three pieces.
+The first piece is simply the overall uncertainty stemming from the fact that the photometric redshift estimators have errors.
+The simplest thing is to make this covariance matrix   \ldots
 
@@
 \centering  
-\includegraphics[width=1.\textwidth]{Paper/corr_CV_0.png}
+\includegraphics[width=1.\textwidth]{./corr_CV_0.png}
 \caption{Cosmic variance (CV) correlation matrix with subtracted diagonal for the first tomographic bin of the HSC data.} 
@@
 \centering  
-\includegraphics[width=1.2\textwidth]{Paper/Tmat.png}
+\includegraphics[width=1.2\textwidth]{./Tmat.png}
 \caption{Visualization of the derivative matrix $T_{mn} = {\partial t_m}/{\partial N_n}$. \todo{Make look nicer: split into tomo bins}} 
@@
   \centering
-  \includegraphics[width=1.\linewidth]{Paper/corr_data.png}
+  \includegraphics[width=1.\linewidth]{./corr_data.png}
   \label{fig:sub1}
@@
   \centering
-  \includegraphics[width=1.\linewidth]{Paper/corr_diff.png}
+  \includegraphics[width=1.\linewidth]{./corr_diff.png}
   \label{fig:sub2}
@@
 \centering  
-\includegraphics[width=1.\textwidth]{Paper/triangle_fid_marg.png}
+\includegraphics[width=1.\textwidth]{./triangle_fid_marg.png}
 \caption{Triangle plot with constraints of the 6 HOD parameters obtained using the covariance matrix from the full HSC dataset (\textit{red}) and from its marginalized version (\textit{blue}).}
@@
 \centering  
-\includegraphics[width=1.\textwidth]{Paper/triangle_marg_tests.png}
+\includegraphics[width=1.\textwidth]{./triangle_marg_tests.png}
 \caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green} is shown the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from \todo{CITE Andrina's paper}; in \textit{red} is shown the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally in \textit{blue} is shown the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 0.25$.}

### 2020-05-16 create_sacc.py
slightly more elegant Tmatrix calculation not changing anything

@@ -30,7 +30,6 @@ dir_write = "data/"+write
 
 # l-cuts
 # Remove this for the actual MCMC run as the sampler makes its own cut TODO
-#lmax = [2000,2000,2600,3200]
 lmax = [2170.58958919, 2515.39193451, 3185.36076391, 4017.39370804]
 
 # choice for noise and smoothing
@@ -82,18 +81,7 @@ if "largenoise" in write:
     noi_fac = 42.
 
 # TEST #4 there is this test too "MARG_nosmooth_largenoise":
-    
-# Implementation of the CV+noise precision matrix from Eq. 7 in the HSC Nz marg overleaf
-# NB: the T matrix here is the transpose of what is written in the equations on overleaf
-# !!!Important!!! Function currently not called since we use the simplified obtain_improved_cov
-def obtain_improved_prec(prec,Tmat,prior):
-    # The expressions inside the bracket, before the bracket and after the bracket
-    bracket = np.linalg.inv((np.dot(np.dot(Tmat,prec),Tmat.T)+prior))
-    prebracket = np.dot(prec,Tmat.T)
-    postbracket = np.dot(Tmat,prec)
-    # Assembling everything into one
-    prec_m = prec - np.dot(np.dot(prebracket,bracket),postbracket)
-    return prec_m
+
 
 # This is the improved covariance with David's simplification (Woodbury matrix identity)
 # NB: the T matrix here is the transpose of what is written in the equations on overleaf
@@ -103,58 +91,23 @@ def obtain_improved_cov(cov,Tmat,prior):
     return cov_m
 
 # The T matrix is simply a matrix containing the first der of Cl with resp to Nz
-def obtain_Tmat(s_data,Nz,Nztr,hod_pars,cosmo_pars,HMCorr,delta=1.e-3):
+def obtain_Tmat(s_data,Nz,Nztr,Nd,hod_pars,cosmo_pars,HMCorr,delta=1.e-3):
     # Create T matrix
-    Tmat = []
+    Tmat = np.zeros((Nz,Nd))
     for i in range(Nz):
-        print ("%i / %i "%(i,Nz))
+        print ("tomo bin, z-sample = %i / %i "%(i//Nztr,i%Nztr))
         s_tmp = copy.deepcopy(s_data)
         s_tmp.tracers[i//Nztr].Nz[i % Nztr] += delta
[…]

### 2020-05-16 plot_cov.py
plotting script

@@ -2,20 +2,14 @@ import sacc
 import numpy as np
 import matplotlib.pyplot as plt
 
-def cov2corr(cov):                                                                                              
-    d = np.diag(cov)
-    stddev = np.sqrt(d)
-    corr = cov.copy()*0.
-    for i in range(cov.shape[0]):
-        for j in range(cov.shape[0]):                                                                           
-            corr[i,j] = cov[i,j]/stddev[i]/stddev[j]
-    return corr
-
-def cov2corr(covmat):
-    corrmat = covmat/np.sqrt(np.diag(covmat)[None,:]*np.diag(covmat)[:,None])-np.diag(np.ones(covmat.shape[0]))
+def cov2corr(covmat,nodiag=True):
+    if nodiag:
+        corrmat = covmat/np.sqrt(np.diag(covmat)[None,:]*np.diag(covmat)[:,None])-np.diag(np.ones(covmat.shape[0]))
+    else:
+        corrmat = covmat/np.sqrt(np.diag(covmat)[None,:]*np.diag(covmat)[:,None])
     return corrmat
     
-def plot2d(data,which,lim=None,levels=None,clip=0,clbar=True,cm=None,label=None,labsize=14,extent=None,ticksize=12,*args):
+def plot2d(data,which,logscale=False,lim=None,levels=None,clip=0,clbar=True,cm=None,label=None,labsize=14,extent=None,ticksize=12,*args):
     # origin at upper left
     Nx = data.shape[0]
     Ny = data.shape[1]
@@ -28,8 +22,9 @@ def plot2d(data,which,lim=None,levels=None,clip=0,clbar=True,cm=None,label=None,
     else:
         limmin=-lim
         limmax = lim
+    if logscale: arr[arr==0.] = 1.e-14; arr = np.log10(np.abs(arr))
     plt.imshow(arr,interpolation="none",vmin=limmin,vmax=limmax,cmap=cm,extent=extent,*args)
-    plt.colorbar()                                                            
+    plt.colorbar()
     plt.savefig(which+".png")
     plt.close()
 
[…]

### 2020-05-16 test_like.py
restructuring test_like.py for readability

@@ -13,21 +13,27 @@ from desclss.halo_mod_corr import HaloModCorrection
 import yaml
 
 import matplotlib.pyplot as plt
+import sys
+from scipy.interpolate import interp1d
 
 #######################################
 ###              MY CODE            ###
 #######################################
 # Names
-#dir_read = "data/COADDED" #orig
-#dir_read = "data/MARG"
+read = sys.argv[1]# NEWCOV_COADDED
+dir_read = "data/"+read
+
 #dir_read = "data/NEWCOV_MARG"
-dir_read = "data/NEWCOV_COADDED"
-#dir_read = "data/MARG_nosmooth"
-#dir_read = "data/MARG_largenoise"
-#dir_read = "data/MARG_nosmooth_largenoise"
-only_6HOD = 1
+#dir_read = "data/NEWCOV_MARG_nosmooth"
+#dir_read = "data/NEWCOV_MARG_largenoise"
+#dir_read = "data/NEWCOV_MARG_nosmooth_largenoise"
+only_6HOD = bool(int(sys.argv[2]))
+test_Tmat = 0
+test_zparams = 0
+verbose = 0
 
 np.random.seed(300)
+
 
 # Function to calculate chi2
 def print_chi2(di,precision):
@@ -102,44 +108,72 @@ dNz = NzVec(s_m)-NzVec(s_d)
 # Slightly improved computation using taylor expansion
 cl_theory_taylor = cl_theory + np.dot(Tmat.T,dNz)
 
-
 # Test how well the Cls are captured by the Tmatrix
-cov_m = np.linalg.inv(prior_m)
-s_new = copy.deepcopy(s_m)
-for i in range(4):
-    nz_new = s_new.tracers[i].Nz
-    nz_size = len(nz_new)
-    if i == 2:
-        gaussian_draw = np.random.multivariate_normal(np.zeros(nz_size), cov_m[i*nz_size:(i+1)*nz_size,i*nz_size:(i+1)*nz_size])
-        nz_new[:] += gaussian_draw
+if test_Tmat:
+    N_tomo = 4
+    zs = s_m.tracers[0].z
+    nz_size = len(zs)
+    cov_m = np.linalg.inv(prior_m)
+    s_new = copy.deepcopy(s_m)
+    for i in range(N_tomo):
+        nz_new = s_new.tracers[i].Nz
+        if i == 2:
[…]

### 2020-05-16 HSC_wrapper.py
new HSC wrapper and modifications to code to use it

@@ -1,0 +1,180 @@
+import sacc
+import copy
+import numpy as np
+import pyccl as ccl
+from hsc_like_mod import HSCLikeModule
+from hsc_core_module import HSCCoreModule
+from desclss.halo_mod_corr import HaloModCorrection
+import yaml
+
+def cutLranges(saccs,saccs_noise):
+    # ell-ranges
+    lmin = [0, 0, 0, 0]
+    lmax = [2170.58958919, 2515.39193451, 3185.36076391, 4017.39370804]
+
+    # do the cuts
+    for i, s in enumerate(saccs):
+        s.cullLminLmax(lmin, lmax)
+        if saccs_noise is not None:
+            saccs_noise[i].cullLminLmax(lmin, lmax)
+    return saccs, saccs_noise
+
+
+def setup(hod_params,z_params):
+    # yaml file
+    config = yaml.load(open("test_like.yaml"))
+
+    # set the HOD parameters to the posterior mean from the paper and the shifts and widths to 0
+    fit_params = config['fit_params']
+    for key in fit_params.keys():
+        if key in hod_params.keys():
+            fit_params[key][1] = hod_params[key]
+        else:
+            fit_params[key][1] = z_params[key]
+
+    # read the parameters in the config file
+    cl_params = config['cl_params']
+    def_params = config['default_params']
+    sacc_params = config['sacc_params']
+
+    # set the parameters
+    param_mapping = {}
+    nparams = len(fit_params.keys())
+    params = np.zeros((nparams, 4))
+    for key in fit_params.keys():
+        param_mapping[key] = fit_params[key][0]
+        params[fit_params[key][0], :] = fit_params[key][1:]
+
+    # update the bias
+    config['default_params'].update(cl_params['bg'])
+
+    return params, param_mapping, sacc_params, cl_params, def_params
+
+
+def _get_cl_theory(s_d,s_n,params,param_mapping,sacc_params,cl_params,def_params,HMCorr,want_Lcuts=False):
+    # sacc files for noise and power spectra
+    saccs = [s_d]
+    saccs_noise = [s_n]
+    Ntomo = len(saccs[0].tracers)
+
+    if want_Lcuts:
+        # do the ell-cutting
[…]

### 2020-05-16 create_sacc.py

@@ -6,18 +6,16 @@ import pyccl as ccl
 import copy
 import shutil
 import matplotlib.pyplot as plt
-from modules.halo_mod_corr import HaloModCorrection
-from modules.theory_cls import get_theory
-from calculate_smooth_s_and_prior import get_smooth_s_and_prior
-import scipy.linalg as la
 
-# Note to self and reader -> this should eventually be run with the real data, i.e.
-# read = 'COADDED'; write = 'MARG'
-# or for realistic NG covariance
+from desclss.halo_mod_corr import HaloModCorrection
+from calculate_smooth_s_and_prior import get_smooth_s_and_prior
+from HSC_wrapper import obtain_Tmat
+from HSC_wrapper import get_theory
+
+# Use realistic NG covariance
 # read = 'NEWCOV_COADDED'; write = 'NEWCOV_MARG'
 # Important REMOVE l-cuts for the real run
 # TODO we are assuming cosmic variance of the tomographic bins is uncorrelated
-# For testing I am using read = 'COADD' and write = 'test_2' 
 
 try:
     # Names of the data read and write directories 
@@ -57,6 +55,16 @@ hod_sigmas = {'lmmin':0.22, 'lmminp':2.0,
               'm0':4.0, 'm0p':5.0,
               'm1':0.27, 'm1p':2.6}
 
+Z_params = {'zshift_bin0':0.000,'zshift_bin1':-0.016,\
+            'zshift_bin2':-0.01,'zshift_bin3':0.01,\
+            'zwidth_bin0':-0.05,'zwidth_bin1':-0.01,\
+            'zwidth_bin2':0.035,'zwidth_bin3':0.05}
+
+Z_params0 = {'zshift_bin0':0.000,'zshift_bin1':-0.0,\
+            'zshift_bin2':-0.0,'zshift_bin3':0.0,\
+            'zwidth_bin0':-0.0,'zwidth_bin1':-0.0,\
+            'zwidth_bin2':0.0,'zwidth_bin3':0.0}
+
 # Here are some choice that we make for the tests we are running
 # TEST #1
 if "1sigma" in write:
@@ -90,21 +98,6 @@ def obtain_improved_cov(cov,Tmat,prior):
[…]

### 2020-05-16 test_like.py

@@ -2,8 +2,7 @@ import sacc
 import copy
 import numpy as np
 import pyccl as ccl
-from modules.halo_mod_corr import HaloModCorrection
-from modules.theory_cls import get_theory
+from HSC_wrapper import get_theory
 from calculate_smooth_s_and_prior import get_smooth_s_and_prior
 
 # HSC modules
@@ -28,7 +27,7 @@ dir_read = "data/"+read
 #dir_read = "data/NEWCOV_MARG_largenoise"
 #dir_read = "data/NEWCOV_MARG_nosmooth_largenoise"
 only_6HOD = bool(int(sys.argv[2]))
-test_Tmat = 0
+test_Tmat = 1
 test_zparams = 0
 verbose = 0
 
@@ -79,6 +78,7 @@ cosmo = ccl.Cosmology(**cosmo_params)
 
 # Load the data whose precision matrix we would like to modify
 s_d = sacc.SACC.loadFromHDF(dir_read+"/power_spectra_wdpj.sacc")
+s_noi = sacc.SACC.loadFromHDF(dir_read+"/noi_bias.sacc")
 
 # Calculate the smooth s_m = (P0+D)^-1 P0 s0 and smooth prior prior_smo = P0+D
 s_m, prior_m = get_smooth_s_and_prior(s_d,cosmo,want_prior=True,A_smooth=0.25,noi_fac=4.)
@@ -88,14 +88,13 @@ s_m, prior_m = get_smooth_s_and_prior(s_d,cosmo,want_prior=True,A_smooth=0.25,no
 s_d.cullLminLmax([0,0,0,0],lmax)
 
 # T matrix
-#Tmat = np.load("Tmat_test_like.npy")
 Tmat = np.load("Tmat_NEWCOV_COADDED.npy")
 
 # Halo Model Correction: have checked that this is the same as what is applied for the MCMC run
 HMCorrection = HaloModCorrection(cosmo, k_range=[1e-4, 1e2], nlk=256, z_range=[0., 3.], nz=50)
 
 # Computing Cls directly from the data
-cl_theory = get_theory(hod_params,cosmo_params, s_d, halo_mod_corrector=HMCorrection)
+cl_theory = get_theory(s_d,s_noi,hod_params,z_params,HMCorrection)
 
 # Get the covariance and precision matrices
 N_data = len(s_d.mean.vector)
@@ -120,14 +119,8 @@ if test_Tmat:
         if i == 2:
[…]

### 2020-05-16 HSC_wrapper.py
name changes

@@ -22,7 +22,7 @@ def cutLranges(saccs,saccs_noise):
 
 def setup(hod_params,z_params):
     # yaml file
-    config = yaml.load(open("test_like.yaml"))
+    config = yaml.load(open("config_6HOD_zparams.yaml"))
 
     # set the HOD parameters to the posterior mean from the paper and the shifts and widths to 0
     fit_params = config['fit_params']

### 2020-05-16 test_like.py

@@ -208,9 +208,9 @@ saccs_noise = [sacc.SACC.loadFromHDF(dir_read+"/noi_bias.sacc")]
 Ntomo = len(saccs[0].tracers) ## number of tomo bins
 
 if only_6HOD:
-    config = yaml.load(open("test_like_6HOD.yaml"))
+    config = yaml.load(open("config_6HOD.yaml"))
 else:
-    config = yaml.load(open("test_like.yaml"))
+    config = yaml.load(open("config_6HOD_zparams.yaml"))
 
 fit_params = config['fit_params']
 # got a raise exception

### 2020-05-18 create_sacc.py
small typo

@@ -204,7 +204,7 @@ for t in s_dp.tracers:
     i+=tl
 
 # Slightly improved computation using taylor expansion
-cl_theory_perturbed = get_theory(s_d,s_noi,hod_params,Z_params,HMCorrection)
+cl_theory_perturbed = get_theory(s_dp,s_noi,hod_params,Z_params,HMCorrection)
 cl_theory_taylor = cl_theory + np.dot(Tmat.T,dNz)
 
 di = s_d.mean.vector - cl_theory_perturbed

### 2020-07-04 Paper/paper.tex
Updates from Overleaf

@@
 \ifpdf
-\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
-             % images in pdf, png or jpg format)
+\pdfoutput=1 
 \fi
@@
 \usepackage{dcolumn}
-\usepackage{amssymb,amsmath,bm}
+\usepackage{amssymb,amsmath,bm,bbold}
 \usepackage{color}
 \usepackage[dvipsnames]{xcolor}
-%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
-%\usepackage[utf8]{inputenc}
 \usepackage{xfrac}
@@
 \newcommand{\nv}{\vec{\theta}}
-%\newcommand{\pasj}{Publications of the ASJ}
-%\newcommand{\aap}{Astronomy and Astrophysics}
-%\newcommand{\apj}{Astrophysical Journal}
-%\newcommand{\apjs}{Astrophysical Journal, Supplement}
-%\newcommand{\mnras}{Monthly Notices of the RAS}
 \newcommand{\todo}[1]{{\bf TODO: #1}}
@@
 
-%\usepackage{lineno}
 %\linenumbers
@@
 
-\section{Introduction}\label{sec:intro}
-
+  \section{Introduction}\label{sec:intro}\label{sec:intro}
     Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky.
@@
 
-\section{Theory}
+  \section{Theory}\label{sec:theory}
+    \subsection{Redshift distribution uncertainties}\label{ssec:theory.nz}
+      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters descring the underlying model.
 
-In typical cosmological analysis problem, we are performing a Bayesian inference on the posterior $p({\rm theory} | {\rm data} ) \propto p({\rm data}|{\rm} theory) p({\rm theory})$ with the data likelihood given by
+      In the case of tomographic analyses of galaxy clustering and large-scale structure, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g.
+galaxy ellipticities or overdensity) and different redshift bins\footnote{We use the term ``redshift bin'' here to denote any given galaxy sample, whether or not it has actually been selected by binning those galaxies into intervals of some redshift estimate.}, and $\vec{\theta}$ includes both cosmological and nuisance parameters needed to produce a forward model of $\vd$ that we will call $\vt(\vec{\theta})$.
+Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian, taking the form
       \begin{equation}
-p({\rm data}|{\rm theory}) =  \mathcal{L} \propto  \exp\left [-\frac{1}{2} \chi^2 \right], \label{eq:like}
+        \mathcal{L} \equiv p({\bf d}|\vec{\theta}) = \frac{\exp\left [-\frac{1}{2}(\vd-\vt(\vec{\theta}))^T{\sf C}^{-1}(\vd-\vt(\vec{\theta})) \right]}{\sqrt{\det{2\pi{\sf C}}}}, \label{eq:like}
      \end{equation}
+      where ${\sf C}$ is the covariance matrix of $\vd$.
+The information contained in the parameter dependence of ${\sf C}$ can be neglected \citep{1811.11584}, and therefore the normalization factor $\sqrt{2\pi\det{\sf C}}$ can be ignored.
+
+      We can make significant progress without specifying explicitly how the theory is calculated given the parameters.
+It suffices to say that, given a set of cosmological parameters, which specify the evolution of the universe and the matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g.
+a particular bias or intrinsic alignment model), and the redshift distributions $N(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$.
+Our main concern in this paper is the way in which the significant uncertainties in $N(z)$, which can potentially dominate the total systematic error budget, should be parametrized and marginalized over.
+
+      The traditional approach (e.g.
+\todo{Add cites}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector.
+This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
+      \begin{equation}
+        N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right), \label{eq:photo-z-model}
+      \end{equation}
+      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width changes.
+$\Delta z_i$ and $z_{w,i}$ for each bin are then inferred and marginalized over together with all other parameters, with priors on them based on calibration uncertainties.
+
+      This approach explicitly marginalizes over two effects that are potentially the two most important systematics stemming from redshift distribution uncertainty: the width of the lens redshift distribution affects the amplitude and shape of the clustering power spectrum, and a shift in the distribution of sources influences the support of the lensing kernel and the amplitude of the shear two-point function.
+However, this methods suffers from two shortcomings.
+First, it is fundamentally ad-hoc: this particular parametric form is not directly mapped onto the main modes of uncertainty in photometric redshift estimators \todo{we may want an expert to vet this statement}, and there is rigorous proof that the parametrization is complete in the sense that it captures all relevant effects.
+The second problem is that it is computationally expensive.
+By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly.
+For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted 50-60\% of all model parameters.
+The methodology described in the next section aims to address both of these shortcomings.
+
+    \subsection{A new approach}\label{ssec:theory.nz_new}
+      Let us consider a discretized description of the $N_i(z)$ as a sum over basis functions $\phi_\alpha(z)$:
+      \begin{equation}
+        N_i(z)=\sum_i N^\alpha_i\phi_\alpha(z).
+      \end{equation}
+      For simplicity here we will just treat $N_i(z)$ as a histogram, and therefore in our case the $\phi_\alpha$ are just top-hat functions centered on an equi-spaced grid of redshifts $z_\alpha$.
+The formalism described below is valid for any choice of basis functions, however.
+In this formalism, full set of redshift distributions is described by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
+In what follows, we will assume that $\vN$ has units of ``counts''.
+The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
+\todo{ref}).
+
+      Let us further distinguish between the true underlying redshift distribution, represented by $\vN$, and a possible measurement of it $\hat{\bf N}$, with measurement uncertainties encapsulated by a covariance matrix $\hat{\sf C}_N$.
+Furthermore, we may have some external prior information on $\vN$, which for simplicity we will assume to be Gaussian with mean $\vN_P$ and covariance $\hat{\sf C}_P$.
+
+      Our data is therefore made up of the combination ${\bf d}=(\hat{\bf c},\hat{\bf N})$, where $\hat{\bf c}$ is a set of two-point function measurements.
+The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redsihft distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters.
+The posterior distribution is therefore given by:
+      \begin{align}
+        p(\vec{\theta}|{\bf d})
+        &=p({\bf q},\vN|\hat{\bf c},\hat{\bf N})\\\label{eq:posterior_full}
+        &=p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
+        &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
+      \end{align}
       where
+      \begin{align}
+        \chi^2_c&=\left(\hat{\bf c}-{\bf t}\right)^T{\sf C}_c^{-1}\left(\hat{\bf c}-{\bf t}\right)\\
+        \chi^2_N&=\left(\hat{\bf N}-\vN\right)^T{\sf C}_N^{-1}\left(\hat{\bf N}-\vN\right)\\
+        \chi^2_P&=\left(\vN-\vN_P\right)^T{\sf C}_P^{-1}\left(\vN-\vN_P\right),
+      \end{align}
+      and ${\bf t}(q,\vN)$ is the theoretical prediction for $\hat{\bf c}$.
+
+      In Eq.
+\ref{eq:posterior_full} we have assumed that $\hat{\bf c}$ and $\hat{\vN}$ are independent at the likelihood level, and that the likelihood of $\hat{\vN}$ is independent of ${\bf q}$.
+The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \todo{cite} or otherwise).
+The method outlined here is still applicable if these assumptions are dropped, albeit with a different expression for the modified covariance in Eq.
+\ref{eq:main2}.
+We have also assumed that the redshift distribution amplitudes $\hat{\vN}$ are Gaussianly distributed.
+\todo{add some blab here and potentially cite Carles Sanchez's latest paper.}
+
+      To derive the main result of this paper, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
       \begin{equation}
-  \chi^2  = (\vd-\vt)^T C^{-1} (\vd-\vt),
+        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\tilde{\vN}\right)^T{\sf P}^{-1}\left(\vN-\tilde{\vN}\right)+K_N,
       \end{equation}
-where $\vd$ is the data vectory, typically containing all the summary statistic of interest like correlation functions ot power spectra and $\vt$ the theory vector through which parameter dependence enters.
-In this set-up the covariance matrix $C$ is fixed (and hence the Gaussian normalization term does not enter equation \ref{eq:like}), which is usually a very good approximation in typical problems that we wish to consider \cite{1811.11584}, but note that the method is easily generalizable to the cases where this is not true.
+      where we have defined the \emph{smoothed mean} $\bar{\vN}$ and combined prior covariance ${\sf P}$ as:
+      \begin{align}\label{eq:priorcov}
+        \bar{\vN}\equiv{\sf P}({\sf C}_N^{-1}\hat{\bf N}+{\sf C}_P^{-1}\vN_P),\hspace{12pt} {\sf P}^{-1}\equiv{\sf C}_N^{-1}+{\sf C}_P^{-1},
+      \end{align}
+      and $K_N$ is independent of the model parameters, and given by
+      \begin{equation}
+        K_N\equiv\hat{\bf N}^T{\sf C}^{-1}_N\hat{\bf N}+\vN_P^T{\sf C}_P^{-1}\vN_P-\bar{\vN}^T{\sf P}^{-1}\bar{\vN}.
+      \end{equation}
+      The full posterior thus takes the form:
+      \begin{align}
+        p({\bf q},\vN|\hat{\bf c},\hat{\bf N})&=Q\,p({\bf q})\exp\left[-\frac{1}{2}(\hat{\bf c}-{\bf t})^T{\sf C}^{-1}_c(\hat{\bf c}-{\bf t})-\frac{1}{2}(\vN-\bar{\vN})^T{\sf P}^{-1}(\vN-\bar{\vN})\right],\\
+        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
+      \end{align}
 
-We can make significant progress without specifying explicitly how theory is calculate given the parameters.
-It suffices to say that given a set of cosmological parameters, which specify the evolution of the universe and thus matter power spectra as a function of scale and redshift, a set of survey properties, such as number densities and redshift distribution of galaxies and weak lensing sources, we can integrate these into theory predictions $\vt=\vt(\theta)$.
-
-The traditional approach is to add to add nuissance parameters that describe our lack of knowledge of galaxy redshift distribution.
-A typical approach is to use shift and width parameters  \cite{1706.09359,1912.08209} \as{did DES Y1 have shift/width?},
-where redshift distributions $N_i(z)$ are shifted according to
-    \begin{equation}
-      N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right),
-      \label{eq:photo-z-model}
-    \end{equation} 
-where the index $i$ runs over the number of tomographic samples.
-Here $N_{i}(z)$ is the true redshift distribution with  $\bar{N}_{i}(z)$ out best estimate of it.
-The parameters $\Delta z_{i}$ and $z_{w,i}$ account for shifts in the means of the distributions, while $z_{c,i} = \left<z\right>_{\bar{N}_i}$ is the mean redshift to act as a sensible pivot for width changes.
-The shift and width parameters for each tomographic bin are then inferred and marginalized over together with all other parameters.
-
-This approach explicitly marginalizes over the two effects that we think are the two most important systematics stemming from redshift distribution uncertainty.
-However, it suffers from two problems.
-First, it is fundamentally ad-hoc.
-This particular parametric form does not stem from consideration of phometric redshift estimators.
-There is also no reason to believe why it would be complete in the sense that it captures all relevant effects.
-The second problem is that it is computationally expensive.
-For example, in the analysis
-    of \cite{1912.08209}, redshift uncertainty parameters accounted for 8 (2 per 4 tomographic bins) of 14 total parameters in the inference process, i.e.
-over the half parameters inferred were dealing with redshift.
-
-\subsection{A new approach}
-
-In this paper we advocate a different approach.
-First, let's consider a binned description of $N_i(z)$ in terms of $N_{ij}$, where $i$ index describes the tomographic bin and $j$ is redshift index, so that $N_{ij} = N_i(z_j)$ at some discrete grid of $z_j$.
-This entire vector can be packed in a vector $\vN$.
-
-We will describe our uncertainty in $\vN$ as a Gaussian
-\begin{equation}
-  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
-\end{equation}
-where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations.
+%We will describe our uncertainty in $\vN$ as a Gaussian
+%\begin{equation}
+%  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
+%\end{equation}
+%where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations.
 We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal).
@@
 
-We can now write a likelihood for the data in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
+      We can now write a likelihood for $\hat{\bf c}$ in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
       \begin{equation}
-\mathcal{L} \propto  \int \exp\left [-\frac{1}{2} (\vd-\vt)^T C^{-1} (\vd-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
+        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T C^{-1} ({\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
       \end{equation}
-  Note that in the equation above, $\vt=\vt(\theta,\vN)$, where $\theta$ are the remaining parameters. It is this dependence that prevents us from performing what looks like a trivial integral.
-
-  In principle, we could solve this problem by making every single element of vector $\vN$ a part of MCMC chain and simply run the MCMC with $\theta$ and additional hundreds of $\vN$ parameters.
-Instead, we pull an Eastern-European trick and Taylor expand theory around $\hat{\vN}$:
+      Note that in the equation above, $\vt=\vt({\bf q},\vN)$.
+It is the non-linear dependence of $\vt$ on $\vN$ that prevents us from performing what looks like a trivial integral analytically.
+In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters \todo{Maybe mention HMC methods that could actually do this}.
+Instead, we pull an Eastern-European trick and Taylor expand theory around $\bar{\vN}$ \todo{we could change this sentence or leave it as is.
+But I think Taylor was British.}:
       \begin{equation}
-    \vt(\theta,\vN) = \vt(\theta,\bar{\vN}) + T \left(\vN - \bar{\vN} \right), \label{eq:taylor}
+        \vt({\bf q},\vN) \simeq \vt({\bf q},\bar{\vN}) + {\sf T} \left(\vN - \bar{\vN} \right), \label{eq:taylor}
       \end{equation}
-  where
+      where the matrix ${\sf T}$ is the gradient of ${\bf t}$ with respect to $\vN$
       \begin{equation}
-    T_{mn} = \frac{\partial t_m}{\partial N_n} \bigg|_{\theta,\hat{\vN}}
+        {\sf T}\equiv \left.\frac{d{\bf t}}{d\vN}\right|_{{\bf q},\bar{\vN}}
       \end{equation}
-  The matrix $T$ is the non-square matrix giving derivatives of theory with respect to each element of $\vN$.
-In other words it contains a full information about how all the correlation functions or power spectra respond to changing $N(z)$ distribution for one tomographic bin at one redshift.
+      In other words it contains a full information about how all the correlation functions or power spectra respond to a small change in $N(z)$.
 
@@
 \ref{eq:taylor} into Eq.
-\ref{eq:like} results in an integral that is now quadratic in $\bar{\vN}$ and can be performed analytically.
-After collecting terms that are quadratic and linear in $\vN$, performing the Gaussian integral and applying the Woodbury matrix identity, we find:
+\ref{eq:like2} results in an integral that is now quadratic in $\vN$ and can be performed analytically.
+After completing squares, and applying the Woodbury matrix identity, we find:
       \begin{equation}
-    \mathcal{L} \propto \frac{1}{|T C^{-1} T^T +P^{-1} |^{1/2}} \exp\left [-\frac{1}{2} (\vd-\vt)^T C_M^{-1} (\vd-\vt) \right], \label{eq:main1}
+        \mathcal{L} \propto \left[{\rm det}\left({\sf T}^T {\sf C}_c^{-1} {\sf T} +{\sf P}^{-1}\right)\right]^{-1/2} \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T {\sf C}_M^{-1} ({\bf c}-\vt) \right], \label{eq:main1}
       \end{equation}
-  where
+      where the  marginalized covariance is
       \begin{equation}
-    C_M = C + T PT^T \label{eq:main2}
+        {\sf C}_M = {\sf C}_c + {\sf T}{\sf P}{\sf T}^T \label{eq:main2}
       \end{equation}
@@
 These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.
-Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far P allows them to go.
+Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go.
+The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether.
+This is the main result of the paper.
 
-  In principle, at every point in likelihood, one should evaluate Eq.
-\ref{eq:main1}, which means recalculating $T$.
-Since this is a very expensive computation, we will further simplify the problem, by assuming $T$ is only weakly deviations from some fiducial cosmology.
-We will therefore pre-compute $T$ once for the fiducial cosmology, which then means that normalization term in equation \ref{eq:main1} is an irrelevant constant and at fixed data covariance matrix $C$, the Eq.
-\ref{eq:main2} also needs to be evaluated just once.
+      Both the marginalized covariance ${\sf C}_M$ normalizing prefactor in Eq.
+\ref{eq:main1} depend in principle on ${\bf q}$ through the parameter dependence of ${\sf T}$.
+This implies that, in principle, ${\sf T}$ should be re-evaluated at every point in the MCMC chain.
+Since the calculation of ${\sf T}$ is expensive (by comparison with that of e.g.
+$\vt$) using standard methods, we will neglect this dependence and evaluate ${\sf T}$ at a fiducial set of parameters.
+We will however explore the impact of the choice of fiducial parameters on the final results.
+This should be a sufficiently good approximation for compact likelihoods, where parameter constraints are driven by the differences between $\hat{\bf c}$ and $\vt$.
+We will therefore pre-compute $T$ once, which then means that for a fixed data covariance matrix ${\sf C}_c$, the normalization term in equation \ref{eq:main1} is an irrelevant constant, and the marginalized covariance in the Eq.
+\ref{eq:main2} also needs to be evaluated only once.
 
-  The procedure is thus as follows:
+      The procedure to obtain constraints from this marginalized likelihood would therefore proceed as follows:
       \begin{itemize}
-  \item Run inference problem once with fixed redshift distribution to determine a sensible fiducial model.
+        \item Run inference problem once with fixed redshift distribution and a standard Gaussian likelihood for $\hat{\bf c}$ with covariance ${\sf C}_c$ to determine a sensible fiducial model.
+        \item Calculate ${\sf T}$ and ${\sf C}_M$ at this fiducial model.
+        \item Re-run inference with a standard Gaussian likelihood using the modified covariance matrix ${\sf C}_M$.
+This will lead to broadened contours due to the marginalization over redshift distribution uncertainties.
+      \end{itemize}
+      If needed, one could repeat the inference at a refined fiducial model although in practice we found this to be unnecessary.
 
-  \item Calculate $T$ and $C_M$ at this fiducial model
-
-  \item Re-run inference with the modified covariance matrix $C_M$, which leads to broadened contours due to marginalization of redshift distributions
-
-  \end{itemize}
-If needed, one could repeat the inference at a refined fiducial model although in pracice we found this to be unnecessary.
-
-Note that there are two distinct approximation here.
-The first is that theory can validly be Taylor expanded in $\vN-\bar{\vN}$ and the second is that $T$ is model independent for models of interest.
+      Before moving on, it is worth emphasizing the two distinct approximation used here.
+The first one is that the theory can validly be Taylor expanded in $\vN-\bar{\vN}$.
+The second is that the model dependence of ${\sf T}$ can be ignored for the models of interest.
 We will examine validity of both of those approximations in Section \ref{sec:hsc}.
 
-\subsection{Determination of prior matrix $P$}
+    \subsection{The prior matrix ${\sf P}$}\label{ssec:theory.prior}
+      A key part of this method is the determination of the $N(z)$ prior covariance ${\sf P}$, which governs the amplitude of the uncertainties on $\vN$.
+As discussed in the previous section, ${\sf P}$ receives two contributions: the covariance associated with the uncertainties in the measured $\hat{\bf N}$, ${\sf C}_N$, and the external prior with covariance ${\sf C}_P$, both combined in an inverse-variance way (see Eq.
+\ref{eq:priorcov}).
+We will describe the models used for both contributions in this section.
 
-Determination of the prior matrix $P$ is where the the actual and often partially subjective prior on our knowledge of $N(z)$ enters.
+      \subsubsection{$N(z)$ uncertainties}\label{sssec:theory.prior.cv}
+        \begin{figure}[ht]
+          \centering  
+          \includegraphics[width=1.\textwidth]{./corr_CV_0.png}
+          \caption{Correlation matrix associated to the cosmic-variance contribution to the total $N(z)$ covariance for the first tomographic bin of the HSC data.
+\todo{Can we make these plots pdfs instead?}}\label{fig:CV}
+        \end{figure}
+        The analysis presented in Section \todo{add ref} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \todo{add ref}, as was done in \todo{add cite}.
+The main sources of uncertainty for this measurement are cosmic variance in the COSMOS field, shot noise and additional uncertainty in the photometric redshifts used to assign galaxies to different bins.
+Here we will associate each of this sources to a separate covariance that will be added in quadrature to form the final ${\sf C}_N$.
+
+        The contribution from the photometric redshift systematic uncertainties is modelled as \ldots\todo{Not sure how you guys implemented this.}
+
+        The cosmic variance uncertainties are caused by the fluctuations in the matter density traced by galaxies in the particular sky patch.
+The corresponding covariance would ideally be estimated from simulations including both non-linear gravitational clustering and a realistic model of the galaxy-halo connection for the specific galaxy sample under study.
+For simplicity in this proof-of-concept analysis we instead use an analytical model for the $N(z)$ covariance.
+Although less precise than a simulation-based calculation, this approach was shown by \todo{cite sanchez} to provide a reasonable prediction for the redsthift distribution uncertainties.
+
+        The covariance matrix element between $N^\alpha_i$ and $N^\beta_j$ two tomographic bins is given by:
+        \begin{equation}\label{eq:nz_cv}
+          {\sf C}_{N,{\rm CV}}^{(i\,\alpha),(j\,\beta)}=\frac{N^\alpha_iN^\beta_j}{2\pi^2} \int_0^\infty dk_\parallel \cos(k_\parallel(\chi_\alpha-\chi_\beta))\int_0^\infty dk_\perp k_\perp W_\alpha(k_\parallel,k_\perp)W_\beta(k_\parallel,k_\perp)P_{gg}({\bf k}),
+        \end{equation}
+        where $\chi_\alpha$ is the radial comoving distance to redshift $z_\alpha$, and we model the galaxy power spectrum using the Kaiser formula \todo{cite} accounting for redshift-space distortions:
+        \begin{equation}
+          P_{gg}({\bf k},z)=\left(b_g(z)+f(z)\frac{k_\parallel^2}{k^2}\right)^2\,P_{mm}(k,z),
+        \end{equation}
+        where $f(z)$ is the logarithmic growth rate and $b_g(z)$ is the linear galaxy bias.
+Following the results of \todo{cite}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor.
+$P_mm(k)$ is the non-linear matter power spectrum, which we model via the \todo{cite} HALOFIT parametrization.
+
+        The overlapping deep HSC patch covering the COSMOS 30-band footprint used to measure the redshift distribution in this work covers a total area of $A_{\rm sky} = 1.7 \ {\rm deg}^2$.
+Modelling this patch as a disc of radius $\theta_{\rm sky}=0.73^\circ$, the corresponding window function in Eq.
+\ref{eq:nz_cv} is given by
+        \begin{equation}
+          W_\alpha(k_\parallel,k_\perp)=j_0(k_\parallel\Delta\chi_\alpha/2)\,\frac{2 J_1(k_\perp \chi_\alpha\theta_{\rm sky})}{k_\perp \chi_\alpha\theta_{\rm sky}},
+        \end{equation}
+        where $\Delta\chi_\alpha$ is the comoving width of the $\alpha$-th histogram bin, $j_0$ is the zero-th order spherical Bessel function and $J_1$ is the order-1 cylindrical Bessel function.
+
+        We assume that the cosmic covariance between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by treating each bin independently.
+We show the resulting correlation matrix of the COSMOS cosmic variance in Fig.
+\ref{fig:CV}.
+
+        The final contribution, from shot noise, is caused by the discrete nature of galaxies as a tracer of the matter fluctuations, and is simply given by
+        \begin{equation}
+          {\sf C}^{(i\,\alpha),(j\,\beta)}_{N,{\rm SN}}=\delta_{\alpha\beta}\delta_{ij}N^\alpha_i.
+        \end{equation}
+        We find this contribution to be subdominant in all cases in comparison with cosmic variance.
+
+    \subsubsection{External priors and smoothness}\label{sssec:theory.prior.smooth}
+      In addition to information from direct measurement of the $N(z)$, it is reasonable to impose certain properties on the underlying true reshift distributions based on physical considerations.
+For instance, there is no reason to expect that the physics of galaxy formation should generate sharp features in the redshift evolution of the abundance of galaxies.
+The transition of features in the spectra of different galaxy types between photometric bands at different redshifts could induce smaller-scale fluctuations in the $N(z)$.
+However, for a sufficiently diverse galaxy sample, one would not expect such fluctuations on scales $\delta z\lesssim0.04$ (corresponding to $\sim100\,{\rm Mpc}$ or $\sim0.2\,{\rm Gyr}$ at $z\sim1$).
+It is therefore a reasonable proposition to impose a certain degree of smoothness on the redshift distributions, which can be achieved through a purposely defined Gaussian prior.
+Applying these types of priors is admittedly a subjective choice to some extent, and we will study its impact on our results in Section \todo{ref}.
+
+      One common way to impose smoothness on a function $f(x)$ is to penalize large values of its firs derivative via $p(f)\propto\exp\left[-(f'(x))^2/(2\sigma_1^2)\right]$.
+In the discrete formalism used here, this is equivalent to imposing a Gaussian prior on $\vN$ with zero mean ($\vN_P=0$) and an inverse covariance given by:
+      \begin{equation}\label{eq:prior_1st}
+        {\sf C}^{-1}_P=\frac{1}{\sigma_1^2}\sum_\alpha {\bf v}^{1T}_\alpha{\bf v}^1_\alpha,
+      \end{equation}
+      where
+      \begin{equation}
+        ({\bf v}^1_\alpha)_\beta=\left\{
+        \begin{array}{ll}
+          1  & {\rm if}\,\,\, \alpha=\beta+1\\
+          -1  &  {\rm if}\,\,\, \alpha=\beta\\
+          0 & {\rm otherwise}
+        \end{array}\right.
+      \end{equation}
+      To understand this, note that, using finite differences, the derivative of $\vN$ is approximately $\vN'\propto {\bf v}^{1T}\vN$.
+
+      The prior in Eq. \ref{eq:prior_1st} effectively penalizes large deviations between adjacent elements of $\vN$. This can be generalized to include all element pairs as
+      \begin{equation}\label{eq:smooth1}
+        {\sf C}^{-1}_P=\sum_{n=1}\sum_{\alpha}p_n\,{\bf v}^{n\,T}_\alpha{\bf v}^n_\alpha,
+      \end{equation}
+      where
+      \begin{equation}
+        ({\bf v}^n_\alpha)_\beta=\left\{
+        \begin{array}{ll}
+          1  & {\rm if}\,\,\, \alpha=\beta+n\\
+          -1  &  {\rm if}\,\,\, \alpha=\beta\\
+          0 & {\rm otherwise}
+        \end{array} \right..
+      \end{equation}
+      The prefactor $p_n$ penalizes differences between neighbors of order $n$, and should therefore be a monotonically decreasing function of $n$ (since there should be no correlation between widely separated histogram bins).
+We therefore choose a functional form
+      \begin{equation}
+        p_n=A\,{\rm exp}\left[-\frac{1}{2}\left(n\frac{\Delta z}{\Delta z_{\rm thr}}\right)^2\right],
+      \end{equation}
+      where $\Delta z$ is the redshift separation between neighbouring histogram bins, and $\Delta z_{\rm thr}$ marks the redshift separation beyond which the $N(z)$ is expected to be uncorrelated.
+In our analysis we choose $A$ and $\Delta z_{\rm thr}$ as \todo{describe and motivate this}.
+
+      %Determination of the prior matrix $P$ is where the the actual and often partially subjective prior on our knowledge of $N(z)$ enters.
 Note that the standard parameterizations can always be recast in this language at least in the perturbative limits.
@@
 
-\subsubsection{Uncertainty due to estimator biases}
-We will build $P$ from three pieces.
-The first piece is simply the overall uncertainty stemming from the fact that the photometric redshift estimators have errors.
-The simplest thing is to make this covariance matrix   \ldots
-
-\subsubsection{Cosmic Variance in calibrator sample}
-The covariance matrix between two tomographic bins is given by:
-\begin{equation}
-Cov_{ij}=\frac{N_i N_j}{2\pi^2} \int_0^\infty dk_\parallel \cos(k_\parallel(\chi_i-\chi_j))\int_0^\infty dk_\perp k_\perp W_i(k_\parallel,k_\perp)W_j(k_\parallel,k_\perp)P_{\bf gg}({\bf k}),
-\end{equation}
-where we can model the power spectra at each bin centre and for all
-transverse and parallel modes using the Kaiser formula (CITE), which corrects for redshift space distortions through
-the expression:
-\begin{equation}
-P_{gg}({\bf k},z)=\left(b_g(z)+f(z)\frac{k_\parallel^2}{k^2}\right)^2\,P_{mm}(k,z),
-\end{equation}
-where $f(z)$ is
-the logarithmic growth
-rate
-and the
-evolution of the linear galaxy bias is given by $b_g(z) = 0.95/D(z)$
-with $D(z)$ the growth factor.
-
-In this work, we use the galaxy sample in the COSMOS survey as calibrator, which
-covers a total area of
-$A_{\rm sky} = 1.7 \ {\rm deg}^2$ for galaxies with redshifts $z < 2$.
-We model the window function for each bin of the dataset through:
-\begin{equation}
-W(k_\parallel,k_\perp)=j_0(k_\parallel\Delta\chi/2)\,\frac{2 J_1(k_\perp R_{\rm mean})}{k_\perp R_{\rm mean}},
-\end{equation}
-where $R_{mean}$ is the
-physical size of the survey measured at the mean
-distance to the
-tomographic bin, 
-$R_{\rm mean} = 
-\sqrt{\frac{A_{\rm sky}}{\pi}} \chi_{\rm mean}$.
-We assume that the
-cosmic covariance
-between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by
-treating each bin independently. We show the resulting correlation matrix of the COSMOS cosmic variance in Fig. \ref{fig:CV}.
-
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./corr_CV_0.png}
-\caption{Cosmic variance (CV) correlation matrix with subtracted diagonal for the first tomographic bin of the HSC data.} 
-\label{fig:CV}
-\end{figure}
-
-\subsubsection{Smoothness prior}
-
-\section{Appplication to HSC clustering data}
-\label{sec:hsc}
+  \section{Appplication to HSC clustering data}\label{sec:hsc}
     \an{Some citations are currently missing but I will put them in later.}
@@
 In the following, we give a very brief summary of the methodology employed in this work and refer the reader to Ref.~\cite{1912.08209} for more details.
-\subsection{Background theory}
-The angular clustering signal for galaxies in redshift bins $i, j$ with photo-$z$ distributions $p^i(z), p^j(z)$ can be modeled using the Limber approximation as \cite{1953ApJ...117..134L, 1992ApJ...388..272K, Kaiser:1998}
+    \subsection{Background theory}\label{ssec:hsc.theory}
+      The angular clustering power spectrum for galaxies in redshift bins $i$, $j$ with can be modeled using the Limber approximation as \cite{1953ApJ...117..134L, 1992ApJ...388..272K, Kaiser:1998}
       \begin{equation}\label{eq:cell_gg_limber}
@@
     where $P_{gg}(z,k)$ denotes the underlying 3D galaxy power spectrum, $\chi(z)$ is the comoving distance and $H(z)$ denotes the Hubble parameter at redshift $z$.
-As can be seen from Eq.~\ref{eq:cell_gg_limber}, the spherical harmonic power spectrum $C_{\ell}$ depends linearly on the photometric redshift distributions $p^i(z)$, which facilitates the computation of the $T$ matrix defined above.
+$p^i(z)$ is the redshift probability distribution of bin $i$ normalized to unit area, and is therefore related to the unnormalized distribution via
+    \begin{equation}
+      p^i(z)=\left[\int dz N_i(z)\right]^{-1}N(z)=\frac{\sum_\alpha N^\alpha_i\phi_\alpha(z)}{\sum_\alpha N^\alpha_i\int dz\phi_\alpha(z)}.
+    \end{equation}
+    The simplicity with which the redshift distribution enters the prediction for the angular power spectrum in Eq.~\ref{eq:cell_gg_limber}, facilitates the computation of the ${\sf T}$ matrix defined above.
 
-We follow Ref.~\cite{1912.08209} and compute theoretical predictions for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}.
+    Finally, following Ref.~\cite{1912.08209} we estimate the theoretical prediction for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}.
+Details about HOD parameterizations can be found in these references, and we only provide a succinct description relevant to the present analysis here
+
+    The galaxy power spectrum receives contributions from the so-called 1-halo and 2-halo terms:
+    \begin{equation}
+      P_{gg}(z,k) = P_{gg,{\rm 1h}}(z,k) + P_{gg,{\rm 2h}}(z,k),
+    \end{equation}
+    where
+    \begin{align}
+      & P_{gg,{\rm 1h}}(k)=\frac{1}{\bar{n}_g^2} \int \mathrm{d}M\,\frac{\mathrm{d}n}{\mathrm{d}M} \bar{N}_c\,\left[\bar{N}_s^2u_s^2(k)+2\bar{N}_su_s(k)\right],\\
+      & P_{gg,{\rm 2h}}(k)=\left(\frac{1}{\bar{n}_g} \int \mathrm{d}M\,\frac{\mathrm{d}n}{\mathrm{d}M}\,b_h(M)\,\bar{N}_c\,\left[1+\bar{N}_su_s(k)\right]\right)^2\,P_{\rm lin}(k).
+    \end{align}
+
+    Here, $\mathrm{d}n/\mathrm{d}M$ is the halo mass function for halo mass $M$, $b_h(M)$ is the linear halo bias, $N_c(M)$ and $N_s(M)$ are the mean number of central and satellite galaxies in halos of mass $M$, $u_s(k)$ is the Fourier transform of the satellite density profile, and $P_{\rm lin}(k)$ is the linear matter power spectrum.
+The number density of galaxies is calculated as
+    \begin{equation}
+      \bar{n}_g=\int \mathrm{d}M\,\frac{\mathrm{d}n}{\mathrm{d}M}\bar{N}_c(M)\left[1+\bar{N}_s(M)\right].
+      \label{eq:ng_hod}
+    \end{equation}
+
+    As in \cite{1912.08209}, we parametrize the number of centrals and satellites as a function of mass as:
+    \begin{align}
+      &\bar{N}_c(M)=\frac{1}{2}\left[1+{\rm erf}\left(\frac{\log_{10}(M/M_{\rm min})}{\sigma_{\log M}}\right)\right],\\
+      &\bar{N}_s(M)=\Theta(M-M_0)\left(\frac{M-M_0}{M_1'}\right)^\alpha,
+    \end{align}
+    where $\Theta(x)$ is the Heavyside step function, and we model the distribution of satellites to follow that of the dark matter, given by a truncated Navarro-Frenk-White profile \cite{Navarro:1996}.
+The choices of mass function parametrization, halo bias and concentration-mass relation used here follow the same models used in \cite{1912.08209}, namely the mass function and halo bias of \cite{Tinker:2010}, and the concentration-mass relation of \cite{Duffy:2008} for a spherical overdensity halo masses with an overdensity parameter $\Delta=200$ with respect to the matter density.
+
+    The HOD model is defined by three characteristic masses, $M_{\rm min}$, $M_0$ and $M_1$.
+We model the redshift dependence of these masses as a linear Taylor expansion in the scale factor around the mean redshift of the sample $z_p=0.65$ as
+    \begin{equation}
+      \log_{10}{M_x(z)} = \mu_x + \mu_{x, p} \left(\frac{1}{1+z} - \frac{1}{1+z_{p}}\right).
+    \end{equation}
+    where $x$ is $\mathrm{min}$, 0 or 1.
+
+    Besides these HOD parameters, the analysis of \cite{1912.08209} marginalized over uncertainties in the redshift distribution using the shift-width parametrization of Eq.
+\ref{eq:photo-z-model}, including two additional shift and width parameters per redshift bin.
+For the four redshift bins described in section \todo{ref}, the complete set of 14 free parameters is
+    \begin{equation}
+      \vec{\theta}=\{\mu_{\rm min},\,\mu_{{\rm min},p},\,\mu_0,\,\mu_{0,p},\,\mu_1,\,\mu_{1,p},\,\Delta z_{\{1,2,3,4\}},\,z_{w,\{1,2,3,4\}}\}.
+    \end{equation}
+
 \subsection{The HSC dataset}
@@
 \label{fig:Tmat}
+\end{figure}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.2\textwidth]{./Nzs.png}
+\caption{Photometric redshift
+distributions $N(z)$
+of the four tomographic HSC bins.
+The \textit{red} curves show the
+default values used in the 
+full HSC analysis (CITE Andrina),
+while the \textit{blue} curves 
+show their smoothed and upsampled 
+version used in this work. The
+\textit{green} show a random 
+Gaussian draw from the smoothing 
+prior described in Section TODO: Section}
+\label{fig:Nzs}
+\end{figure}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.2\textwidth]{./Cls.png}
+\caption{Galaxy power spectrum, $C_\ell^{gg}$, between the four
+tomographic bins. The \textit{black} curves show the HSC-measured
+power spectrum and error bars, while the
+\textit{red} curves present the $C_\ell$s
+computed using the posterior mean values of the 6 HOD parameters from the
+full HSC analysis (CITE Andrina). In 
+\textit{green} is shown the power
+spectrum obtained by using the smoothed
+$N(z)$ distributions (see Fig. \ref{fig:Nzs}). The \textit{blue}
+curves correspond to the exact
+calculation of the $C_\ell$s using the
+perturbed $N(z)$ values in
+Fig. \ref{fig:Nzs}, and the 
+\textit{lightblue} dotted lines
+show the power spectra derived via
+Taylor expansion: $C_\ell' = C_\ell +
+ {\sf T} \left(\vN' - {\vN} \right)$.
+TODO: put in a table
+Chi2 original model, original cov: 98.97
+Chi2 original model, new cov  89.12
+m1, m1p =  13.05 0.79
+Chi2 original model, smoothed Nz, original cov: 91.33
+Chi2 original model, smoothed Nz, new cov  86.25
+Chi2 perturbed model smooth, original cov: 106.11
+Chi2 perturbed model smooth Nz, new cov  88.37
+Chi2 taylor model smooth nz, original cov: 93.48
+Chi2 taylor model smooth , new cov  86.04
+Ratio of determinants of original precision to CV+noise+smooth precision per dof =  1.145}
+\label{fig:Cls}
 \end{figure}
@@
 \label{fig:triangle_marg_tests}
+\end{figure}
+
+\begin{figure}[ht]
+\centering  
+\includegraphics[width=1.\textwidth]{./importance_noburn.png}
+\caption{Same as Fig.
+\ref{fig:triangle_fid_marg} except 
+here we apply importance sampling 
+to the full 14-parameter chain 
+imposing Gaussian priors on the
+shift parameters with variances
+[0.008,0.006,0.006,0.007] and the
+width parameters with
+variances [0.05,0.045,0.04,0.03],
+derived by estimating their
+distribution due to 
+cosmic variance.}
+\label{fig:triangle_fid_marg}
 \end{figure}
@@
 
+\appendix
+\section{Oversampling}
+We may want to sub-sample the originally measured $N_{\alpha}$ ($\alpha\in[0,N_z^{\rm coarse})$) into a finer grid $n_{\mu}$ ($\mu\in[0,N_z^{\rm fine})$).
+This is essentially interpolation, which we can write as a linear operator $n_\mu = O_{\mu\alpha}N_{\alpha}$.
+The CV covariance of the $n_\mu$ is then given by ${\rm Cov}^{\rm fine}_{\mu\nu}=O_{\mu\alpha}O_{\nu\beta}{\rm Cov}^{\rm coarse}_{\alpha\beta}$.
+
+For nearest-neighbor interpolation, the kernel $O_{\mu\alpha}$ is simply $O_{\mu\alpha}=\delta_{\alpha\alpha_\mu}$, where $\alpha_\mu$ is the index of the $N_\alpha$ that lies closest to $n_\mu$.
+In practice, the easiest thing to do will be to just use the same linear interpolation function (e.g.
+scipy's {\tt interp1d}) on all the rows of ${\rm Cov}^{\rm coarse}$ and then all the columns of the resulting matrix.
+
 \bibliography{bibliography}

### 2020-07-07 Paper/paper.tex
latest plots
Updates from Overleaf

@@
         {\sf T}\equiv \left.\frac{d{\bf t}}{d\vN}\right|_{{\bf q},\bar{\vN}}
+        \label{eq:Tmat}
       \end{equation}
@@
           \centering  
-          \includegraphics[width=1.\textwidth]{./corr_CV_0.png}
-          \caption{Correlation matrix associated to the cosmic-variance contribution to the total $N(z)$ covariance for the first tomographic bin of the HSC data.
-\todo{Can we make these plots pdfs instead?}}\label{fig:CV}
+          \includegraphics[width=1.\textwidth]{./corr_CV_0}
+          \caption{Correlation matrix associated with the cosmic-variance contribution to the total $N(z)$ covariance for the first tomographic bin of the HSC data.}\label{fig:CV}
         \end{figure}
@@
 \centering  
-\includegraphics[width=1.2\textwidth]{./Tmat.png}
+\includegraphics[width=1.\textwidth]{./Tmat}
 \caption{Visualization of the derivative matrix $T_{mn} = {\partial t_m}/{\partial N_n}$. \todo{Make look nicer: split into tomo bins}} 
@@
 \centering  
-\includegraphics[width=1.2\textwidth]{./Nzs.png}
+\includegraphics[width=1.\textwidth]{./Nzs}
 \caption{Photometric redshift
@@
 \centering  
-\includegraphics[width=1.2\textwidth]{./Cls.png}
+\includegraphics[width=1.\textwidth]{./Cls}
 \caption{Galaxy power spectrum, $C_\ell^{gg}$, between the four
@@
 Taylor expansion: $C_\ell' = C_\ell +
- {\sf T} \left(\vN' - {\vN} \right)$.
-TODO: put in a table
-Chi2 original model, original cov: 98.97
-Chi2 original model, new cov  89.12
-m1, m1p =  13.05 0.79
-Chi2 original model, smoothed Nz, original cov: 91.33
-Chi2 original model, smoothed Nz, new cov  86.25
-Chi2 perturbed model smooth, original cov: 106.11
-Chi2 perturbed model smooth Nz, new cov  88.37
-Chi2 taylor model smooth nz, original cov: 93.48
-Chi2 taylor model smooth , new cov  86.04
-Ratio of determinants of original precision to CV+noise+smooth precision per dof =  1.145}
+ {\sf T} \left(\vN' - {\vN} \right)$.}
 \label{fig:Cls}
@@
 
+\begin{table}
+\begin{center}
+ \begin{tabular}{c | c c } 
+ \hline\hline
+ Model & HSC cov. & Marginalized $N(z)$ cov. \\ [0.5ex] 
+ \hline
+  $\chi^2$, original $N(z)$, evaluated at $C_\ell$ & 98.97 & 88.97 \\
+
+ $\chi^2$, smoothed $N(z)$, evaluated at $C_\ell$ & 91.33 & 86.44 \\
+
+ $\chi^2$, perturbed $N(z)$, evaluated at $C_\ell'$ & 99.78 & 89.57 \\
+
+ $\chi^2$, perturbed $N(z)$, taylor-expanded $C_\ell +
+ {\sf T} \Delta \vN$ & 100.62 & 89.64 \\
+
+ \hline
+ \hline
+\end{tabular}
+\end{center}
+\label{tab:lin}
+\caption{$\chi^2$ values with respect to the HSC data of the four curves shown in Fig. \ref{fig:Cls}. $[\mu_1, \ \mu_{1,p}] =  [13.05, 0.79]$
+Ratio of determinants of original precision to CV+noise+smooth precision per dof =  1.15
+94}
+\end{table}
+
+Fig. \ref{fig:Nzs} shows the distribution of the
+photometric redshifts of the HSC data
+
+Fig. \ref{fig:Cls} displays the exact $C_\ell$ vs. 
+linear prediction using the derivative matrix, ${\sf T}$.
+
+In Fig. \ref{fig:Tmat}, we show the
+derivative matrix, ${\sf T}$,
+defined in Eq. \ref{eq:Tmat} for the first tomographic
+bin in the HSC data and its cross-correlations with
+all four bins, i.e. . The rows of the matrix correspond
+to the $\ell$ multipoles, while the columns to the
+redshift samples of $N(z)^1$.
+
+\subsection{Impact of N(z) uncertainties on final parameters}
+To study the effect on the covariance matrix
+of the marginalization procedure described in detail
+in sections TODO: CITE, we visualize the original
+correlation matrix and its difference with the
+marginalized matrix in Fig. \ref{fig:fid_marg_cov}. The
+main effects on the original matrix is
+a positive contribution to the off-diagonal entries
+for redshift samples in close proximity. We set the free
+parameters of our model to the following values: 
+$A_{\rm smooth} = 1$, $A_{\rm noise} = 4.$,
+$N_{\rm upsample} = 3$, and $\Delta z_{\rm thresh} = 
+0.06$. These values were chosen so as to reduce the
+$\chi^2$ of the perturbed model 
+using the new covariance matrix when compared with the
+original covariance matrix. The modest value of the
+smoothing parameter, $A_{\rm smooth}$, ensure that we 
+are not oversmoothing the $N(z)$ distribution, while
+the choice of photometric noise as four times 
+larger than the variance in the different photometric
+measures provides a reasonable guess for the expected
+variation in $N(z)$ given the uncertainty.
+
+To assess whether the marginalization procedure 
+proposed in this work provides better constraints
+on the
+6 HOD parameters than the original 14-parameter
+model, we employ the ensemble sampler for Markov
+chain Monte Carlo (MCMC) $\textsc{emcee}$ CITE Goodman 
+\& Weare (2010) and other paper with 6 parameters and
+a walkers' ratio of 6 for 12000 iterations. We assess
+the convergence of the chains with a simple G-R test
+having computed the auto-correlation and thinned the
+chains beforehand. In all cases, we find that the
+chains are reasonably well converged.
+
+In Fig. \ref{fig:fid_marg_cov}, we show the key 
+comparison between
+
+We can
+see that the full analysis leads to 
+worse constraints on the parameters
+when compared with the marginalized
+covariance. Furthermore, the proximity
+of the \textit{blue} and \textit{red} 
+contours serves both as a sanity check 
+that our method yields physical constraints
+of the 6 parameters (i.e. not exceeding
+the cosmic variance limits) as well as 
+nearly optimal such
+
+We, furthermore, run the following several 
+tests to check the robustness of our result:
 \begin{itemize}
-  \item Plot showing $T$
-  \item Plot showing exact $C_\ell$ vs linear prediction. \todo{Make your plots into subplots; verify the Tmatrix}
-\end{itemize}{}
-\subsection{Impact of N(z) uncertainties on final parameters}
+\item We study the effect of the smoothing
+by turning it off, $A_{\rm smooth} = 0$. 
+\item We vary the choice for the
+diagonal photometric noise parameter
+by increasing it to $A_{\rm noise} = 42$.
+\item We compute the ${\sf T}$ matrix 
+using a different set of 
+HOD parameters perturbed by $1\sigma$
+away from the fiducial ones.
+\end{itemize}
+We show these comparisons with the
+fiducial marginalized $N(z)$ model in Fig.
+\ref{fig:triangle_marg_tests}.
+One can notice that the effect of the
+smoothing on the constraints is minimal,
+while increasing the diagonal noise
+leads to weakening them. Neither of these
+tests changes significantly the
+mean of the posterior distribution.
+On the other hand, offsetting the 
+HOD parameters used in the ${\sf T}$
+matrix computation leads to a shift to
+the mean and a slight widening of the
+contours. That is because the covariance
+matrix used in the analysis no longer
+favors the HOD values preferred by the
+data. This slight tension
+results in sub-par convergence
+of the chains and shifted mean posteriors.
 
+\ref{fig:triangle_fid_marg}
 \begin{figure}[ht]
@@
   \centering
-  \includegraphics[width=1.\linewidth]{./corr_data.png}
-  \label{fig:sub1}
+  \includegraphics[width=1.\linewidth]{./corr_data}
 \end{subfigure}%
@@
   \centering
-  \includegraphics[width=1.\linewidth]{./corr_diff.png}
-  \label{fig:sub2}
+  \includegraphics[width=1.\linewidth]{./corr_diff}
 \end{subfigure}
-\caption{Correlation matrix for the fiducial HSC dataset (\textit{left}) and difference with its marginalized version (\textit{right}).}
+\caption{Correlation matrix of the original HSC dataset (\textit{left} panel) and difference between the
+original and the marginalized (this work)
+correlation matrices (\textit{right} panel). Most of the
+effect of marginalizing for the shift and width
+parameters is manifested as a positive contribution
+to the original matrix near the diagonal (i.e. there
+is an increase in correlation for close redshift
+values). Note that we have subtracted the diagonal
+off the two matrices.}
 \label{fig:fid_marg_cov}
@@
 \centering  
-\includegraphics[width=1.\textwidth]{./triangle_fid_marg.png}
-\caption{Triangle plot with constraints of the 6 HOD parameters obtained using the covariance matrix from the full HSC dataset (\textit{red}) and from its marginalized version (\textit{blue}).}
+\includegraphics[width=1.\textwidth]{./triangle_fid_marg}
+\caption{Triangle plot with constraints of 
+the 6 HOD parameters obtained using the 
+covariance matrix from the full
+14-parameter original HSC analysis 
+(\textit{gray contours}). The \textit{red
+contours} are obtained by applying
+importance sampling to the full original 
+chain with weights determined by the
+cosmic variance constraints of the
+shift and width parameters. In \textit{blue},
+we show the constraints on the 6 HOD parameters
+obtained using the covariance
+matrix marginalized to account for the
+$N(z)$ uncertainty (i.e. without the
+8 shift and width parameters). We can
+see that the full analysis leads to 
+worse constraints on the parameters
+when compared with the marginalized
+covariance.}
 \label{fig:triangle_fid_marg}
@@
 \centering  
-\includegraphics[width=1.\textwidth]{./triangle_marg_tests.png}
-\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green} is shown the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from \todo{CITE Andrina's paper}; in \textit{red} is shown the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally in \textit{blue} is shown the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 0.25$.}
+\includegraphics[width=1.\textwidth]{./triangle_marg_tests}
+\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green}, we show the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from \todo{CITE Andrina's paper}; in \textit{red}, we display the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally the \textit{blue} contours show the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 1$.}
 \label{fig:triangle_marg_tests}
@@
 
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./importance_noburn.png}
-\caption{Same as Fig.
-\ref{fig:triangle_fid_marg} except 
-here we apply importance sampling 
-to the full 14-parameter chain 
-imposing Gaussian priors on the
-shift parameters with variances
-[0.008,0.006,0.006,0.007] and the
-width parameters with
-variances [0.05,0.045,0.04,0.03],
-derived by estimating their
-distribution due to 
-cosmic variance.}
-\label{fig:triangle_fid_marg}
-\end{figure}
-
-Plots showing the effects of different things:
-\begin{itemize}
-\item Compare different choices of prior (smoothness, no smoothing, different degrees of noise etc.)
-\item Triangle plot with T matrix computed from a different fiducial.
-\item Compare CV-only covariance vs. no covariance.
-\item Compare old marginalization vs. new one (maybe same plot).
-\item Compare old marginalization with CV priors (importance sample).
-\end{itemize}
-
-Other results:
-\begin{itemize}
-  \item smoothness does nothing
-  \item quote $\chi^2$ values \todo{Make a table}
-\end{itemize}
-
+\begin{table}
 \begin{center}
- \begin{tabular}{c | c c c} 
+\begin{tabular}{c | c c c c} 
  \hline\hline
- Parameter & HSC covariance, full & HSC covariance, HOD & Marginalized covariance, HOD \\ [0.5ex] 
+ Parameter & HSC cov. [68\%, 95\%] & CV constraints [68\%, 95\%] & Marg. $N(z)$ [68\%, 95\%] \\ [0.5ex] 
  \hline
- $\chi^2/\nu$ & 87.49/80 & 94.97/88 & 91.06/88 \\
-
- $\mu_{{\rm min},p}$ & -0.52 & -1.47 & -1.42 \\
-
- $\mu_{\rm min}$ & 11.88 & 11.73 & 11.73 \\
-
- $\mu_{0,p}$ & 1.68 & 4.86 & 5.03 \\
-
- $\mu_{0}$ & 0.02 & 11.53 & 11.41 \\
-
- $\mu_{1,p}$ & 0.87 & -0.58 & -0.51 \\
-
- $\mu_{1}$ & 13.073 & 12.866 & 12.873 \\ [1ex] 
+ $\chi^2/\nu$ & 87.49/80 & --- & 88.54/88 \\ 
+$\mu_{{\rm min},p}$ & $-0.5^{+1.7}_{-2.0}$, $^{+4.0}_{-4.0}$ & $-0.4^{+1.1}_{-1.4}$, $^{+2.5}_{-2.3}$ & $-0.3^{+1.5}_{-1.8}$, $^{+3.2}_{-2.8}$ \\ [1ex]
+$\mu_{\rm min}$ & $11.90^{+0.26}_{-0.15}$, $^{+0.39}_{-0.46}$ & $11.86^{+0.19}_{-0.16}$, $^{+0.28}_{-0.30}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.32}_{-0.36}$ \\ [1ex]
+$\mu_{0,p}$ & $2.4\pm 4.3$, $^{+7.2}_{-7.0}$ & $2.1^{+4.2}_{-5.1}$, $^{+6.6}_{-7.2}$ & $2.7^{+6.7}_{-4.4}$, $^{+6.9}_{-7.1}$ \\ [1ex]
+$\mu_{0}$ & $5.7\pm 3.3$, $^{+5.5}_{-5.5}$ & $6.1^{+4.6}_{-3.8}$, $^{+5.2}_{-4.8}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ \\ [1ex]
+$\mu_{1,p}$ & $0.9^{+2.0}_{-2.8}$, $^{+5.1}_{-4.7}$ & $0.9^{+1.3}_{-1.8}$, $^{+3.4}_{-2.9}$ & $1.0^{+1.6}_{-2.6}$, $^{+4.1}_{-3.5}$ \\ [1ex]
+$\mu_{1}$ & $13.10^{+0.30}_{-0.21}$, $^{+0.48}_{-0.54}$ & $13.04\pm 0.21$, $^{+0.38}_{-0.40}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ \\ [1ex]
  \hline
@@
 \end{center}
-%[-1.42408101 11.73454764  5.0302023  11.41301207 -0.50980494 12.87323354]
-%[-1.47048791 11.73097558  4.85774675 11.52813388 -0.58053445 12.86613972]
-%[ 2.18173486e-02 -1.39414027e-02 -4.61928378e-03  1.27879857e-02
-% -5.18739083e-01  1.18844059e+01  1.67767553e+00  2.01822301e-02
-%  8.65219859e-01  1.30726966e+01 -1.64307267e-01 -3.19342441e-02
-%  3.97310971e-02  3.88321407e-02]
+\label{tab:chi2_tests}
+\caption{Table showing the minimum
+$\chi^2$ value as well as the 68\% and 95\% constraints on the 6 HOD parameters
+in three regimes: using the original
+HSC covariance matrix with 14 parameters (6 HOD + 8 $N(z)$ ones),
+importance-sampling the original
+chain to get cosmic variance constraints on the parameters, and finally using the
+marginalized $N(z)$ covariance matrix with the 6 HOD parameters developed in this work.}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c | c c c c} 
+ \hline\hline
+ Parameter & Fiducial [68\%, 95\%] & Test 1 [68\%, 95\%] & Test 2 [68\%, 95\%] & Test 3 [68\%, 95\%] \\ [0.5ex] 
+ \hline
+ $\chi^2/\nu$ & 88.54/88 & 87.67/88 & 88.56/88 & 78.61/88 \\ 
+$\mu_{{\rm min},p}$ & $-0.3^{+1.5}_{-1.8}$, $^{+3.2}_{-2.8}$ & $0.0\pm 1.7$, $^{+3.2}_{-3.3}$ & $-0.3^{+1.4}_{-1.8}$, $^{+3.2}_{-2.8}$ & $-0.8^{+1.8}_{-2.1}$, $^{+3.7}_{-3.5}$ \\ [1ex]
+$\mu_{\rm min}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.32}_{-0.36}$ & $11.85^{+0.22}_{-0.13}$, $^{+0.31}_{-0.37}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.33}_{-0.36}$ & $11.76^{+0.26}_{-0.17}$, $^{+0.38}_{-0.43}$ \\ [1ex]
+$\mu_{0,p}$ & $2.7^{+6.7}_{-4.4}$, $^{+6.9}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ \\ [1ex]
+$\mu_{0}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.8^{+3.6}_{-5.2}$, $^{+5.6}_{-5.5}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.9\pm 3.4$, $^{+5.5}_{-5.6}$ \\ [1ex]
+$\mu_{1,p}$ & $1.0^{+1.6}_{-2.6}$, $^{+4.1}_{-3.5}$ & $1.5^{+2.1}_{-2.6}$, $^{+4.3}_{-3.9}$ & $0.98^{+1.5}_{-2.5}$, $^{+4.2}_{-3.5}$ & $0.6^{+1.9}_{-2.9}$, $^{+4.6}_{-4.0}$ \\ [1ex]
+$\mu_{1}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ & $13.03^{+0.26}_{-0.20}$, $^{+0.41}_{-0.45}$ & $13.00^{+0.25}_{-0.22}$, $^{+0.42}_{-0.44}$ & $12.93^{+0.29}_{-0.24}$, $^{+0.47}_{-0.50}$ \\ [1ex]
+ \hline
+ \hline
+\end{tabular}
+\end{center}
+\label{tab:chi2_tests}
+\caption{Table}
+\end{table}
 

### 2020-07-17 Paper/paper.tex
Updates from Overleaf

@@
         We assume that the cosmic covariance between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by treating each bin independently.
-We show the resulting correlation matrix of the COSMOS cosmic variance in Fig.
-\ref{fig:CV}.
+We show the resulting correlation matrix of the COSMOS cosmic variance in Fig.~\ref{fig:CV}.
 
@@
 Finally, following Ref.~\cite{2019PASJ...71...43H}, photometric redshift distributions have been estimated by cross-matching HSC galaxies to galaxies in the COSMOS 30-band photometric catalog presented in Ref.~\cite{2016ApJS..224...24L}.
-\subsection{Validating the linear expansion}
+
+\subsection{Validating the linear expansion}\label{ssec:hsc.lin}
+In this section, we study how the linear
+approximation performed via the derivative 
+matrix ${\sf T}$ (defined in Eq. \ref{eq:Tmat})
+compares with the exact calculation. The
+derivative matrix can be seen in Fig.~\ref{fig:Tmat}
+for the first tomographic
+bin and its cross-correlations with
+all four bins in the HSC data, 
+i.e. the tomographic pairs [1,1], [1,2],
+[1,3], and [1,4]. The rows of the matrix correspond
+to the $\ell$ multipoles available for each pair, 
+while the columns to the 300 equally
+spaced redshift samples of $N(z)$
+between $z = 0$ and $z = 4$. We can notice that
+as we move to pairs with bins at higher redshifts,
+the peak of the matrix appears at different locations:
+namely, the high-$\ell$ high-$z$ regions of the matrix.
+In the $z$ direction, these coincide with the maxima
+of the photometric redshifts of the HSC data, shown in
+Fig.~\ref{fig:Nzs}. Furthermore, we illustrate the
+effect of the smoothing prior (see 
+Section~\ref{sssec:theory.prior.smooth})
+on the $N(z)$ distributions as well as a perturbed $N(z)$
+set, obtained by adding the noise as
+a Gaussian random draw from the new marginalized $N(z)$ covariance matrix to the smoothed $N(z)$'s.
+The corresponding angular power spectra, $C_\ell$,
+for the three cases are shown in Fig.~\ref{fig:Cls}
+in their respective colors. Furthermore, 
+Fig.~\ref{fig:Cls} illustrates the comparison
+between the exact $C_\ell$'s (\textit{yellow line}) with 
+the linear prediction using the derivative matrix, ${\sf T}$,
+(\textit{dotted light-blue line}). We find that the
+two are in reasonably good agreement provided the current 
+dataset degree of precision. See Appendix B \todo{to be done} 
+for a short explanation of
+why the auto-correlation cannot be well-approximated
+using linear expansion. The $\chi^2$ values for all four
+curves with respect to the HSC data are shown in Table~
+\ref{tab:lin}. Comparing the $\chi^2$ 
+values for the original and smoothed $N(z)$ distributions,
+we notice that they decrease for both the
+HSC covariance matrix and also for the marginalized one.
+As expected, perturbing the photometric distributions
+gets penalized, but the change in the $\chi^2$ value
+is much larger for the original HSC covariance, which
+indicates that the marginalized covariance is more lenient in
+allowing small (and expected) $N(z)$ fluctuations.
+The difference between the last two lines suggests 
+that the Taylor expansion works sufficiently well,
+as the difference in $\chi^2$ is negligible. Finally,
+comparing the $\chi^2$ for the new
+and old covariances, we are led to conclude that the
+new covariance provides a better fit to the data.
+We note that we have fixed the values of
+all the shift and width parameter to be zero and set the
+6 HOD parameters to their \textbf{fiducial}
+values of:
+\begin{itemize}
+    \item $\mu_{\rm min} = 11.88$, $\mu_{\rm min,p} = -0.5$,
+    \item $\mu_{\rm 0} = 5.7$, $\mu_{\rm 0,p} = 2.5$,
+    \item $\mu_{\rm 1} = 13.08$, $\mu_{\rm 1,p} = 0.9$.
+\end{itemize}
+These HOD values
+are selected to match the 6 posterior means reported in Ref.
+\cite{1912.08209}. For the smoothed and perturbed
+$N(z)$ values only, we minimize the likelihood to find the best
+fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$,
+obtaining 13.05 and 0.79, respectively. The ratio of 
+the original HSC precision matrix to the marginalized $N(z)$ 
+one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total
+degrees of freedom of the system (i.e. size of the $C_\ell$
+vector for all cross correlations between the tomographic bins).
 
@@
 \includegraphics[width=1.\textwidth]{./Tmat}
-\caption{Visualization of the derivative matrix $T_{mn} = {\partial t_m}/{\partial N_n}$. \todo{Make look nicer: split into tomo bins}} 
+\caption{Visualization of the linear derivative matrix ${\sf T}
+= {\partial \vt}/{\partial \vN}$ divided by the angular power
+spectrum $C_{\ell}$ for all cross-correlation pairs between
+the first tomographic bin and all four others.} 
 \label{fig:Tmat}
@@
 default values used in the 
-full HSC analysis (CITE Andrina),
+full HSC analysis \cite{1912.08209},
 while the \textit{blue} curves 
@@
 Gaussian draw from the smoothing 
-prior described in Section TODO: Section}
+prior described in Section~\ref{sssec:theory.prior.smooth}.}
 \label{fig:Nzs}
@@
 \caption{Galaxy power spectrum, $C_\ell^{gg}$, between the four
-tomographic bins. The \textit{black} curves show the HSC-measured
+tomographic bins. The \textit{black} curves illustrate the HSC-measured
 power spectrum and error bars, while the
-\textit{red} curves present the $C_\ell$s
-computed using the posterior mean values of the 6 HOD parameters from the
-full HSC analysis (CITE Andrina). In 
-\textit{green} is shown the power
+\textit{red} curves show the $C_\ell$s
+computed using the posterior mean values
+of the 6 HOD parameters from the
+full HSC analysis \cite{1912.08209}. In 
+\textit{green}, we show the angular power
 spectrum obtained by using the smoothed
-$N(z)$ distributions (see Fig. \ref{fig:Nzs}). The \textit{blue}
+$N(z)$ distributions (see Fig.~\ref{fig:Nzs}). The \textit{blue}
 curves correspond to the exact
@@
 perturbed $N(z)$ values in
-Fig. \ref{fig:Nzs}, and the 
+Fig.~\ref{fig:Nzs}, and the 
 \textit{lightblue} dotted lines
@@
 \end{center}
+\caption{$\chi^2$ values with respect to the HSC data of the 
+four curves shown in Fig.~\ref{fig:Cls} using the original 
+covariance used in the HSC analysis and the marginalized
+$N(z)$ covariance developed in this work.}
 \label{tab:lin}
-\caption{$\chi^2$ values with respect to the HSC data of the four curves shown in Fig. \ref{fig:Cls}. $[\mu_1, \ \mu_{1,p}] =  [13.05, 0.79]$
-Ratio of determinants of original precision to CV+noise+smooth precision per dof =  1.15
-94}
 \end{table}
 
-Fig. \ref{fig:Nzs} shows the distribution of the
-photometric redshifts of the HSC data
+\subsection{Impact of N(z) uncertainties on final parameters}
+In this section, we summarize the main results of this
+paper, showing in particular the effect of marginalization
+of the photometric uncertainties on the resulting constraints
+of cosmological parameters of interest. In our particular
+case, these are the 6 HOD parameters described in the model
+in Section~\ref{ssec:hsc.theory}. The marginalized $N(z)$ 
+covariance is computed
+assuming a fiducial set of values for the HOD parameters
+listed in the previous section.
 
-Fig. \ref{fig:Cls} displays the exact $C_\ell$ vs. 
-linear prediction using the derivative matrix, ${\sf T}$.
-
-In Fig. \ref{fig:Tmat}, we show the
-derivative matrix, ${\sf T}$,
-defined in Eq. \ref{eq:Tmat} for the first tomographic
-bin in the HSC data and its cross-correlations with
-all four bins, i.e. . The rows of the matrix correspond
-to the $\ell$ multipoles, while the columns to the
-redshift samples of $N(z)^1$.
-
-\subsection{Impact of N(z) uncertainties on final parameters}
 To study the effect on the covariance matrix
 of the marginalization procedure described in detail
-in sections TODO: CITE, we visualize the original
+in Section~\ref{ssec:theory.nz_new}, we visualize the original
 correlation matrix and its difference with the
-marginalized matrix in Fig. \ref{fig:fid_marg_cov}. The
-main effects on the original matrix is
+marginalized matrix in Fig.~\ref{fig:fid_marg_cov}. The
+main difference with the original matrix is
 a positive contribution to the off-diagonal entries
@@
 parameters of our model to the following values: 
-$A_{\rm smooth} = 1$, $A_{\rm noise} = 4.$,
-$N_{\rm upsample} = 3$, and $\Delta z_{\rm thresh} = 
-0.06$. These values were chosen so as to reduce the
+\begin{itemize}
+    \item $A_{\rm smooth} = 1$, 
+    \item $A_{\rm noise} = 4.$,
+    \item $N_{\rm upsample} = 3$, 
+    \item $\Delta z_{\rm thresh} = 0.06$.
+\end{itemize}
+These values were chosen so as to reduce the
 $\chi^2$ of the perturbed model 
@@
 original covariance matrix. The modest value of the
-smoothing parameter, $A_{\rm smooth}$, ensure that we 
+smoothing parameter, $A_{\rm smooth}$, ensures that we 
 are not oversmoothing the $N(z)$ distribution, while
@@
 proposed in this work provides better constraints
-on the
-6 HOD parameters than the original 14-parameter
-model, we employ the ensemble sampler for Markov
-chain Monte Carlo (MCMC) $\textsc{emcee}$ CITE Goodman 
-\& Weare (2010) and other paper with 6 parameters and
-a walkers' ratio of 6 for 12000 iterations. We assess
-the convergence of the chains with a simple G-R test
-having computed the auto-correlation and thinned the
-chains beforehand. In all cases, we find that the
-chains are reasonably well converged.
+on the 6 HOD parameters than the original 14-parameter
+model, we employ the MCMC ensemble sampler
+$\textbf{emcee}$ \cite{2013PASP..125..306F}
+setting the
+ walkers' ratio to 6 and
+ the number of iterations to 12000. We test
+the convergence of the chains using a simple Gelman-Rubin convergence diagnostic
+after analyzing the auto-correlation statistics and thinning the
+chains appropriately. In all cases, we find that the
+chains are reasonably well converged,
+which gives more credibility to the 
+conclusions we draw.
 
-In Fig. \ref{fig:fid_marg_cov}, we show the key 
-comparison between
-
-We can
-see that the full analysis leads to 
-worse constraints on the parameters
+In Fig.~\ref{fig:fid_marg_cov} and Table~\ref{tab:chi2_fid},
+we illustrate the key  comparison between the original HSC
+covariance matrix and the marginalized
+one developed in this paper. One can
+notice that the full analysis (\textit{in gray}) leads to 
+suboptimal constraints on the parameters
 when compared with the marginalized
-covariance. Furthermore, the proximity
+covariance (\textit{in blue}) in almost all cases
+except for those HOD parameters which are poorly
+constrained by the HSC dataset. 
+Furthermore, the proximity
 of the \textit{blue} and \textit{red} 
@@
 of the 6 parameters (i.e. not exceeding
-the cosmic variance limits) as well as 
-nearly optimal such
+the cosmic variance limits) while at the same time 
+nearly optimal such (as 
+our inferences are always limited
+by cosmic variance).
 
-We, furthermore, run the following several 
-tests to check the robustness of our result:
+To check the robustness of our result, we furthermore
+run the following several tests:
 \begin{itemize}
-\item We study the effect of the smoothing
+\item[\textbf{Test 1}:] We study the effect of the smoothing
 by turning it off, $A_{\rm smooth} = 0$. 
-\item We vary the choice for the
+\item[\textbf{Test 2}:] We vary the choice for the
 diagonal photometric noise parameter
-by increasing it to $A_{\rm noise} = 42$.
-\item We compute the ${\sf T}$ matrix 
+by increasing it ten-fold, $A_{\rm noise} = 42$.
+\item[\textbf{Test 3}:] We compute the ${\sf T}$ matrix 
 using a different set of 
 HOD parameters perturbed by $1\sigma$
-away from the fiducial ones.
+relative to the \textbf{fiducial} ones (listed in Section \ref{ssec:hsc.lin}).
 \end{itemize}
-We show these comparisons with the
+We show the comparisons between
+these three tests and the
 fiducial marginalized $N(z)$ model in Fig.
-\ref{fig:triangle_marg_tests}.
+\ref{fig:triangle_marg_tests} and Table~\ref{tab:chi2_tests}.
 One can notice that the effect of the
-smoothing on the constraints is minimal,
+smoothing on the constraints is negligible,
 while increasing the diagonal noise
-leads to weakening them. Neither of these
+weakens them. Neither of these
 tests changes significantly the
@@
 
-\ref{fig:triangle_fid_marg}
 \begin{figure}[ht]
@@
 \includegraphics[width=1.\textwidth]{./triangle_marg_tests}
-\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green}, we show the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from \todo{CITE Andrina's paper}; in \textit{red}, we display the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally the \textit{blue} contours show the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 1$.}
+\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green}, we show the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from Ref.~\cite{1912.08209}; in \textit{red}, we display the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally the \textit{blue} contours show the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 1$.}
 \label{fig:triangle_marg_tests}
@@
 \end{center}
-\label{tab:chi2_tests}
-\caption{Table showing the minimum
-$\chi^2$ value as well as the 68\% and 95\% constraints on the 6 HOD parameters
-in three regimes: using the original
-HSC covariance matrix with 14 parameters (6 HOD + 8 $N(z)$ ones),
+\caption{Minimum $\chi^2$ values (divided by the degrees
+of freedom) reached in each of the three MCMC runs
+and [68\%, 95\%] constraints on the 6 HOD parameters
+in the following scenarios: 1) adopting the original
+HSC covariance matrix with 14 
+parameters (6 HOD + 8 $N(z)$ ones), 2)
 importance-sampling the original
-chain to get cosmic variance constraints on the parameters, and finally using the
-marginalized $N(z)$ covariance matrix with the 6 HOD parameters developed in this work.}
+chain to get cosmic variance constraints on 
+the parameters, and 3) using the
+marginalized $N(z)$ covariance matrix
+developed in this work to constrain
+the 6 HOD parameters.}
+\label{tab:chi2_fid}
 \end{table}
@@
 \end{center}
+\caption{Minimum $\chi^2$ values (divided by the degrees
+of freedom) attained in each of the four MCMC runs
+and [68\%, 95\%] constraints on the 6 HOD parameters. The
+``fiducial'' column displays the results for the marginalized
+$N(z)$ covariance matrix adopting the fiducial parameter
+choices described in the text, while the three tests
+refer to the marginalized $N(z)$ covariance matrix
+obtained with no smoothing, large diagonal noise and
+$1\sigma$-offset HOD parameters, respectively.}
 \label{tab:chi2_tests}
-\caption{Table}
 \end{table}

### 2020-07-17 Paper/paper.tex
AS fixes

@@
 
-    it has been long recognized that the accuracy of photometric surveys relies on the accuracy of photometric redshift distributions and that these are likely to continue to be a dominant source of systematic uncertainty.
+    It has been long recognized that the accuracy of photometric surveys relies on the accuracy of photometric redshift distributions and that these are likely to continue to be a dominant source of systematic uncertainty.
 The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number density of galaxies as a function of redshift for each tomographic sample.
 It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}.
-Improving photometric redshift techniques and calibrations is an area of active research \cite{2004.09542}\as{add citations}, but for the foreseable future we will need to rely on marginalization to control residual systematic errors.
+Improving photometric redshift techniques and calibrations is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseable future we will need to rely on marginalization to control residual systematic errors.
 
@@
 
-    We applyt this method to the HSC data as an example and  discuss some of the advantages, limitations and possible extensions of this model.
+    We apply this method to the HSC data as an example and  discuss some of the advantages, limitations and possible extensions of this model.
 
@@
       The traditional approach (e.g.
-\todo{Add cites}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector.
+\cite{2018PhRvD..98d3526A,1912.08209}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector.
 This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
@@
 However, this methods suffers from two shortcomings.
-First, it is fundamentally ad-hoc: this particular parametric form is not directly mapped onto the main modes of uncertainty in photometric redshift estimators \todo{we may want an expert to vet this statement}, and there is rigorous proof that the parametrization is complete in the sense that it captures all relevant effects.
+First, it is fundamentally ad-hoc: this particular parametric form, while intuitively the most important, is not directly mapped onto the main modes of uncertainty in photometric redshift estimators, and there is no rigorous proof that the parametrization is complete in the sense that it captures all relevant effects.
 The second problem is that it is computationally expensive.
 By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly.
-For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted 50-60\% of all model parameters.
+For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted for approximately half of all model parameters.
 The methodology described in the next section aims to address both of these shortcomings.
@@
       For simplicity here we will just treat $N_i(z)$ as a histogram, and therefore in our case the $\phi_\alpha$ are just top-hat functions centered on an equi-spaced grid of redshifts $z_\alpha$.
-The formalism described below is valid for any choice of basis functions, however.
-In this formalism, full set of redshift distributions is described by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
-In what follows, we will assume that $\vN$ has units of ``counts''.
+The formalism described below is however valid for any choice of basis functions.
+In particular, it is likely that future photometric estimators will provide their uncertainty in terms of a several important components and a corresponding covariance matrix.
+We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
+In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?}.
 The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
-\todo{ref}).
+\ref{eq:cell_gg_limber}).
 
@@
       Our data is therefore made up of the combination ${\bf d}=(\hat{\bf c},\hat{\bf N})$, where $\hat{\bf c}$ is a set of two-point function measurements.
-The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redsihft distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters.
+The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redshift distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters.
 The posterior distribution is therefore given by:
@@
 \ref{eq:posterior_full} we have assumed that $\hat{\bf c}$ and $\hat{\vN}$ are independent at the likelihood level, and that the likelihood of $\hat{\vN}$ is independent of ${\bf q}$.
-The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \todo{cite} or otherwise).
+The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \cite{2008ApJ...684...88N} or otherwise).
 The method outlined here is still applicable if these assumptions are dropped, albeit with a different expression for the modified covariance in Eq.
 \ref{eq:main2}.
-We have also assumed that the redshift distribution amplitudes $\hat{\vN}$ are Gaussianly distributed.
-\todo{add some blab here and potentially cite Carles Sanchez's latest paper.}
+We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed.
+While this is likely not true in detail for the current $N(z)$, it will likely become a better approximation in the future as our data become more precise.
+Moreover, it is also unlikely to  matter much in practice, as long as $C_N$ captures the relevant directions of uncertainty.
 
-      To derive the main result of this paper, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
+      To proceed, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
       \begin{equation}
@@
 It is the non-linear dependence of $\vt$ on $\vN$ that prevents us from performing what looks like a trivial integral analytically.
-In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters \todo{Maybe mention HMC methods that could actually do this}.
-Instead, we pull an Eastern-European trick and Taylor expand theory around $\bar{\vN}$ \todo{we could change this sentence or leave it as is.
-But I think Taylor was British.}:
+In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters.
+This might actually be possible in the future using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}.
+Instead, faced with queue waiting times at NERSC we Taylor expand theory around $\bar{\vN}$:
       \begin{equation}
@@
       \end{equation}
-      This is a fascinatingly simple equation and the main result of this paper.
+      This  fascinatingly simple equation and the main result of this paper.
 This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$.
@@
 The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether.
-This is the main result of the paper.
 
@@
         \end{figure}
-        The analysis presented in Section \todo{add ref} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \todo{add ref}, as was done in \todo{add cite}.
+        The analysis presented in Section ref{sec:hsc} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \cite{2016ApJS..224...24L}, as was done in \cite{1912.08209}.
 The main sources of uncertainty for this measurement are cosmic variance in the COSMOS field, shot noise and additional uncertainty in the photometric redshifts used to assign galaxies to different bins.
@@
 
-        The contribution from the photometric redshift systematic uncertainties is modelled as \ldots\todo{Not sure how you guys implemented this.}
-
-        The cosmic variance uncertainties are caused by the fluctuations in the matter density traced by galaxies in the particular sky patch.
+        The sample variance uncertainties are caused by the fluctuations in the matter density traced by galaxies in the particular sky patch.
 The corresponding covariance would ideally be estimated from simulations including both non-linear gravitational clustering and a realistic model of the galaxy-halo connection for the specific galaxy sample under study.
 For simplicity in this proof-of-concept analysis we instead use an analytical model for the $N(z)$ covariance.
-Although less precise than a simulation-based calculation, this approach was shown by \todo{cite sanchez} to provide a reasonable prediction for the redsthift distribution uncertainties.
+Although less precise than a simulation-based calculation, this approach was shown by \cite{2004.09542} to provide a reasonable prediction for the redsthift distribution uncertainties.
 
@@
         \end{equation}
-        where $\chi_\alpha$ is the radial comoving distance to redshift $z_\alpha$, and we model the galaxy power spectrum using the Kaiser formula \todo{cite} accounting for redshift-space distortions:
+        where $\chi_\alpha$ is the radial comoving distance to redshift $z_\alpha$, and we model the galaxy power spectrum using the Kaiser formula \cite{1987MNRAS.227....1K} accounting for redshift-space distortions:
         \begin{equation}
@@
 Following the results of \todo{cite}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor.
-$P_mm(k)$ is the non-linear matter power spectrum, which we model via the \todo{cite} HALOFIT parametrization.
+$P_mm(k)$ is the non-linear matter power spectrum, which we model via the HALOFIT parametrization \cite{2003MNRAS.341.1311S}.
 
@@
         \end{equation}
-        We find this contribution to be subdominant in all cases in comparison with cosmic variance.
+We find this contribution to be subdominant in all cases in comparison with sample variance.
+
+To estimate the systematic error associated with the consider alternative phometric redshift codes.
+Following \cite{1912.08209}, we consider  {\tt DEmP}, {\tt Ephor}, {\tt Ephor\_AB} and {\tt FRANKEN-Z}.
+These are some of the best-performing algorithms presented in \cite{2018PASJ...70S...9T}, and constitute a fair representation of the underlying photo-$z$ uncertainties.
+Note however that all of these codes rely on COSMOS-30 for calibration.
+We calculate the variance between $N(z)$s implied for each photometric redshift code.
+We multiply this variance by a factor of 4 and smooth by convolving with a Gaussian with $\sigma_z=0.1$.
+The resulting vector is added to the diagonal of covariance matrix for the pure sample variance contribution.
+While the choices of these parameter is somewhat arbitrary, we think our selection is a conservative while maintaining that variance between codes is representative of systematic uncertainty.
 

### 2020-07-17 create_sacc.py

@@ -34,7 +34,7 @@ lmax = [2170.58958919, 2515.39193451, 3185.36076391, 4017.39370804]
 # choice for noise and smoothing
 #noi_fac, dz_thr, pert, A_smooth, upsample, cov_cv = 42., 0.04, 'cov', 1., 3, True
 #noi_fac, dz_thr, pert, A_smooth, upsample, cov_cv = 4., 0.04, 'cov', 1., 3, True
-noi_fac, dz_thr, pert, A_smooth, upsample, cov_cv = 4., 0.06, 'cov', 1., 3, True
+noi_fac, dz_thr, pert, A_smooth, upsample, cov_cv = 4., 0.06, 'cov', 1., int(sys.argv[3]), True
 
 #noi_fac, A_smooth, pert = 42., 0.1, 'cov'
 #noi_fac, A_smooth, pert = 4000., 0.03, 'cov'

### 2020-07-17 Paper/paper.tex
conclusion

@@
 
+In this method we have presented a new method for marginalizing of redshift distribution uncertainties.
+This method relies by modifying the data covariance matrix by projecting the uncertainties in the redshift distribution to the appropriate data space basis.
+
+We have demonstrate this method on the photometric galaxy clustering data from HSC and compared to previous approaches.
+Method performs as expected without having to include many new parameters in the chain.
+In general this method is applicable for any 3x2pt analysis and will likely perform even better, because galaxy clustering is considerably more sensitive to redsfhit distributions than weak lensing (since the later has a much wider kernel).
+
+This method relies on two approximations.
+The first approximation is that first order expansion of theory with respect to the change in redshift distribution is sufficient and second that the matrix of first derivatives can be approximated to be constant over the parameter space of interest.
+We have explicitly shown that the first approximation is good enough.
+With improving data, the contours will shrink and this will become even better approximation.
+The same is true for the second approximation.
+Note however, that the second approximation could be improved by considering a second order expansion, that would allow us to calculate $T$ as a first order expansion from the fiducial model, i.e.
+\begin{equation}
+        {\sf T} ({{\bf q},\bar{\vN}}) = {\sf T}_{\rm fid} + \left. \frac{\partial \sf T}{\partial {\bf q}}\right|_{{\bf q_{\rm fid}}} \left({\bf q} - {\bf q_{\rm fid}}\right).
+        \label{eq:Tmat}
+\end{equation}
+We leave this for the future work. But in short, these approximations are sufficient and can be further improved if necessary.
+
+The most difficult issue is this determination of the covariance matrix $C_N$.
+In this paper we have considered contributions from the sample variance of COSMOS-30 field and an additional diagonal component given by the scatter between diffierent redshift codes and mutliplied by a ``safety'' factor.
+With the development of new methods for  photmetric redshifit, characterization of uncertainties will most certainly become an integral part of the photometric redshift estimation.
+These uncertainties might be characterised either by a covariance matrix (like in this paper) or perhaps with a set of samples drawn from the probability distribution.
+Even if these samples are not necessarily normal distributed to a good approximation, the implied covariance matrix might still be sufficient to propagate these uncertainties.
+Again this is something that we leave for the future work.
+
+Finally, in this paper we went through some pain to implement a smoothness prior following the intuition that the true $N(z)$ must be smooth and hence it does not make sense to needlessly  marginalize over non-physical $N(z)$s.
+As we have shown, the effect is largerly negligible.
+This is because the high frequency oscillations in $N(z)$ cancel out to a good approximation.
+
 \appendix

### 2020-07-18 chain_analysis/CV_triangle_plot.py
latest code just in case I accidentally rm *

@@ -10,6 +10,7 @@ import sys
 # names and choices
 name_fid = "fid_newcov_april28"
 lab_fid = 'Full 14-parameter fit (HSC paper)'#'fiducial, new covmat'
+name_CV = "CV_newcov_july17"
 lab_CV = 'Cosmic variance constraint'
 name_marg = "marg_newcov_june19"
 lab_marg = 'Marginalized N(z)'#, new covmat'
@@ -33,40 +34,36 @@ def get_par_names(name):
     return n_iter, w_rat, n_par, b_iter, par_names, lab_names
 
 
+# load the fiducial dataset
 # walkers ratio, number of params and burn in iterations
 fiducial_outfile = name_fid+"/"+name_fid+".txt"
 n_iter, w_rat, n_par, b_iter, par_names, lab_names = get_par_names(name_fid)
 fiducial_chains = np.loadtxt(dir_chains+fiducial_outfile)
 
-# parameters for CV importance sampling
-shift_sigmas = np.array([0.008,0.006,0.006,0.007])
-width_sigmas = np.array([0.05,0.045,0.04,0.03])
-
-def gaussian(param,sigma):
-    return np.exp(-0.5*param**2/sigma**2)
-
 # removing burn-in
 fiducial_chains = fiducial_chains[w_rat*n_par*b_iter:]
 
-# compute weights
-weights = np.ones(fiducial_chains.shape[0])
-for i in range(4):
-    sig = shift_sigmas[i]
-    weights[:] *= gaussian(fiducial_chains[:,i],sig)
-for i in range(4):
-    sig = width_sigmas[i]
-    weights[:] *= gaussian(fiducial_chains[:,-4+i],sig)
+# load the CV dataset
+CV_outfile = name_CV+"/"+name_CV+".txt"
+n_iter, w_rat, n_par, b_iter, par_names, lab_names = get_par_names(name_CV)
+CV_chains = np.loadtxt(dir_chains+CV_outfile)
+
+# removing burn-in
+CV_chains = CV_chains[w_rat*n_par*b_iter:]
 
 # feed samples to getdist
-fiducial_hsc_reweighted = MCSamples(samples=fiducial_chains,names=par_names,labels=lab_names,name_tag='Fid',weights=weights)
[…]

### 2020-07-18 create_sacc.py

@@ -193,17 +193,6 @@ prec_n = s_n.precision.getPrecisionMatrix()
 cl_theory = get_theory(s_d,s_noi,hod_params,Z_params,HMCorrection)
 cl_theory_s = get_theory(s_m,s_noi,hod_params,Z_params,HMCorrection)
 
-
-
-
-# xar = np.arange(len(cl_theory))
-# errv = np.sqrt(s_d.precision.getCovarianceMatrix().diagonal())
-# #plt.plot(xar,cl_theory,'b-')
-# plt.errorbar(xar,(s_d.mean.vector-cl_theory)/errv)
-
-# plt.show()
-
-
 di = s_d.mean.vector - cl_theory
 chi2o = get_chi2(di,prec_o)
 chi2n = get_chi2(di,prec_n)
@@ -274,10 +263,12 @@ else:
 print ("Sanity: ",np.dot(dNz,np.dot(np.linalg.inv(prior_cov),dNz)))
 
 
-
-    
-
-if True:
+np.save("z_arr_up.npy",zaru)
+np.save("z_arr.npy",zar)
+np.save("Nz.npy",Nz)
+np.save("Nz_s.npy",Nz_s)
+np.save("Nz_p.npy",NzP)
+if False:
     fig, ax = plt.subplots(4,1, facecolor="w",
             gridspec_kw={"hspace": 0.0},
             figsize=(10, 6))
@@ -330,9 +321,12 @@ log_det_prec_CVnoismo = np.linalg.slogdet(prec_CVnoismo)[1]
 ratio_det_CVnoismo = log_det_prec-log_det_prec_CVnoismo
 ratio_det_CVnoismo = np.exp(ratio_det_CVnoismo)
 print("Ratio of determinants of original precision to CV+noise+smooth precision per dof = ", ratio_det_CVnoismo**(1./dof))
-# for the test data, the answer used to be ~1.12 and is now ~1.3 with COADD
 
-if True:
+np.save("cl_theory.npy",cl_theory)
+np.save("cl_theory_s.npy",cl_theory_s)
+np.save("cl_theory_perturbed.npy",cl_theory_perturbed)
+np.save("cl_theory_taylor.npy",cl_theory_taylor)
+if False:
     fig, ax = plt.subplots(4,4, facecolor="w",
             gridspec_kw={"hspace": 0.0, "wspace":0},
             figsize=(10, 10))
@@ -358,11 +352,3 @@ if True:
[…]

### 2020-07-18 plot_cl.py

@@ -1,0 +1,96 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sacc
+
+dir_read = "data/NEWCOV_COADDED"
+s_d = sacc.SACC.loadFromHDF(dir_read+"/power_spectra_wdpj.sacc")
+s_noi = sacc.SACC.loadFromHDF(dir_read+"/noi_bias.sacc")
+shot_noise = s_noi.mean.vector
+cov = s_d.precision.getCovarianceMatrix()
+upsample = 3
+
+import plotparams
+plotparams.buba()
+
+import distinct_colours
+cs = distinct_colours.get_distinct(4)
+
+zaru = np.load("data_figs/z_arr_up.npy")
+zar = np.load("data_figs/z_arr.npy")
+Nz = np.load("data_figs/Nz.npy")
+Nz_s = np.load("data_figs/Nz_s.npy")
+NzP = np.load("data_figs/Nz_p.npy")
+
+fig, ax = plt.subplots(4,1, facecolor="w",
+          gridspec_kw={"hspace": 0.0},
+          figsize=(14, 10))
+for i in range(4):
+    li = i*100
+    hi = li+100
+    ax[i].plot(zar,Nz[li:hi],cs[0])
+    ax[i].plot(zaru,Nz_s[li*upsample:hi*upsample],cs[1])
+    ax[i].plot(zaru,NzP[li*upsample:hi*upsample],cs[2])
+    ax[i].set_xlim(0,2.6)
+    ax[i].set_ylabel(r'$N_%i (z)$'%(i+1))
+
+    if i < 3:
+        ax[i].xaxis.set_ticklabels([])
+ax[3].set_xlabel(r'$z$')
+ax[0].plot([],[],cs[0],label='original')
+ax[0].plot([],[],cs[1],label='smoothed')
+ax[0].plot([],[],cs[2],label='perturbed')
+ax[0].legend(ncol=3)
+plt.savefig('Paper/Nzs.pdf')
+
+plt.close()
+
+
+cl_theory = np.load("data_figs/cl_theory.npy")-shot_noise
+cl_theory_s = np.load("data_figs/cl_theory_s.npy")-shot_noise
+cl_theory_perturbed = np.load("data_figs/cl_theory_perturbed.npy")-shot_noise
+cl_theory_taylor = np.load("data_figs/cl_theory_taylor.npy")-shot_noise
+
+fig, ax = plt.subplots(4,4, facecolor="w",
+        gridspec_kw={"hspace": 0.0, "wspace":0},
[…]

### 2020-07-18 plottools/plotparams.py

@@ -1,0 +1,154 @@
+from pylab import *
+import matplotlib.pyplot as plt
+
+def dark():
+
+    params = {'axes.labelsize': 30,
+          'axes.titlesize': 22,
+          'font.size': 25,
+          'legend.fontsize': 25,
+          'xtick.labelsize': 40,
+          'ytick.labelsize': 40,
+          'text.usetex': True,
+          'figure.subplot.left'    : 0.14,
+          'figure.subplot.right'   : 0.94  ,
+          'figure.subplot.bottom'  : 0.14  ,
+          'figure.subplot.top'     : 0.96  ,
+          'figure.subplot.wspace'  : 0.  ,
+          'figure.subplot.hspace'  : 0.  ,
+          #'lines.markersize' : 6,
+          #'lines.markeredgewidth'  : 50,
+          'lines.linewidth' : 2.5,
+          'text.latex.unicode': True,
+          'xtick.direction': 'in',
+          'ytick.direction': 'in',
+          'xtick.minor.size': 4,
+          'xtick.minor.width': 2,
+          'ytick.minor.size': 4,
+          'ytick.minor.width': 2,
+          'xtick.major.size': 7.5,
+          'xtick.major.width': 2,
+          'ytick.major.size': 7.5,
+          'ytick.major.width': 2,
+          'axes.facecolor'   : 'black',
+          'grid.color'       : 'white',
+          'xtick.color'      : 'white',
+          'ytick.color'      : 'white',
+          'figure.facecolor' : 'black',
+          'figure.edgecolor' : 'white',
+          'axes.edgecolor'   : 'white',
+          'axes.labelcolor'  : 'white'    
+          }
+
+    rc('font', **{'family': 'sans-serif', 'sans-serif': ['sans-serif']})
+    #plt.minorticks_on()
+    ion() # Turn on interactive plotting
+    rcParams.update(params)
+
+    return "Now using the dark theme"
+
+def default():
+
+    params = {'axes.labelsize': 25,
+          'axes.titlesize': 22,
+          'axes.linewidth': 1.5,
+          'font.size': 20,
+          'legend.fontsize': 20,
+          'xtick.labelsize': 25,
+          'ytick.labelsize': 25,
+          'xtick.major.pad': 8,
+          'ytick.major.pad': 8,
+          'text.usetex': True,
+          'figure.subplot.left'    : 0.18,
+          'figure.subplot.right'   : 0.94  ,
+          'figure.subplot.bottom'  : 0.15  ,
[…]

### 2020-07-21 chain_analysis/CV_triangle_plot.py
slight plotting changes

@@ -89,7 +89,7 @@ print("\\begin{tabular}{c | c c c c} ")
 print(" \\hline\\hline")
 print(" Parameter & HSC cov. [68\%, 95\%] & CV constraints [68\%, 95\%] & Marg. $N(z)$ [68\%, 95\%] \\\\ [0.5ex] ")
 print(" \\hline")
-print(" $\\chi^2/\\nu$ & 87.49/80 & SUBSTITUTE ME & 88.54/88 \\\\ ")
+print(" $\\chi^2/\\nu$ & 87.49/80 & 88.29/80 & 88.54/88 \\\\ ")
 margs = [fiducial_hsc, CV_hsc, marg_hsc]
 print_bounds(margs)
 print(" \\hline")

### 2020-07-21 plot_cl.py

@@ -8,6 +8,7 @@ s_noi = sacc.SACC.loadFromHDF(dir_read+"/noi_bias.sacc")
 shot_noise = s_noi.mean.vector
 cov = s_d.precision.getCovarianceMatrix()
 upsample = 3
+lw = 2.5
 
 import plotparams
 plotparams.buba()
@@ -27,9 +28,9 @@ fig, ax = plt.subplots(4,1, facecolor="w",
 for i in range(4):
     li = i*100
     hi = li+100
-    ax[i].plot(zar,Nz[li:hi],cs[0])
-    ax[i].plot(zaru,Nz_s[li*upsample:hi*upsample],cs[1])
-    ax[i].plot(zaru,NzP[li*upsample:hi*upsample],cs[2])
+    ax[i].plot(zar,Nz[li:hi],lw=lw,c=cs[0])
+    ax[i].plot(zaru,Nz_s[li*upsample:hi*upsample],ls='--',lw=lw,c=cs[1])
+    ax[i].plot(zaru,NzP[li*upsample:hi*upsample],ls='--',lw=lw,c=cs[2])
     ax[i].set_xlim(0,2.6)
     ax[i].set_ylabel(r'$N_%i (z)$'%(i+1))
 
@@ -41,7 +42,6 @@ ax[0].plot([],[],cs[1],label='smoothed')
 ax[0].plot([],[],cs[2],label='perturbed')
 ax[0].legend(ncol=3)
 plt.savefig('Paper/Nzs.pdf')
-
 plt.close()
 
 

### 2020-07-21 Paper/paper.tex
Updates from Overleaf

@@
 \newcommand{\da}[1]{{\textcolor{red}{[DA: #1]}}}
+\newcommand{\bh}[1]{{\textcolor{teal}{[BH: #1]}}}
 
@@
 
-\abstract{We present a new method for marginalization over uncertainties in redshift distributions of tomographic bins applicable to current and coming photometric galaxy surveys.
-We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing incompleteness in our knowledge of redshift distributions.
-In principle, this is marginalization over potentially hundreds of new parameters describing potential deviations as a function of redshift and tomographic bin.
-However, by Taylor expanding theory predictions around a fiducial model, we can perform the marginalization analytically, resulting in a modified data covariance matrix.
-We demonstrate  this method by applying to the galaxy clustering measurements from the Year One of Hyper SupremeCam data demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods.
-We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations and show that the only deviations that matter are smooth deviations.
+\abstract{We present a new method to marginalize over uncertainties in redshift distributions, $N(z)$, within tomographic cosmological analyses applicable to current and upcoming photometric galaxy surveys.
+We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing the uncertainty in our knowledge of redshift distributions.
+In principle, this is marginalization over potentially hundreds or thousands of new parameters describing potential deviations as a function of redshift and tomographic bin.
+However, by linearly expanding the theory predictions around a fiducial model, this marginalization can be performed analytically, resulting in a modified data covariance matrix that effectively downweights the modes of the data vector that are more sensitive to redshift distribution variations.
+We showcase this method by applying it to the galaxy clustering measurements from the Hyper Suprime-Cam first data release.
+We demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods, and explore the impact of priors imposing smoothness in the redshift distributions.
+We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations, and show that the only deviations that matter are smooth deviations \da{are we really doing this?}.
 }
@@
   \section{Introduction}\label{sec:intro}\label{sec:intro}
-    Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky.
-These catalogs can be used to constrain cosmological parameters through measurements of clustering and weak lensing as demonstrated by numerous surveys that have completed observations including  Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201}.
+    Photometric galaxy surveys  have the potential to transform our understanding of the Universe by measuring the properties of millions and soon billions of galaxies on the sky.
+These catalogs can be used to constrain cosmological parameters from measurements of galaxy clustering and weak gravitational lensing as demonstrated by numerous surveys, including the Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey (HSC) \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201} \da{We could also cite the actual analysis papers}.
 
-    It has been long recognized that the accuracy of photometric surveys relies on the accuracy of photometric redshift distributions and that these are likely to continue to be a dominant source of systematic uncertainty.
-The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number density of galaxies as a function of redshift for each tomographic sample.
+    It has been long recognized that the constraining power of photometric surveys relies heavily on the accuracy of photometric redshift distributions and that this is likely to continue to be a dominant source of systematic uncertainty.
+The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number of galaxies as a function of redshift for each tomographic sample.
 It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}.
-Improving photometric redshift techniques and calibrations is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseable future we will need to rely on marginalization to control residual systematic errors.
+Improving photometric redshift techniques and calibration methods for redshift distributions is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseeable future, reliable methods to marginalize over uncertainties in the $N(z)$ well be crucial to control residual systematic errors.
 
-    The traditional approach is to add nuissiance parameters that shift the best guess $N(z)$ or additionally change its width or a relative contribution from a secondary peak in $N(z)$.
-This usually works well, but in additional to adding potentially large number of nuissance parameters suffers from model completeness problem: how do we know that any given parameterization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?
+    The traditional approach has been to add nuisance parameters that shift the mean of the best guess $N(z)$ or, additionally, change its width or a relative contribution from a secondary peak in $N(z)$.
+This approach has so far been sufficient but, in addition to adding a potentially large number of nuisance parameters, the method suffers from a model completeness problem: how do we know that any particular parametrization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?
 
-    In this paper we propose a new technique that performs a marginalization around all possible functional deviations from the best fit $N(z)$.
-The space of possible deviations in $N(z)$ is described as a Gaussian function and under some controlled approximations, the integrals can be performed analytically.
+    In this paper we propose a new technique that performs a marginalization around all possible functional deviations from the best-guess $N(z)$.
+The space of possible deviations in $N(z)$ is described as a Gaussian function and, under some controlled approximations, these deviations can be marginalized over analytically.
 This results in a simple change to the data covariance matrix with a simple explanation: directions in the space of predictions that are degenerate with changes in $N(z)$ are given large variance, which means that information in these direction is not used for inferring the parameters of interest.
-This then marginalizes over $N(z)$ uncertainty in a robust manner that does not add new niussance parameters.
+The method then marginalizes over the $N(z)$ uncertainty in a robust manner, without including a large number of new nuisance parameters.
 
-    We apply this method to the HSC data as an example and  discuss some of the advantages, limitations and possible extensions of this model.
+    We apply this method to the HSC first public data release as an example, and discuss some of the advantages, limitations and possible extensions of this model.
+
+    \da{The paper is structured as follows...}
 
@@
     \subsection{Redshift distribution uncertainties}\label{ssec:theory.nz}
-      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters descring the underlying model.
+      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters describing the underlying model.
 
-      In the case of tomographic analyses of galaxy clustering and large-scale structure, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g.
+      In the case of tomographic large-scale structure analyses, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g.
 galaxy ellipticities or overdensity) and different redshift bins\footnote{We use the term ``redshift bin'' here to denote any given galaxy sample, whether or not it has actually been selected by binning those galaxies into intervals of some redshift estimate.}, and $\vec{\theta}$ includes both cosmological and nuisance parameters needed to produce a forward model of $\vd$ that we will call $\vt(\vec{\theta})$.
-Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian, taking the form
+Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian \cite{2018MNRAS.473.2355S}, taking the form
       \begin{equation}
@@
       We can make significant progress without specifying explicitly how the theory is calculated given the parameters.
-It suffices to say that, given a set of cosmological parameters, which specify the evolution of the universe and the matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g.
-a particular bias or intrinsic alignment model), and the redshift distributions $N(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$.
+It suffices to say that, given a set of cosmological parameters, which specify the expansion history of the universe and the growth of matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g.
+a particular bias or intrinsic alignment model), and the redshift distributions $N_i(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$.
 Our main concern in this paper is the way in which the significant uncertainties in $N(z)$, which can potentially dominate the total systematic error budget, should be parametrized and marginalized over.
@@
       The traditional approach (e.g.
-\cite{2018PhRvD..98d3526A,1912.08209}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector.
+\cite{2018PhRvD..98d3526A,2019PASJ...71...43H,1912.08209,2020A&A...637A.100W}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector.
 This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
@@
 However, this methods suffers from two shortcomings.
-First, it is fundamentally ad-hoc: this particular parametric form, while intuitively the most important, is not directly mapped onto the main modes of uncertainty in photometric redshift estimators, and there is no rigorous proof that the parametrization is complete in the sense that it captures all relevant effects.
+First, it is fundamentally ad-hoc: while this particular parametric form, intuitively characterizes the most important effects, is not directly mapped onto the main modes of uncertainty in photometric redshift estimators, and there is no rigorous proof that the parametrization is complete in the sense that it captures all relevant effects.
 The second problem is that it is computationally expensive.
@@
 We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
-In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?}.
+In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?} \da{This only affects the specific form of the shot-noise covariance Eq.
+\ref{eq:cov_nz_shot}}.
 The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
@@
         &=p({\bf q},\vN|\hat{\bf c},\hat{\bf N})\\\label{eq:posterior_full}
-        &=p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
+        &\propto p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
         &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
@@
 \ref{eq:main2}.
-We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed.
-While this is likely not true in detail for the current $N(z)$, it will likely become a better approximation in the future as our data become more precise.
+We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed \cite{2004.09542}.
+While this is not true in detail for the current $N(z)$, it will likely become a better approximation in the future as data become more precise.
 Moreover, it is also unlikely to  matter much in practice, as long as $C_N$ captures the relevant directions of uncertainty.
 
-      To proceed, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
+      To proceed further, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
       \begin{equation}
@@
 
-%We will describe our uncertainty in $\vN$ as a Gaussian
-%\begin{equation}
-%  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
-%\end{equation}
-%where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations.
-We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal).
-Other possible parameterizations are sub-summed in this approach.
-For example, a general shift-width is allowed with the probability $p(\vN(\Delta z_i, z_{w,i}))$.
-
       We can now write a likelihood for $\hat{\bf c}$ in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
@@
 In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters.
-This might actually be possible in the future using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}.
-Instead, faced with queue waiting times at NERSC we Taylor expand theory around $\bar{\vN}$:
+This should be feasible using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}.
+Instead, faced with queue waiting times at NERSC \da{I'm fine leaving this XD, but the referee or journal will complain} we can Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
       \begin{equation}
@@
       \end{equation}
-      In other words it contains a full information about how all the correlation functions or power spectra respond to a small change in $N(z)$.
+      In other words ${\sf T}$ contains a full information about how all the correlation functions or power spectra respond to a small change in $N_i(z)$.
 
@@
       \end{equation}
-      This  fascinatingly simple equation and the main result of this paper.
-This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$.
-These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.
+      This is a fascinatingly simple equation and the main result of this paper.
+This calculation can be understood as follows: for each deviation around $\bar{\vN}$, there is a corresponding deviation around $\vt$.
+These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N_i(z)$.
 Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go.
@@
 
-      Both the marginalized covariance ${\sf C}_M$ normalizing prefactor in Eq.
+      Both the marginalized covariance ${\sf C}_M$, and the normalizing prefactor in Eq.
 \ref{eq:main1} depend in principle on ${\bf q}$ through the parameter dependence of ${\sf T}$.
@@
       \begin{itemize}
-        \item Run inference problem once with fixed redshift distribution and a standard Gaussian likelihood for $\hat{\bf c}$ with covariance ${\sf C}_c$ to determine a sensible fiducial model.
+        \item Solve the parameter inference problem once with fixed redshift distribution and a standard Gaussian likelihood for $\hat{\bf c}$ with covariance ${\sf C}_c$ to determine a sensible fiducial model.
         \item Calculate ${\sf T}$ and ${\sf C}_M$ at this fiducial model.
-        \item Re-run inference with a standard Gaussian likelihood using the modified covariance matrix ${\sf C}_M$.
+        \item Solve again the inference problem with a standard Gaussian likelihood using the modified covariance matrix ${\sf C}_M$ instead of ${\sf C}_c$.
 This will lead to broadened contours due to the marginalization over redshift distribution uncertainties.
@@
         \end{figure}
-        The analysis presented in Section ref{sec:hsc} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \cite{2016ApJS..224...24L}, as was done in \cite{1912.08209}.
+        The analysis presented in Section \ref{sec:hsc} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \cite{2016ApJS..224...24L}, as was done in \cite{1912.08209}.
 The main sources of uncertainty for this measurement are cosmic variance in the COSMOS field, shot noise and additional uncertainty in the photometric redshifts used to assign galaxies to different bins.
 Here we will associate each of this sources to a separate covariance that will be added in quadrature to form the final ${\sf C}_N$.
+\da{Note that there is one fourth source of uncertainty that we do not address here: although the redshifts provided in the COSMOS 30-band catalog have a high accuracy, they are photometric and therefore subject to potential biases.
+This has been identified as an important source of the relative differences between ongoing cosmic shear surveys \citep{2020A&A...638L...1J}.
+Calibrating this bias or incorporating it into the $N(z)$ error model is an important task that requires a detailed study of the COSMOS 30-band sample, and that we leave for future work.}
 
@@
         where $f(z)$ is the logarithmic growth rate and $b_g(z)$ is the linear galaxy bias.
-Following the results of \todo{cite}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor.
-$P_mm(k)$ is the non-linear matter power spectrum, which we model via the HALOFIT parametrization \cite{2003MNRAS.341.1311S}.
+Following the results of \cite{1912.08209}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor.
+$P_{mm}(k)$ is the non-linear matter power spectrum, which we model via the HALOFIT parametrization \cite{2003MNRAS.341.1311S,2012ApJ...761..152T}.
 
@@
         We assume that the cosmic covariance between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by treating each bin independently.
-We show the resulting correlation matrix of the COSMOS cosmic variance in Fig.~\ref{fig:CV}.
+Figure \ref{fig:CV} shows the correlation matrix associated with this cosmic variance contribution in the COSMOS 30-band sample for the first tomographic bin used in Section \ref{ssec:hsc.data}.
 
         The final contribution, from shot noise, is caused by the discrete nature of galaxies as a tracer of the matter fluctuations, and is simply given by
-        \begin{equation}
+        \begin{equation}\label{eq:cov_nz_shot}
           {\sf C}^{(i\,\alpha),(j\,\beta)}_{N,{\rm SN}}=\delta_{\alpha\beta}\delta_{ij}N^\alpha_i.
@@
 
-To estimate the systematic error associated with the consider alternative phometric redshift codes.
-Following \cite{1912.08209}, we consider  {\tt DEmP}, {\tt Ephor}, {\tt Ephor\_AB} and {\tt FRANKEN-Z}.
-These are some of the best-performing algorithms presented in \cite{2018PASJ...70S...9T}, and constitute a fair representation of the underlying photo-$z$ uncertainties.
+        Finally we estimate the systematic error associated with the choice of photo-$z$ code as follows.
+Following \cite{1912.08209}, we consider  {\tt DEmP}, {\tt Ephor}, {\tt Ephor\_AB} and {\tt FRANKEN-Z} as some of the best-performing algorithms presented in \cite{2018PASJ...70S...9T}, and constitute a fair representation of the underlying photo-$z$ uncertainties.
 Note however that all of these codes rely on COSMOS-30 for calibration.
-We calculate the variance between $N(z)$s implied for each photometric redshift code.
-We multiply this variance by a factor of 4 and smooth by convolving with a Gaussian with $\sigma_z=0.1$.
-The resulting vector is added to the diagonal of covariance matrix for the pure sample variance contribution.
-While the choices of these parameter is somewhat arbitrary, we think our selection is a conservative while maintaining that variance between codes is representative of systematic uncertainty.
+We calculate the variance between $N(z)$s estimated from each code by stacking the redshift probability density functions of each source in the HSC data.
+We multiply this variance by a factor of 4 and smooth it by convolving it with a Gaussian kernel with standard deviation $\sigma_z=0.1$.
+The resulting variance vector is added to the diagonal of the pure sample variance covariance matrix.
+While the choices made when estimating this contribution are somewhat arbitrary, we believe our selection is conservative while maintaining that variance between codes is representative of the underlying systematic uncertainties \da{I think the referee will get hung up on this...
+Can we say that we've explore the sensitivity to these choices?}.
 
@@
 
-      One common way to impose smoothness on a function $f(x)$ is to penalize large values of its firs derivative via $p(f)\propto\exp\left[-(f'(x))^2/(2\sigma_1^2)\right]$.
-In the discrete formalism used here, this is equivalent to imposing a Gaussian prior on $\vN$ with zero mean ($\vN_P=0$) and an inverse covariance given by:
+      One common way to impose smoothness on a function $f(x)$ is to penalize large values of its first derivative via a Gaussian prior of the form $p(f)\propto\sum_x\exp\left[-(f'(x))^2/(2\sigma_1^2)\right]$.
+In the discrete formalism used here, using first order finite differences, this is equivalent to imposing a Gaussian prior on $\vN$ with zero mean ($\vN_P=0$) and an inverse covariance given by:
       \begin{equation}\label{eq:prior_1st}
@@
 
-      The prior in Eq. \ref{eq:prior_1st} effectively penalizes large deviations between adjacent elements of $\vN$. This can be generalized to include all element pairs as
+      The prior in Eq. \ref{eq:prior_1st} effectively penalizes large deviations between adjacent elements of $\vN$. This can be generalized to include all possible pairs as
       \begin{equation}\label{eq:smooth1}
@@
 
-      %Determination of the prior matrix $P$ is where the the actual and often partially subjective prior on our knowledge of $N(z)$ enters.
-Note that the standard parameterizations can always be recast in this language at least in the perturbative limits.
-For example, a vector of derivatives $\lambda_i = dN_i/dz$ when added to $\vN$ will produce a shift in the the shape of $N_i(z)$.
-Therefore, adding a contribution $\sigma_{zi}^2 \lambda_i \lambda_i^T$ will allow shifts with variance $\sigma_{zi}$.
-Of course, when shifts become so large that the perturbative expansion breaks, the approach will work only approximately.
-But in this paper we are not interesting in shifts; instead we will try to build $P$ is a more systematic manner.
+  \section{Appplication to galaxy clustering in HSC}\label{sec:hsc}
+    \an{Some citations are currently missing but I will put them in later.}
+    We apply the method outlined in the previous section to the data presented in Ref.~\cite{1912.08209}.
+This work measured the angular galaxy clustering power spectrum from the first data release (PDR1) of the Hyper Suprime-Cam survey.
+In the following, we give a very brief summary of the methodology employed in Ref.~\cite{1912.08209} and refer the reader to the original paper for further details.
 
-  \section{Appplication to HSC clustering data}\label{sec:hsc}
-    \an{Some citations are currently missing but I will put them in later.}
-    We apply the method outlined above to the data obtained in Ref.~\cite{1912.08209}.
-This work measured the angular galaxy clustering power spectrum from HSC DR1 data.
-In the following, we give a very brief summary of the methodology employed in this work and refer the reader to Ref.~\cite{1912.08209} for more details.
     \subsection{Background theory}\label{ssec:hsc.theory}
@@
 
-    Finally, following Ref.~\cite{1912.08209} we estimate the theoretical prediction for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}.
-Details about HOD parameterizations can be found in these references, and we only provide a succinct description relevant to the present analysis here
+      Following Ref.~\cite{1912.08209} we estimate the theoretical prediction for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution (HOD) modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}.
+Details about HOD parametrizations can be found in these references, and we only provide a succinct description relevant to the present analysis here
 
@@
 
-    Here, $\mathrm{d}n/\mathrm{d}M$ is the halo mass function for halo mass $M$, $b_h(M)$ is the linear halo bias, $N_c(M)$ and $N_s(M)$ are the mean number of central and satellite galaxies in halos of mass $M$, $u_s(k)$ is the Fourier transform of the satellite density profile, and $P_{\rm lin}(k)$ is the linear matter power spectrum.
+      Here, $\mathrm{d}n/\mathrm{d}M$ is the halo mass function for halo mass $M$, $b_h(M)$ is the linear halo bias, $\bar{N}_c(M)$ and $\bar{N}_s(M)$ are the mean number of central and satellite galaxies in halos of mass $M$, $u_s(k)$ is the Fourier transform of the satellite density profile, and $P_{\rm lin}(k)$ is the linear matter power spectrum.
 The number density of galaxies is calculated as
@@
 
-\subsection{The HSC dataset}
+    \subsection{The HSC dataset}\label{ssec:hsc.data}
       The Hyper-Suprime Cam survey is an on-going photometric galaxy survey survey focused mainly on weak gravitational lensing.
-The analysis in Ref.~\cite{1912.08209} is based on the publicly-available HSC DR1 data, whose so-called wide fields cover approximately 108 square degrees on the sky, subdivided into seven distinct patches.
+The analysis in Ref.~\cite{1912.08209} is based on the publicly-available HSC DR1 data \cite{2018PASJ...70S...8A}, whose so-called wide fields cover approximately 108 square degrees on the sky, subdivided into seven distinct patches.
 Ref.~\cite{1912.08209} used these data to compute spherical harmonic galaxy clustering power spectra for four tomographic redshift bins between $z=0.15$ and $z=1.5$, taking both auto- and cross-correlations into account.
@@
     \subsection{Validating the linear expansion}\label{ssec:hsc.lin}
-In this section, we study how the linear
-approximation performed via the derivative 
-matrix ${\sf T}$ (defined in Eq. \ref{eq:Tmat})
-compares with the exact calculation. The
-derivative matrix can be seen in Fig.~\ref{fig:Tmat}
-for the first tomographic
-bin and its cross-correlations with
-all four bins in the HSC data, 
-i.e. the tomographic pairs [1,1], [1,2],
-[1,3], and [1,4]. The rows of the matrix correspond
-to the $\ell$ multipoles available for each pair, 
-while the columns to the 300 equally
-spaced redshift samples of $N(z)$
-between $z = 0$ and $z = 4$. We can notice that
-as we move to pairs with bins at higher redshifts,
-the peak of the matrix appears at different locations:
-namely, the high-$\ell$ high-$z$ regions of the matrix.
-In the $z$ direction, these coincide with the maxima
-of the photometric redshifts of the HSC data, shown in
-Fig.~\ref{fig:Nzs}. Furthermore, we illustrate the
-effect of the smoothing prior (see 
-Section~\ref{sssec:theory.prior.smooth})
-on the $N(z)$ distributions as well as a perturbed $N(z)$
-set, obtained by adding the noise as
-a Gaussian random draw from the new marginalized $N(z)$ covariance matrix to the smoothed $N(z)$'s.
-The corresponding angular power spectra, $C_\ell$,
-for the three cases are shown in Fig.~\ref{fig:Cls}
-in their respective colors. Furthermore, 
-Fig.~\ref{fig:Cls} illustrates the comparison
-between the exact $C_\ell$'s (\textit{yellow line}) with 
-the linear prediction using the derivative matrix, ${\sf T}$,
-(\textit{dotted light-blue line}). We find that the
-two are in reasonably good agreement provided the current 
-dataset degree of precision. See Appendix B \todo{to be done} 
-for a short explanation of
-why the auto-correlation cannot be well-approximated
-using linear expansion. The $\chi^2$ values for all four
-curves with respect to the HSC data are shown in Table~
-\ref{tab:lin}. Comparing the $\chi^2$ 
-values for the original and smoothed $N(z)$ distributions,
-we notice that they decrease for both the
-HSC covariance matrix and also for the marginalized one.
-As expected, perturbing the photometric distributions
-gets penalized, but the change in the $\chi^2$ value
-is much larger for the original HSC covariance, which
-indicates that the marginalized covariance is more lenient in
-allowing small (and expected) $N(z)$ fluctuations.
-The difference between the last two lines suggests 
-that the Taylor expansion works sufficiently well,
-as the difference in $\chi^2$ is negligible. Finally,
-comparing the $\chi^2$ for the new
-and old covariances, we are led to conclude that the
-new covariance provides a better fit to the data.
-We note that we have fixed the values of
-all the shift and width parameter to be zero and set the
-6 HOD parameters to their \textbf{fiducial}
-values of:
+      \begin{figure}[ht]
+        \centering
+        \includegraphics[width=1.\textwidth]{./Tmat}
+        \caption{Visualization of the linear derivative matrix ${\sf T} = {\partial \vt}/{\partial \vN}$ divided by the angular power spectrum $C_{\ell}$ for all cross-correlation pairs between the first tomographic bin and all four others.} \label{fig:Tmat}
+      \end{figure}
+
+      \begin{figure}[ht]
+        \centering
+        \includegraphics[width=1.\textwidth]{./Nzs}
+        \caption{Photometric redshift distributions $N(z)$ of the four tomographic HSC bins.
+The \textit{red} curves show the default values used in the full HSC analysis \cite{1912.08209}, while the \textit{blue} curves show their smoothed and upsampled version used in this work.
+The \textit{green} show a random Gaussian draw from the smoothing prior described in Section~\ref{sssec:theory.prior.smooth}.}\label{fig:Nzs}
+      \end{figure}
+
+      In this section, we study how the linear approximation performed via the derivative matrix ${\sf T}$ (defined in Eq. \ref{eq:Tmat}) compares with the exact $C_\ell$ calculation.
+
+      The derivative matrix can be seen in Fig.~\ref{fig:Tmat} for the first tomographic bin and its cross-correlations with all four bins in the HSC data, i.e.
+the tomographic pairs [1,1], [1,2], [1,3], and [1,4].
+The rows of the matrix correspond to the $\ell$ multipole bandpowers available for each pair, and the columns to the 300 equally-spaced redshift samples of $N(z)$ between $z = 0$ and $z = 4$.
+We can notice that as we move to pairs with bins at higher redshifts, the peak of the matrix appears at different locations: namely, the high-$\ell$ high-$z$ regions of the matrix \da{I find this kind of confusing.
+If I look at the matrix, what I see is that there is a dominant column  which moves with the mean redshift of each bin, not so much about things moving in the $\ell$ direction in a coherent manner.}.
+In the $z$ direction, these coincide with the maxima of the corresponding redshift distribution.
+
+      These $N(z)$s are shown as blue solid lines in Fig.~\ref{fig:Nzs} \da{We should describe either here or in the previous section what we define as the fiducial $N(z)$s.
+I guess, from the fact that the original and smoothed versions aren't so different, that the original ones are the sum of all photo-$z$ codes.
+Is that the case?
+Should we also show the COSMOS $N(z)$s and corresponding smoothed version?
+I'd be really curious to see what they look like.
+Right now it seems like the smoothing does absolutely nothing.}\bh{Resolved?}.
+The figure also shows the {\sl smoothed} distribution $\bar{N}(z)$s in solid red \da{maybe we should make it dashed so we can actually see the red line}\bh{Did that help?}, defined in Eq.
+\ref{eq:priorcov}.
+Our choice of smoothing prior has only a mild effect on the original redshift distribution, mostly removing sharp features at the edges of adjacent $N(z)$ measurements.
+Also shown in yellow is a realization of the $N(z)$ drawn from a multivariate Gaussian distribution with a mean given by the red line and a covariance given by ${\sl P}$ in Eq.
+\ref{eq:priorcov} \da{Is this true?
+If the smoothing prior has any relevance, then it should only allow "smooth" fluctuations.
+Also, the third bin is kind of suspicious in that all the fluctuations seem to be low.}
+      \bh{It is a Gaussian draw from the inverted smoothing prior multiplied by 0.6, see this code line \url{https://github.com/damonge/WeePeeZee/blob/ce50049c6be90ca40b3a903ba782d8bff4b46c78/create_sacc.py##L246}, and \\ $\mathbf{calculate\_smooth\_s\_and\_prior.py}$ for details on the smoothing prior but it should be same as the equation being referenced}.
+
+      The angular power spectra, $C_\ell$, estimated for the same three sets of $N(z)$s are shown in Fig.~\ref{fig:Cls} using the same color scheme.
+The figure also shows the comparison between the exact prediction for the $N(z)$s (yellow line) and the linear prediction around the smoothed distribution using the derivative matrix, ${\sf T}$, (\textit{dotted light-blue line}).
+We find that the two are in reasonably good agreement given the statistical uncertainties in the data.
+\da{we could highlight that, better data should come with better $N(z)$ calibration samples so this statement is also true in the future.} It is worth noting that the agreement between exact prediction and linear model is noticeably worse for auto-correlations.
+As described in Appendix \ref{app:autos}, this is due to the fact that a single $N(z)$ enters the corresponding auto-correlation quadratically, making higher-order terms in Taylor expansion more relevant than in the case of cross-correlations.
+The $\chi^2$ values for all four curves with respect to the HSC data are shown in Table~\ref{tab:lin}.
+\da{I stopped here today} Comparing the $\chi^2$ values for the original and smoothed $N(z)$ distributions, we notice that they decrease for both the HSC covariance matrix and also for the marginalized one.
+As expected, perturbing the photometric distributions gets penalized, but the change in the $\chi^2$ value is much larger for the original HSC covariance, which indicates that the marginalized covariance is more lenient in allowing small (and expected) $N(z)$ fluctuations.
+The difference between the last two lines suggests that the Taylor expansion works sufficiently well, as the difference in $\chi^2$ is negligible.
+Finally, comparing the $\chi^2$ for the new and old covariances, we are led to conclude that the new covariance provides a better fit to the data.
+We note that we have fixed the values of all the shift and width parameter to be zero and set the 6 HOD parameters to their \textbf{fiducial} values of:
       \begin{itemize}
@@
       \end{itemize}
-These HOD values
-are selected to match the 6 posterior means reported in Ref.
-\cite{1912.08209}. For the smoothed and perturbed
-$N(z)$ values only, we minimize the likelihood to find the best
-fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$,
-obtaining 13.05 and 0.79, respectively. The ratio of 
-the original HSC precision matrix to the marginalized $N(z)$ 
-one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total
-degrees of freedom of the system (i.e. size of the $C_\ell$
-vector for all cross correlations between the tomographic bins).
-
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./Tmat}
-\caption{Visualization of the linear derivative matrix ${\sf T}
-= {\partial \vt}/{\partial \vN}$ divided by the angular power
-spectrum $C_{\ell}$ for all cross-correlation pairs between
-the first tomographic bin and all four others.} 
-\label{fig:Tmat}
-\end{figure}
-
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./Nzs}
-\caption{Photometric redshift
-distributions $N(z)$
-of the four tomographic HSC bins.
-The \textit{red} curves show the
-default values used in the 
-full HSC analysis \cite{1912.08209},
-while the \textit{blue} curves 
-show their smoothed and upsampled 
-version used in this work. The
-\textit{green} show a random 
-Gaussian draw from the smoothing 
-prior described in Section~\ref{sssec:theory.prior.smooth}.}
-\label{fig:Nzs}
-\end{figure}
+      These HOD values are selected to match the 6 posterior means reported in Ref.
+\cite{1912.08209}.
+For the smoothed and perturbed $N(z)$ values only, we minimize the likelihood to find the best fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$, obtaining 13.05 and 0.79, respectively.
+The ratio of  the original HSC precision matrix to the marginalized $N(z)$ one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total degrees of freedom of the system (i.e.
+size of the $C_\ell$ vector for all cross correlations between the tomographic bins).
 
@@
 \includegraphics[width=1.\textwidth]{./Cls}
-\caption{Galaxy power spectrum, $C_\ell^{gg}$, between the four
+\caption{Galaxy power spectrum with subtracted shot noise, $C_\ell^{gg}$, between the four
 tomographic bins. The \textit{black} curves illustrate the HSC-measured
@@
 (\textit{gray contours}). The \textit{red
-contours} are obtained by applying
-importance sampling to the full original 
-chain with weights determined by the
-cosmic variance constraints of the
-shift and width parameters. In \textit{blue},
-we show the constraints on the 6 HOD parameters
-obtained using the covariance
+contours} are obtained by running an MCMC
+chain with the original HSC covariance and
+a Gaussian prior on the 4 shift and 4
+width parameters with a zero mean and an RMS
+width for each parameter determined by its 
+statistical distribution over a number different 
+realizations of the cosmic variance. In 
+\textit{blue}, we show the constraints on the 6 
+HOD parameters obtained using the covariance
 matrix marginalized to account for the
@@
  \hline
+ $\chi^2/\nu$ & 87.49/80 & 88.29/80 & 88.54/88 \\ 
+$\mu_{{\rm min},p}$ & $-0.5^{+1.7}_{-2.0}$, $^{+4.0}_{-4.0}$ & $-0.44^{+0.96}_{-1.6}$, $^{+2.9}_{-2.4}$ & $-0.3^{+1.5}_{-1.8}$, $^{+3.2}_{-2.8}$ \\ [1ex]
+$\mu_{\rm min}$ & $11.90^{+0.26}_{-0.15}$, $^{+0.39}_{-0.46}$ & $11.83^{+0.19}_{-0.15}$, $^{+0.31}_{-0.31}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.32}_{-0.35}$ \\ [1ex]
+$\mu_{0,p}$ & $2.4\pm 4.3$, $^{+7.2}_{-7.0}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ & $2.7^{+6.7}_{-4.4}$, $^{+6.9}_{-7.1}$ \\ [1ex]
+$\mu_{0}$ & $5.7\pm 3.3$, $^{+5.5}_{-5.5}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.6}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ \\ [1ex]
+$\mu_{1,p}$ & $0.9^{+2.0}_{-2.8}$, $^{+5.1}_{-4.7}$ & $0.8^{+1.0}_{-2.2}$, $^{+4.0}_{-3.0}$ & $1.0^{+1.6}_{-2.6}$, $^{+4.1}_{-3.5}$ \\ [1ex]
+$\mu_{1}$ & $13.10^{+0.30}_{-0.21}$, $^{+0.48}_{-0.54}$ & $12.99\pm 0.24$, $^{+0.41}_{-0.39}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ \\ [1ex]
+ \hline
+ \hline
+\end{tabular}
+\end{center}
+\label{tab:chi2_tests}
+\caption{Table}
+\end{table}
+
+\iffalse
+\begin{table}
+\begin{center}
+\begin{tabular}{c | c c c c} 
+ \hline\hline
+ Parameter & HSC cov. [68\%, 95\%] & CV constraints [68\%, 95\%] & Marg. $N(z)$ [68\%, 95\%] \\ [0.5ex] 
+ \hline
  $\chi^2/\nu$ & 87.49/80 & --- & 88.54/88 \\ 
@@
 parameters (6 HOD + 8 $N(z)$ ones), 2)
-importance-sampling the original
-chain to get cosmic variance constraints on 
-the parameters, and 3) using the
-marginalized $N(z)$ covariance matrix
-developed in this work to constrain
-the 6 HOD parameters.}
+adopting the original HSC covariance matrix 
+with Gaussian priors on the $N(z)$ parameters
+determined by their statistical distributions
+due to cosmic variance, and 3) using the 
+marginalized $N(z)$ covariance matrix developed 
+in this work to constrain the 6 HOD parameters.}
 \label{tab:chi2_fid}
 \end{table}
+\fi
 

### 2020-07-22 Paper/paper.tex

@@
 
-    We apply this method to the HSC first public data release as an example, and discuss some of the advantages, limitations and possible extensions of this model.
-
-    \da{The paper is structured as follows...}
+    The paper is structured as follows.
+In Section \ref{sec:theory} we derive the basic equations and discuss how the priors on the allowed fluctuations should be determined.
+In Section \ref{sec:hsc} we  apply this method to the HSC first public data release as an example to demonstrate its applicability in a practical settings.
+We conclude in Section \ref{sec:conclusiosn} and discuss some of the advantages, limitations and possible extensions of this model.
 
@@
 We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
-In what follows, we will assume that $\vN$ has units of galaxy ``counts'' \as{Do we actually need to say this?} \da{This only affects the specific form of the shot-noise covariance Eq.
-\ref{eq:cov_nz_shot}}.
-The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
+Normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
 \ref{eq:cell_gg_limber}).
+      %In what follows, we will assume that $\vN$ has units of galaxy ``counts''
+      %\as{Do we actually need to say this?} \da{This only affects the specific form of the shot-noise covariance Eq. \ref{eq:cov_nz_shot}}.
 
@@
 This should be feasible using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}.
-Instead, faced with queue waiting times at NERSC \da{I'm fine leaving this XD, but the referee or journal will complain} we can Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
+Instead, faced with queue waiting times at NERSC we Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
       \begin{equation}
@@
 Here we will associate each of this sources to a separate covariance that will be added in quadrature to form the final ${\sf C}_N$.
-\da{Note that there is one fourth source of uncertainty that we do not address here: although the redshifts provided in the COSMOS 30-band catalog have a high accuracy, they are photometric and therefore subject to potential biases.
+There is fourth source of uncertainty that we do not address here: although the redshifts provided in the COSMOS 30-band catalog have a high accuracy, they are photometric and therefore subject to potential biases.
 This has been identified as an important source of the relative differences between ongoing cosmic shear surveys \citep{2020A&A...638L...1J}.
-Calibrating this bias or incorporating it into the $N(z)$ error model is an important task that requires a detailed study of the COSMOS 30-band sample, and that we leave for future work.}
+Calibrating this bias or incorporating it into the $N(z)$ error model is an important task that requires a detailed study of the COSMOS 30-band sample, and that we leave for future work.
 
@@
 The resulting variance vector is added to the diagonal of the pure sample variance covariance matrix.
-While the choices made when estimating this contribution are somewhat arbitrary, we believe our selection is conservative while maintaining that variance between codes is representative of the underlying systematic uncertainties \da{I think the referee will get hung up on this...
-Can we say that we've explore the sensitivity to these choices?}.
+We discuss sensitivity to some of these partially subjective choices below.
 
@@
 Also, the third bin is kind of suspicious in that all the fluctuations seem to be low.}
-      \bh{It is a Gaussian draw from the inverted smoothing prior multiplied by 0.6, see this code line \url{https://github.com/damonge/WeePeeZee/blob/ce50049c6be90ca40b3a903ba782d8bff4b46c78/create_sacc.py##L246}, and \\ $\mathbf{calculate\_smooth\_s\_and\_prior.py}$ for details on the smoothing prior but it should be same as the equation being referenced}.
+
+%      \bh{It is a Gaussian draw from the inverted smoothing prior multiplied by 0.6, see this code %line\url{https://github.com/damonge/WeePeeZee/blob/ce50049c6be90ca40b3a903ba782d8bff4b46c78/create_sac%c.py##L246}}
+%      % , and \\ \texttt{calculate\_smooth\_s\_and\_prior.py} for details on the smoothing prior but %it should be same as the equation being referenced}.
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2020-07-22 calculate_smooth_s_and_prior.py
weirdly jumbled up files by overleag

@@ -1,0 +1,196 @@
+import sacc
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+from compute_CV_cov import compute_covmat_cv
+from smoothness import obtain_generalized_D
+from scipy.ndimage import gaussian_filter
+from scipy.interpolate import interp1d
+import pyccl as ccl
+from scipy import interpolate
+
+def get_mean_cov(s_data,Ntr,Nztr,noi_fac,upsample):
+    ## THIS IMPLEMENTS UPSAMPLING + SECTION 2.2.1
+    # photo-z codes
+    pz_codes = ['nz_demp', 'nz_ephor', 'nz_ephor_ab', 'nz_frankenz']
+
+    # store new tracers and noise covariance
+    myNztr=Nztr*upsample
+    cov_all = np.zeros((myNztr*Ntr,myNztr*Ntr))
+    tr = []
+    for i in range(Ntr):
+        # get nz for all pz codes
+        zs = s_data.tracers[i].z
+        nzs = s_data.tracers[i].Nz
+        nzs /= np.sum(nzs)
+        if (upsample>=1):
+            minz,maxz= zs[0],zs[-1]
+            newzs = zs[0] + np.arange(myNztr) * (maxz-minz)/(Nztr-1)/upsample
+            newnzs = np.zeros(myNztr)
+            w = np.where(newzs<=maxz)
+            newnzs[w] = [interp1d(zs,nzs, kind='cubic')(newzs[w])]
+            nzs = [newnzs]
+        else:
+            newzs = zs
+            nzs = [nzs]
+
+            
+            
+        for pn in pz_codes:
+            n = s_data.tracers[i].extra_cols[pn]
+            n /= np.sum(n)
+            newn = np.zeros(myNztr)
+            newn[w] = interp1d(zs,n, kind='cubic')(newzs[w])
+            nzs.append(newn)
+        nzs = np.array(nzs)
+
+        # get mean and variance
+        #nz_mean = np.mean(nzs, axis=0)
+        # TESTING
+        nz_mean = np.zeros(myNztr)
+        nz_mean[w] = interp1d(zs,s_data.tracers[i].Nz,kind='cubic')(newzs[w])
+        nz_var = np.var(nzs,axis=0)
+        # TODO: is this necessary? 
+        nz_var = gaussian_filter(nz_var, 2.5*upsample)
+        # TODO: is this correct?
+        # used to be
+        #corr = np.eye(Nztr)
[…]

### 2020-07-22 coadd_saccs.py

@@ -1,0 +1,47 @@
+import sacc
+import os
+import sys
+import numpy as np
+
+# Name of directory where sacc files will be saved
+write = sys.argv[1] # I am currently choosing COADDED, NEWCOV_COADDED
+dir_write = "data/"+write
+
+# Want to include the new covariance matrix?
+want_newcovmat = bool(sys.argv[2])
+if want_newcovmat:
+    newcov = np.load("/users/boryanah/repos/WeePeeZee/data/NEW_COVMAT/covmat.npy")
+
+# Loading the saccs
+sacc_names = ['/users/boryanah/HSC_data/HSC/GAMA15H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/GAMA09H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/WIDE12H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/VVDS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/XMMLSS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc']
+saccs = [sacc.SACC.loadFromHDF(fn) for fn in sacc_names]
+print("We are coadding "+str(len(sacc_names))+" files.")
+
+# Loading the noise saccs
+saccs_noise_names = [os.path.join(os.path.split(fn)[0], 'noi_bias.sacc') for fn in sacc_names]
+print('Reading noise saccs {}.'.format(saccs_noise_names))
+saccs_noise = [sacc.SACC.loadFromHDF(fn) for fn in saccs_noise_names]
+
+# Add precision matrix to noise saccs
+for i, s in enumerate(saccs):
+    saccs_noise[i].precision = s.precision
+
+# Coadding the files normalizing by area
+sacc_coadded = sacc.coadd(saccs, mode='area')
+sacc_noise_coadded = sacc.coadd(saccs_noise, mode='area')
+
+# Change the covmat to the new one if requested
[…]

### 2020-07-22 Paper/paper.tex

@@
 The rows of the matrix correspond to the $\ell$ multipole bandpowers available for each pair, and the columns to the 300 equally-spaced redshift samples of $N(z)$ between $z = 0$ and $z = 4$.
-We can notice that as we move to pairs with bins at higher redshifts, the peak of the matrix appears at different locations: namely, the high-$\ell$ high-$z$ regions of the matrix \da{I find this kind of confusing.
-If I look at the matrix, what I see is that there is a dominant column  which moves with the mean redshift of each bin, not so much about things moving in the $\ell$ direction in a coherent manner.}.
-In the $z$ direction, these coincide with the maxima of the corresponding redshift distribution.
+We can notice that as we move to pairs with bins at higher redshifts, the corresponding block of ${\sf T}$ peaks at a redshift column roughly corresponding to the maximum of the deepest distribution of the pair.
 
-      These $N(z)$s are shown as blue solid lines in Fig.~\ref{fig:Nzs} \da{We should describe either here or in the previous section what we define as the fiducial $N(z)$s.
-I guess, from the fact that the original and smoothed versions aren't so different, that the original ones are the sum of all photo-$z$ codes.
-Is that the case?
-Should we also show the COSMOS $N(z)$s and corresponding smoothed version?
-I'd be really curious to see what they look like.
-Right now it seems like the smoothing does absolutely nothing.}\bh{Resolved?}.
-The figure also shows the {\sl smoothed} distribution $\bar{N}(z)$s in solid red \da{maybe we should make it dashed so we can actually see the red line}\bh{Did that help?}, defined in Eq.
+      These $N(z)$s are shown in Fig.~\ref{fig:Nzs}.
+We obtain the fiducial ``measured'' distributions $\hat{\vN}$, shown in solid blue in the Figure, using the COSMOS 30-band catalog re-weighted to account for the different colour space distribution of that sample using a nearest-neighbours approach as described in \da{cite}.
+The figure also shows the {\sl smoothed} distribution $\bar{N}$ in dashed red, defined in Eq.
 \ref{eq:priorcov}.
 Our choice of smoothing prior has only a mild effect on the original redshift distribution, mostly removing sharp features at the edges of adjacent $N(z)$ measurements.
-Also shown in yellow is a realization of the $N(z)$ drawn from a multivariate Gaussian distribution with a mean given by the red line and a covariance given by ${\sl P}$ in Eq.
+Also shown in solid yellow is a realization of $\vN$ drawn from a multivariate Gaussian distribution with a mean given by the red line and a covariance given by ${\sl P}$ in Eq.
 \ref{eq:priorcov} \da{Is this true?
@@
 We find that the two are in reasonably good agreement given the statistical uncertainties in the data.
-\da{we could highlight that, better data should come with better $N(z)$ calibration samples so this statement is also true in the future.} It is worth noting that the agreement between exact prediction and linear model is noticeably worse for auto-correlations.
+This may not be the case for future datasets with higher statistical power, although the expectation is that those data will be accompanied by larger spectroscopic samples \da{cite some of the Euclid efforts?} needed to reduce systematic photo-$z$ uncertainties.
+It is worth noting that the agreement between exact prediction and linear model is noticeably worse for auto-correlations.
 As described in Appendix \ref{app:autos}, this is due to the fact that a single $N(z)$ enters the corresponding auto-correlation quadratically, making higher-order terms in Taylor expansion more relevant than in the case of cross-correlations.
@@
 
+\section{Precision of the linear expansion for auto- and cross-correlations}\label{app:autos}
+\da{Write this}
+
 \bibliography{bibliography}

### 2020-07-22 Paper/calculate_smooth_s_and_prior.py
let's hope that didn't screw up anything

@@ -1,196 +1,0 @@
-import sacc
-import numpy as np
-import matplotlib.pyplot as plt
-import os
-from compute_CV_cov import compute_covmat_cv
-from smoothness import obtain_generalized_D
-from scipy.ndimage import gaussian_filter
-from scipy.interpolate import interp1d
-import pyccl as ccl
-from scipy import interpolate
-
-def get_mean_cov(s_data,Ntr,Nztr,noi_fac,upsample):
-    ## THIS IMPLEMENTS UPSAMPLING + SECTION 2.2.1
-    # photo-z codes
-    pz_codes = ['nz_demp', 'nz_ephor', 'nz_ephor_ab', 'nz_frankenz']
-
-    # store new tracers and noise covariance
-    myNztr=Nztr*upsample
-    cov_all = np.zeros((myNztr*Ntr,myNztr*Ntr))
-    tr = []
-    for i in range(Ntr):
-        # get nz for all pz codes
-        zs = s_data.tracers[i].z
-        nzs = s_data.tracers[i].Nz
-        nzs /= np.sum(nzs)
-        if (upsample>=1):
-            minz,maxz= zs[0],zs[-1]
-            newzs = zs[0] + np.arange(myNztr) * (maxz-minz)/(Nztr-1)/upsample
-            newnzs = np.zeros(myNztr)
-            w = np.where(newzs<=maxz)
-            newnzs[w] = [interp1d(zs,nzs, kind='cubic')(newzs[w])]
-            nzs = [newnzs]
-        else:
-            newzs = zs
-            nzs = [nzs]
-
-            
-            
-        for pn in pz_codes:
-            n = s_data.tracers[i].extra_cols[pn]
-            n /= np.sum(n)
-            newn = np.zeros(myNztr)
-            newn[w] = interp1d(zs,n, kind='cubic')(newzs[w])
-            nzs.append(newn)
-        nzs = np.array(nzs)
-
-        # get mean and variance
-        #nz_mean = np.mean(nzs, axis=0)
-        # TESTING
-        nz_mean = np.zeros(myNztr)
-        nz_mean[w] = interp1d(zs,s_data.tracers[i].Nz,kind='cubic')(newzs[w])
-        nz_var = np.var(nzs,axis=0)
-        # TODO: is this necessary? 
-        nz_var = gaussian_filter(nz_var, 2.5*upsample)
-        # TODO: is this correct?
-        # used to be
-        #corr = np.eye(Nztr)
[…]

### 2020-07-22 Paper/coadd_saccs.py

@@ -1,47 +1,0 @@
-import sacc
-import os
-import sys
-import numpy as np
-
-# Name of directory where sacc files will be saved
-write = sys.argv[1] # I am currently choosing COADDED, NEWCOV_COADDED
-dir_write = "data/"+write
-
-# Want to include the new covariance matrix?
-want_newcovmat = bool(sys.argv[2])
-if want_newcovmat:
-    newcov = np.load("/users/boryanah/repos/WeePeeZee/data/NEW_COVMAT/covmat.npy")
-
-# Loading the saccs
-sacc_names = ['/users/boryanah/HSC_data/HSC/GAMA15H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/GAMA09H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/WIDE12H/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/VVDS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc', '/users/boryanah/HSC_data/HSC/XMMLSS/CovAna_NoiAna_MskSirius_ClFit_Dpj0_DpjBands1_newDepth/power_spectra_wdpj.sacc']
-saccs = [sacc.SACC.loadFromHDF(fn) for fn in sacc_names]
-print("We are coadding "+str(len(sacc_names))+" files.")
-
-# Loading the noise saccs
-saccs_noise_names = [os.path.join(os.path.split(fn)[0], 'noi_bias.sacc') for fn in sacc_names]
-print('Reading noise saccs {}.'.format(saccs_noise_names))
-saccs_noise = [sacc.SACC.loadFromHDF(fn) for fn in saccs_noise_names]
-
-# Add precision matrix to noise saccs
-for i, s in enumerate(saccs):
-    saccs_noise[i].precision = s.precision
-
-# Coadding the files normalizing by area
-sacc_coadded = sacc.coadd(saccs, mode='area')
-sacc_noise_coadded = sacc.coadd(saccs_noise, mode='area')
-
-# Change the covmat to the new one if requested
[…]

### 2020-07-24 Paper/paper.tex
Updates from Overleaf

@@
 \bibliographystyle{JHEP}
-\title{Marginalization of photometric redshift uncertainties}
+\title{Analytic marginalization of $N(z)$ uncertainties in tomographic galaxy surveys}
 
@@
 \author[c]{Andrina Nicola,}
-\author[d]{An\v{z}e Slosar,}
+\author[d]{An\v{z}e Slosar}
 
@@
 We allow for arbitrary deviations from the best guess $N(z)$ governed by a general covariance matrix describing the uncertainty in our knowledge of redshift distributions.
-In principle, this is marginalization over potentially hundreds or thousands of new parameters describing potential deviations as a function of redshift and tomographic bin.
+In principle, this is marginalization over hundreds or thousands of new parameters describing potential deviations as a function of redshift and tomographic bin.
 However, by linearly expanding the theory predictions around a fiducial model, this marginalization can be performed analytically, resulting in a modified data covariance matrix that effectively downweights the modes of the data vector that are more sensitive to redshift distribution variations.
@@
 We demonstrate how we can marginalize over sample-variance of the PZ calibration sample and a large general systematic uncertainty in photometric determination methods, and explore the impact of priors imposing smoothness in the redshift distributions.
-We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations, and show that the only deviations that matter are smooth deviations \da{are we really doing this?}.
+We find that this method gives results that are comparable to the ad-hoc shift and width parameterizations, and discuss the applicability of the method to ``3$\times$2-point'' analyses and future surveys (and show that the only deviations that matter are smooth deviations \da{are we really doing this?}).
 }
@@
 
-  \section{Introduction}\label{sec:intro}\label{sec:intro}
+  \section{Introduction}\label{sec:intro}
     Photometric galaxy surveys  have the potential to transform our understanding of the Universe by measuring the properties of millions and soon billions of galaxies on the sky.
-These catalogs can be used to constrain cosmological parameters from measurements of galaxy clustering and weak gravitational lensing as demonstrated by numerous surveys, including the Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey (HSC) \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201} \da{We could also cite the actual analysis papers}.
+These catalogs can be used to constrain cosmological parameters from measurements of galaxy clustering and weak gravitational lensing as demonstrated by numerous surveys, including the Sloan Digital Sky Survey \cite{astro-ph/0006396,2013MNRAS.432.1544M}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254,2020A&A...633A..69H}, Dark Energy Survey \cite{1601.00329,2018PhRvD..98d3526A}, Hyper Suprime-Cam survey (HSC) \cite{2012SPIE.8446E..0ZM,2019PASJ...71...43H,1912.08209}, and will play a crucial role in the upcoming Vera Rubin Observatory Legacy Survey of Space and Time \cite{0912.0201}.
 
@@
 The crucial quantity is the redshift distribution $N(z)$, which is simply the mean number of galaxies as a function of redshift for each tomographic sample.
-It is known that the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}.
-Improving photometric redshift techniques and calibration methods for redshift distributions is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseeable future, reliable methods to marginalize over uncertainties in the $N(z)$ well be crucial to control residual systematic errors.
+It is known that the accuracy in the shape of $N(z)$ delivered by the current photometric redshift codes and subsequent calibrations is insufficient for retiring this source of systematic error completely \cite{1809.01669}.
+Improving photometric redshift techniques and calibration methods for redshift distributions is an area of active research \cite{2006A&A...457..841I,2008ApJ...684...88N,2012MNRAS.423..909C,2013MNRAS.431.1547B,2016MNRAS.460.4258L,2019ApJ...877..117H,2018MNRAS.478..592H,2019MNRAS.489..820B,2020MNRAS.491.4768R,2019ApJ...881...80L,2019MNRAS.483.2487J,2019MNRAS.483.2801S,2019arXiv191007127A,2020A&A...637A.100W,2004.09542}, but for the foreseeable future, reliable methods to marginalize over uncertainties in the $N(z)$ will be crucial to control residual systematic errors and fully propagate this uncertainty to final parameter constraints.
 
-    The traditional approach has been to add nuisance parameters that shift the mean of the best guess $N(z)$ or, additionally, change its width or a relative contribution from a secondary peak in $N(z)$.
+    The traditional approach has been to add nuisance parameters that shift the mean of the best guess $N(z)$ or, additionally, change its width or a relative contribution from a secondary peak in $N(z)$ (e.g.
+\cite{2018PhRvD..98d3526A,0912.0201}).
 This approach has so far been sufficient but, in addition to adding a potentially large number of nuisance parameters, the method suffers from a model completeness problem: how do we know that any particular parametrization encompasses all possible ways in which our best-guess $N(z)$ can be wrong?
@@
 In Section \ref{sec:hsc} we  apply this method to the HSC first public data release as an example to demonstrate its applicability in a practical settings.
-We conclude in Section \ref{sec:conclusiosn} and discuss some of the advantages, limitations and possible extensions of this model.
+We conclude in Section \ref{sec:conclusions} and discuss some of the advantages, limitations and possible extensions of this model.
 
@@
       \begin{equation}
-        N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right), \label{eq:photo-z-model}
+        N_{i}(z) \propto \bar{N}_{i}\big(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\big), \label{eq:photo-z-model}
       \end{equation}
-      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width changes.
+      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width variations.
 $\Delta z_i$ and $z_{w,i}$ for each bin are then inferred and marginalized over together with all other parameters, with priors on them based on calibration uncertainties.
@@
 By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly.
-For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted for approximately half of all model parameters.
+For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted for more than half of all model parameters.
 The methodology described in the next section aims to address both of these shortcomings.
@@
 The formalism described below is however valid for any choice of basis functions.
-In particular, it is likely that future photometric estimators will provide their uncertainty in terms of a several important components and a corresponding covariance matrix.
+In particular, it is likely that future photometric estimators will provide their uncertainty in terms of a several important components and a corresponding covariance matrix for their amplitude.
 We describe the full set of redshift distributions by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$.
-Normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
+The normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq.
 \ref{eq:cell_gg_limber}).
-      %In what follows, we will assume that $\vN$ has units of galaxy ``counts''
-      %\as{Do we actually need to say this?} \da{This only affects the specific form of the shot-noise covariance Eq. \ref{eq:cov_nz_shot}}.
 
@@
         &\propto p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
-        &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
+        &=p({\bf q})\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
       \end{align}
@@
       \end{align}
-      and ${\bf t}(q,\vN)$ is the theoretical prediction for $\hat{\bf c}$.
+      and ${\bf t}({\bf q},\vN)$ is the theoretical prediction for $\hat{\bf c}$.
 
@@
 We have also assumed that the deviations from the best guess redshift distribution $\hat{\vN}$ are Gaussianly distributed \cite{2004.09542}.
-While this is not true in detail for the current $N(z)$, it will likely become a better approximation in the future as data become more precise.
-Moreover, it is also unlikely to  matter much in practice, as long as $C_N$ captures the relevant directions of uncertainty.
+While this is not true in detail for the current $N(z)$, it will likely become a better approximation in the future as more spectroscopic data becomes available.
+Moreover, this is also unlikely to  matter much in practice, as long as ${\sf C}_N$ captures the relevant directions and amplitudes of uncertainty.
 
@@
       \begin{equation}
-        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\tilde{\vN}\right)^T{\sf P}^{-1}\left(\vN-\tilde{\vN}\right)+K_N,
+        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\bar{\vN}\right)^T{\sf P}^{-1}\left(\vN-\bar{\vN}\right)+K_N,
       \end{equation}
@@
         p({\bf q},\vN|\hat{\bf c},\hat{\bf N})&=Q\,p({\bf q})\exp\left[-\frac{1}{2}(\hat{\bf c}-{\bf t})^T{\sf C}^{-1}_c(\hat{\bf c}-{\bf t})-\frac{1}{2}(\vN-\bar{\vN})^T{\sf P}^{-1}(\vN-\bar{\vN})\right],\\
-        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
+        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}}.
       \end{align}
@@
       \begin{equation}
-        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T C^{-1} ({\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
+        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} (\hat{\bf c}-\vt)^T C^{-1} (\hat{\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
       \end{equation}
@@
 This should be feasible using methods that can sample very large parameter spaces, such as Hamiltonian Monte Carlo \cite{1987PhLB..195..216D}.
-Instead, faced with queue waiting times at NERSC we Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
+Instead, faced with queue waiting times at NERSC, we Taylor expand the theory in $\vN$ to first order around $\bar{\vN}$:
       \begin{equation}
@@
       \end{equation}
-      In other words ${\sf T}$ contains a full information about how all the correlation functions or power spectra respond to a small change in $N_i(z)$.
+      In other words ${\sf T}$ contains response of all the correlation functions or power spectra respond to small changes in $N_i(z)$.
 
@@
 This calculation can be understood as follows: for each deviation around $\bar{\vN}$, there is a corresponding deviation around $\vt$.
-These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N_i(z)$.
-Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go.
+These are directions in the space of theory predictions that are perfectly degenerate with the changes in shape of $N_i(z)$.
+Equation \ref{eq:main2}  increases the variance for those linear combinations commensurately with how far ${\sf P}$ allows them to go.
 The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether.
@@
 This implies that, in principle, ${\sf T}$ should be re-evaluated at every point in the MCMC chain.
-Since the calculation of ${\sf T}$ is expensive (by comparison with that of e.g.
-$\vt$) using standard methods, we will neglect this dependence and evaluate ${\sf T}$ at a fiducial set of parameters.
+Since the calculation of ${\sf T}$ is expensive using standard methods (by comparison with that of e.g.
+$\vt$), we will neglect this dependence and evaluate ${\sf T}$ at a fiducial set of parameters.
 We will however explore the impact of the choice of fiducial parameters on the final results.
 This should be a sufficiently good approximation for compact likelihoods, where parameter constraints are driven by the differences between $\hat{\bf c}$ and $\vt$.
-We will therefore pre-compute $T$ once, which then means that for a fixed data covariance matrix ${\sf C}_c$, the normalization term in equation \ref{eq:main1} is an irrelevant constant, and the marginalized covariance in the Eq.
+We will therefore pre-compute $T$ once, which then means that, for a fixed data covariance matrix ${\sf C}_c$, the normalization prefactor in equation \ref{eq:main1} is an irrelevant constant, and the marginalized covariance in the Eq.
 \ref{eq:main2} also needs to be evaluated only once.
@@
 Here we will associate each of this sources to a separate covariance that will be added in quadrature to form the final ${\sf C}_N$.
-There is fourth source of uncertainty that we do not address here: although the redshifts provided in the COSMOS 30-band catalog have a high accuracy, they are photometric and therefore subject to potential biases.
+There is a fourth source of uncertainty that we do not address here: although the redshifts provided in the COSMOS 30-band catalog have a high accuracy, they are photometric and therefore subject to potential biases.
 This has been identified as an important source of the relative differences between ongoing cosmic shear surveys \citep{2020A&A...638L...1J}.
@@
 
-        The covariance matrix element between $N^\alpha_i$ and $N^\beta_j$ two tomographic bins is given by:
+        The covariance matrix element between two tomographic bins $N^\alpha_i$ and $N^\beta_j$ is given by:
         \begin{equation}\label{eq:nz_cv}
@@
 Following the results of \cite{1912.08209}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor.
-$P_{mm}(k)$ is the non-linear matter power spectrum, which we model via the HALOFIT parametrization \cite{2003MNRAS.341.1311S,2012ApJ...761..152T}.
+$P_{mm}(k)$ is the real-space non-linear matter power spectrum, which we model via the HALOFIT parametrization \cite{2003MNRAS.341.1311S,2012ApJ...761..152T}.
 
@@
 Note however that all of these codes rely on COSMOS-30 for calibration.
-We calculate the variance between $N(z)$s estimated from each code by stacking the redshift probability density functions of each source in the HSC data.
-We multiply this variance by a factor of 4 and smooth it by convolving it with a Gaussian kernel with standard deviation $\sigma_z=0.1$.
+We calculate the variance between $N(z)$s estimated from each code by stacking the redshift probability density functions (pdfs) of each source in the HSC data.
+We multiply this variance by a factor $A_{\rm noise}$ and smooth it by convolving it with a Gaussian kernel with standard deviation $\sigma_z=0.1$.
 The resulting variance vector is added to the diagonal of the pure sample variance covariance matrix.
-We discuss sensitivity to some of these partially subjective choices below.
+We discuss sensitivity to some of these partially subjective choices in Section \ref{ssec:hcs.mcmc}.
 
@@
 It is therefore a reasonable proposition to impose a certain degree of smoothness on the redshift distributions, which can be achieved through a purposely defined Gaussian prior.
-Applying these types of priors is admittedly a subjective choice to some extent, and we will study its impact on our results in Section \todo{ref}.
+Applying these types of priors is admittedly a subjective choice to some extent, and we will study its impact on our results in Section \ref{ssec:hcs.mcmc}.
 
       One common way to impose smoothness on a function $f(x)$ is to penalize large values of its first derivative via a Gaussian prior of the form $p(f)\propto\sum_x\exp\left[-(f'(x))^2/(2\sigma_1^2)\right]$.
-In the discrete formalism used here, using first order finite differences, this is equivalent to imposing a Gaussian prior on $\vN$ with zero mean ($\vN_P=0$) and an inverse covariance given by:
+In the discrete formalism used here, using first-order finite differences, this is equivalent to imposing a Gaussian prior on $\vN$ with zero mean ($\vN_P=0$) and an inverse covariance given by:
       \begin{equation}\label{eq:prior_1st}
@@
 
-      The prior in Eq. \ref{eq:prior_1st} effectively penalizes large deviations between adjacent elements of $\vN$. This can be generalized to include all possible pairs as
+      The prior in Eq. \ref{eq:prior_1st} effectively penalizes large deviations between adjacent elements of $\vN$. This can be generalized to include all possible pairs of elements as
       \begin{equation}\label{eq:smooth1}
@@
       \begin{equation}
-        p_n=A\,{\rm exp}\left[-\frac{1}{2}\left(n\frac{\Delta z}{\Delta z_{\rm thr}}\right)^2\right],
+        p_n=A_{\rm smooth}\,{\rm exp}\left[-\frac{1}{2}\left(n\frac{\Delta z}{\Delta z_{\rm thr}}\right)^2\right],
       \end{equation}
-      where $\Delta z$ is the redshift separation between neighbouring histogram bins, and $\Delta z_{\rm thr}$ marks the redshift separation beyond which the $N(z)$ is expected to be uncorrelated.
-In our analysis we choose $A$ and $\Delta z_{\rm thr}$ as \todo{describe and motivate this}.
+      where $\Delta z$ is the redshift separation between neighbouring histogram bins, and $\Delta z_{\rm thr}$ marks the redshift separation beyond which different elements of the $N(z)$ are expected to be uncorrelated.
+Our fiducial analysis uses $A_{\rm smooth}=1$ and $\Delta z_{\rm thr}=0.06$.
+These values were empirically chosen to cause a mild smoothing of the redshift distribution directly measured from the COSMOS catalog.
+As we discuss in Section \ref{ssec:hcs.mcmc}, this choice has no practical impact on the final results.
+
+    \subsubsection{Up-sampling}\label{sssec:theory.prior.ups}
+      Depending on the size of the spectroscopic sample used to estimate the initial redshift distribution, the measured $\hat{\vN}$ may be provided with a relatively coarse redshift spacing to reduce the jaggedness in the fiducial $N(z)$.
+However, it may often be desirable to explore the impact of variations in the $N(z)$ on scales smaller than that, which implies artificially increasing the size of $\vN$ by up-sampling the original distributions into a finer grid of $z$.
+
+      This up-sampling can be done in different ways, but in general can be expressed as a linear operation of the form
+      \begin{equation}
+        \vN_{\rm fine}={\sf O}\,\vN_{\rm coarse}.
+      \end{equation}
+      The prior covariance of $\vN_{\rm fine}$ is then related to that of $\vN_{\rm coarse}$ via the bilinear operation
+      \begin{equation}
+        {\sf P}_{\rm fine}={\sf O}\,{\sf P}_{\rm coarse}\,{\sf O}^T.
+      \end{equation}
+
+      For nearest-neigh interpolation, the linear kernel $O_{\mu\alpha}$ is simply $O_{\mu\alpha}=\delta_{\alpha\alpha_\mu}$, where $\alpha_\mu$ is the index of the coarse redshift distribution element that lies closest to the finer grid element with index $\mu$.
+Higher-order interpolation methods can be described in terms of different kernels.
+In practice, the easiest procedure is to simply apply the same interpolating function used to up-sample $\vN$ to all the rows and then all the columns of ${\sf P}_{\rm coarse}$.
+
+      The original $N(z)$s obtained from the COSMOS catalog were measured in bins of $\Delta z=0.04$ in the range $z\in(0,4)$.
+We up-sampled them using linear interpolation by a factor $N_{\rm up}=3$ to a resolution $\Delta z=0.0133$.
+For the total 4 bins used in this analysis, the final up-sampled $\vN$ has 1200 elements.
 
   \section{Appplication to galaxy clustering in HSC}\label{sec:hsc}
-    \an{Some citations are currently missing but I will put them in later.}
+    \an{Some citations are currently missing but I will put them in later.} \da{I added a few, have a look.}
     We apply the method outlined in the previous section to the data presented in Ref.~\cite{1912.08209}.
-This work measured the angular galaxy clustering power spectrum from the first data release (PDR1) of the Hyper Suprime-Cam survey.
+This work measured the angular galaxy clustering power spectrum from the first data release (PDR1) of the Hyper Suprime-Cam survey, described in detail in \cite{2018PASJ...70S...8A,2018PASJ...70S..25M,2018PASJ...70S...5B}.
 In the following, we give a very brief summary of the methodology employed in Ref.~\cite{1912.08209} and refer the reader to the original paper for further details.
@@
       \begin{equation}
-        p^i(z)=\left[\int dz N_i(z)\right]^{-1}N(z)=\frac{\sum_\alpha N^\alpha_i\phi_\alpha(z)}{\sum_\alpha N^\alpha_i\int dz\phi_\alpha(z)}.
+        p^i(z)=\frac{N_i(z)}{\int dz' N_i(z')}=\frac{\sum_\alpha N^\alpha_i\phi_\alpha(z')}{\sum_\alpha N^\alpha_i\int dz\phi_\alpha(z')}.
       \end{equation}
-      The simplicity with which the redshift distribution enters the prediction for the angular power spectrum in Eq.~\ref{eq:cell_gg_limber}, facilitates the computation of the ${\sf T}$ matrix defined above.
+      The simplicity with which the redshift distribution amplitudes $N_i^\alpha$ enter the prediction for the angular power spectrum in Eq.~\ref{eq:cell_gg_limber}, facilitates the computation of the ${\sf T}$ matrix defined in Eq.
+\ref{eq:Tmat}.
 
@@
 These power spectra have been corrected for observational and extragalactic systematics by deprojection at the map-level \cite{2019MNRAS.484.4127A}.
+We also used the same scale cuts described in \cite{1912.08209}.
 Finally, following Ref.~\cite{2019PASJ...71...43H}, photometric redshift distributions have been estimated by cross-matching HSC galaxies to galaxies in the COSMOS 30-band photometric catalog presented in Ref.~\cite{2016ApJS..224...24L}.
@@
         \caption{Photometric redshift distributions $N(z)$ of the four tomographic HSC bins.
-The \textit{red} curves show the default values used in the full HSC analysis \cite{1912.08209}, while the \textit{blue} curves show their smoothed and upsampled version used in this work.
+The \textit{red} curves show the default values used in the full HSC analysis \cite{1912.08209}, while the \textit{blue} curves show their smoothed and up-sampled version used in this work.
 The \textit{green} show a random Gaussian draw from the smoothing prior described in Section~\ref{sssec:theory.prior.smooth}.}\label{fig:Nzs}
+      \end{figure}
+
+      \begin{figure}[ht]
+        \centering
+        \includegraphics[width=1.\textwidth]{./Cls}
+        \caption{Galaxy power spectrum with subtracted shot noise, $C_\ell^{gg}$, between the four tomographic bins.
+The \textit{black} curves illustrate the HSC-measured power spectrum and error bars, while the \textit{red} curves show the $C_\ell$s computed using the posterior mean values of the 6 HOD parameters from the full HSC analysis \cite{1912.08209}.
+In \textit{green}, we show the angular power spectrum obtained by using the smoothed $N(z)$ distributions (see Fig.~\ref{fig:Nzs}).
+The \textit{blue} curves correspond to the exact calculation of the $C_\ell$s using the perturbed $N(z)$ values in Fig.~\ref{fig:Nzs}, and the \textit{lightblue} dotted lines show the power spectra derived via Taylor expansion: $C_\ell' = C_\ell + {\sf T} \left(\vN' - {\vN} \right)$.
+\da{Isn't the fit of the blue line to the data super horrible in the first bin?
+Maybe the NG covariance actually allows this...}}\label{fig:Cls}
       \end{figure}
@@
 As described in Appendix \ref{app:autos}, this is due to the fact that a single $N(z)$ enters the corresponding auto-correlation quadratically, making higher-order terms in Taylor expansion more relevant than in the case of cross-correlations.
+
       The $\chi^2$ values for all four curves with respect to the HSC data are shown in Table~\ref{tab:lin}.
-\da{I stopped here today} Comparing the $\chi^2$ values for the original and smoothed $N(z)$ distributions, we notice that they decrease for both the HSC covariance matrix and also for the marginalized one.
+Comparing the $\chi^2$ values for the original and smoothed $N(z)$ distributions, we notice that they decrease for both the HSC covariance matrix and also for the marginalized one.
 As expected, perturbing the photometric distributions gets penalized, but the change in the $\chi^2$ value is much larger for the original HSC covariance, which indicates that the marginalized covariance is more lenient in allowing small (and expected) $N(z)$ fluctuations.
@@
 Finally, comparing the $\chi^2$ for the new and old covariances, we are led to conclude that the new covariance provides a better fit to the data.
-We note that we have fixed the values of all the shift and width parameter to be zero and set the 6 HOD parameters to their \textbf{fiducial} values of:
-      \begin{itemize}
-        \item $\mu_{\rm min} = 11.88$, $\mu_{\rm min,p} = -0.5$,
-        \item $\mu_{\rm 0} = 5.7$, $\mu_{\rm 0,p} = 2.5$,
-        \item $\mu_{\rm 1} = 13.08$, $\mu_{\rm 1,p} = 0.9$.
-      \end{itemize}
-      These HOD values are selected to match the 6 posterior means reported in Ref.
+We note that we have fixed the values of all the shift and width parameter to be zero and set the 6 HOD parameters to their fiducial values of
+      \begin{equation}\label{eq:par_fid}
+        (\mu_{\rm min},\mu_{{\rm min},p},\mu_0,\mu_{0,p},\mu_1,\mu_{1,p})=(11.88,-0.5,5.7,2.5,13.08,0.9).
+      \end{equation}
+
+      These HOD values correspond to the posterior means reported in Ref.
 \cite{1912.08209}.
-For the smoothed and perturbed $N(z)$ values only, we minimize the likelihood to find the best fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$, obtaining 13.05 and 0.79, respectively.
-The ratio of  the original HSC precision matrix to the marginalized $N(z)$ one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total degrees of freedom of the system (i.e.
+For the smoothed and perturbed $N(z)$ values only, we minimize the likelihood to find the best fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$, obtaining 13.05 and 0.79, respectively \da{It's not clear if you get the same values for both the smoothed and perturbed $N(z)$s (which would be weird) or just for one of them.
+It's also not clear if you're saying that you did this for the results in Table \ref{tab:lin} or just now.}.
+The ratio of  the original HSC precision matrix to the marginalized $N(z)$ one \da{Not clear at all how you calculate this number or why it's relevant} is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total degrees of freedom of the system (i.e.
 size of the $C_\ell$ vector for all cross correlations between the tomographic bins).
-
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./Cls}
-\caption{Galaxy power spectrum with subtracted shot noise, $C_\ell^{gg}$, between the four
-tomographic bins. The \textit{black} curves illustrate the HSC-measured
-power spectrum and error bars, while the
-\textit{red} curves show the $C_\ell$s
-computed using the posterior mean values
-of the 6 HOD parameters from the
-full HSC analysis \cite{1912.08209}. In 
-\textit{green}, we show the angular power
-spectrum obtained by using the smoothed
-$N(z)$ distributions (see Fig.~\ref{fig:Nzs}). The \textit{blue}
-curves correspond to the exact
-calculation of the $C_\ell$s using the
-perturbed $N(z)$ values in
-Fig.~\ref{fig:Nzs}, and the 
-\textit{lightblue} dotted lines
-show the power spectra derived via
-Taylor expansion: $C_\ell' = C_\ell +
- {\sf T} \left(\vN' - {\vN} \right)$.}
-\label{fig:Cls}
-\end{figure}
 
@@
         \begin{center}
- \begin{tabular}{c | c c } 
+          \begin{tabular}{l| c c }
             \hline\hline
@@
             \hline
-  $\chi^2$, original $N(z)$, evaluated at $C_\ell$ & 98.97 & 88.97 \\
-
- $\chi^2$, smoothed $N(z)$, evaluated at $C_\ell$ & 91.33 & 86.44 \\
-
- $\chi^2$, perturbed $N(z)$, evaluated at $C_\ell'$ & 99.78 & 89.57 \\
-
- $\chi^2$, perturbed $N(z)$, taylor-expanded $C_\ell +
- {\sf T} \Delta \vN$ & 100.62 & 89.64 \\
-
+            $\chi^2$, original $N(z)$, exact $C_\ell$ & 98.97 & 88.97 \\ 
+            $\chi^2$, smoothed $N(z)$, exact $C_\ell$ & 91.33 & 86.44 \\
+            $\chi^2$, perturbed $N(z)$, exact $C_\ell$ & 99.78 & 89.57 \\
+            $\chi^2$, perturbed $N(z)$, $C_\ell(\bar{\vN}) + {\sf T} \Delta \vN$ & 100.62 & 89.64 \\
             \hline
@@
         \end{center}
-\caption{$\chi^2$ values with respect to the HSC data of the 
-four curves shown in Fig.~\ref{fig:Cls} using the original 
-covariance used in the HSC analysis and the marginalized
-$N(z)$ covariance developed in this work.}
-\label{tab:lin}
+        \caption{$\chi^2$ values with respect to the HSC data of the four curves shown in Fig.~\ref{fig:Cls} using the original covariance used in the HSC analysis and the marginalized $N(z)$ covariance developed in this work.}\label{tab:lin}
       \end{table}
 
-\subsection{Impact of N(z) uncertainties on final parameters}
-In this section, we summarize the main results of this
-paper, showing in particular the effect of marginalization
-of the photometric uncertainties on the resulting constraints
-of cosmological parameters of interest. In our particular
-case, these are the 6 HOD parameters described in the model
-in Section~\ref{ssec:hsc.theory}. The marginalized $N(z)$ 
-covariance is computed
-assuming a fiducial set of values for the HOD parameters
-listed in the previous section.
-
-To study the effect on the covariance matrix
-of the marginalization procedure described in detail
-in Section~\ref{ssec:theory.nz_new}, we visualize the original
-correlation matrix and its difference with the
-marginalized matrix in Fig.~\ref{fig:fid_marg_cov}. The
-main difference with the original matrix is
-a positive contribution to the off-diagonal entries
-for redshift samples in close proximity. We set the free
-parameters of our model to the following values: 
-\begin{itemize}
-    \item $A_{\rm smooth} = 1$, 
-    \item $A_{\rm noise} = 4.$,
-    \item $N_{\rm upsample} = 3$, 
-    \item $\Delta z_{\rm thresh} = 0.06$.
-\end{itemize}
-These values were chosen so as to reduce the
-$\chi^2$ of the perturbed model 
-using the new covariance matrix when compared with the
-original covariance matrix. The modest value of the
-smoothing parameter, $A_{\rm smooth}$, ensures that we 
-are not oversmoothing the $N(z)$ distribution, while
-the choice of photometric noise as four times 
-larger than the variance in the different photometric
-measures provides a reasonable guess for the expected
-variation in $N(z)$ given the uncertainty.
-
-To assess whether the marginalization procedure 
-proposed in this work provides better constraints
-on the 6 HOD parameters than the original 14-parameter
-model, we employ the MCMC ensemble sampler
-$\textbf{emcee}$ \cite{2013PASP..125..306F}
-setting the
- walkers' ratio to 6 and
- the number of iterations to 12000. We test
-the convergence of the chains using a simple Gelman-Rubin convergence diagnostic
-after analyzing the auto-correlation statistics and thinning the
-chains appropriately. In all cases, we find that the
-chains are reasonably well converged,
-which gives more credibility to the 
-conclusions we draw.
-
-In Fig.~\ref{fig:fid_marg_cov} and Table~\ref{tab:chi2_fid},
-we illustrate the key  comparison between the original HSC
-covariance matrix and the marginalized
-one developed in this paper. One can
-notice that the full analysis (\textit{in gray}) leads to 
-suboptimal constraints on the parameters
-when compared with the marginalized
-covariance (\textit{in blue}) in almost all cases
-except for those HOD parameters which are poorly
-constrained by the HSC dataset. 
-Furthermore, the proximity
-of the \textit{blue} and \textit{red} 
-contours serves both as a sanity check 
-that our method yields physical constraints
-of the 6 parameters (i.e. not exceeding
-the cosmic variance limits) while at the same time 
-nearly optimal such (as 
-our inferences are always limited
-by cosmic variance).
-
-To check the robustness of our result, we furthermore
-run the following several tests:
-\begin{itemize}
-\item[\textbf{Test 1}:] We study the effect of the smoothing
-by turning it off, $A_{\rm smooth} = 0$. 
-\item[\textbf{Test 2}:] We vary the choice for the
-diagonal photometric noise parameter
-by increasing it ten-fold, $A_{\rm noise} = 42$.
-\item[\textbf{Test 3}:] We compute the ${\sf T}$ matrix 
-using a different set of 
-HOD parameters perturbed by $1\sigma$
-relative to the \textbf{fiducial} ones (listed in Section \ref{ssec:hsc.lin}).
-\end{itemize}
-We show the comparisons between
-these three tests and the
-fiducial marginalized $N(z)$ model in Fig.
-\ref{fig:triangle_marg_tests} and Table~\ref{tab:chi2_tests}.
-One can notice that the effect of the
-smoothing on the constraints is negligible,
-while increasing the diagonal noise
-weakens them. Neither of these
-tests changes significantly the
-mean of the posterior distribution.
-On the other hand, offsetting the 
-HOD parameters used in the ${\sf T}$
-matrix computation leads to a shift to
-the mean and a slight widening of the
-contours. That is because the covariance
-matrix used in the analysis no longer
-favors the HOD values preferred by the
-data. This slight tension
-results in sub-par convergence
-of the chains and shifted mean posteriors.
-
-\begin{figure}[ht]
+    \subsection{Impact of N(z) uncertainties on final parameters}\label{ssec:hcs.mcmc}
+      \begin{figure}
         \centering
@@
         \end{subfigure}
-\caption{Correlation matrix of the original HSC dataset (\textit{left} panel) and difference between the
-original and the marginalized (this work)
-correlation matrices (\textit{right} panel). Most of the
-effect of marginalizing for the shift and width
-parameters is manifested as a positive contribution
-to the original matrix near the diagonal (i.e. there
-is an increase in correlation for close redshift
-values). Note that we have subtracted the diagonal
-off the two matrices.}
-\label{fig:fid_marg_cov}
+        \caption{Correlation matrix of the original HSC dataset (\textit{left} panel) and difference between the original and the marginalized (this work) correlation matrices (\textit{right} panel).
+Most of the effect of marginalizing for the shift and width parameters is manifested as a positive contribution to the original matrix near the diagonal (i.e.
+there is an increase in correlation for close redshift values).
+Note that we have subtracted the diagonal off the two matrices.}\label{fig:fid_marg_cov}
       \end{figure}
-
-\begin{figure}[ht]
+      \begin{figure}
         \centering  
         \includegraphics[width=1.\textwidth]{./triangle_fid_marg}
-\caption{Triangle plot with constraints of 
-the 6 HOD parameters obtained using the 
-covariance matrix from the full
-14-parameter original HSC analysis 
-(\textit{gray contours}). The \textit{red
-contours} are obtained by running an MCMC
-chain with the original HSC covariance and
-a Gaussian prior on the 4 shift and 4
-width parameters with a zero mean and an RMS
-width for each parameter determined by its 
-statistical distribution over a number different 
-realizations of the cosmic variance. In 
-\textit{blue}, we show the constraints on the 6 
-HOD parameters obtained using the covariance
-matrix marginalized to account for the
-$N(z)$ uncertainty (i.e. without the
-8 shift and width parameters). We can
-see that the full analysis leads to 
-worse constraints on the parameters
-when compared with the marginalized
-covariance.}
-\label{fig:triangle_fid_marg}
+        \caption{Triangle plot with constraints of the 6 HOD parameters obtained using the  covariance matrix from the full 14-parameter original HSC analysis (\textit{gray contours}).
+The \textit{red contours} are obtained by running an MCMC chain with the original HSC covariance and a Gaussian prior on the 4 shift and 4 width parameters with a zero mean and an RMS width for each parameter determined by its statistical distribution over a number different realizations of the cosmic variance.
+In \textit{blue}, we show the constraints on the 6 HOD parameters obtained using the covariance matrix marginalized to account for the $N(z)$ uncertainty (i.e.
+without the 8 shift and width parameters).
+We can see that the full analysis leads to worse constraints on the parameters when compared with the marginalized covariance.} \label{fig:triangle_fid_marg}
       \end{figure}
+      In this section, we summarize the main results of this paper, showing in particular the effect of marginalization of the photometric uncertainties on the final parameter constraints.
+In our particular case, these are the 6 HOD parameters described in Section~\ref{ssec:hsc.theory}.
+The ${\sf T}$ matrix used to calculate the marginalized $N(z)$ covariance was computed assuming the fiducial HOD parameters listed in Eq.
+\ref{eq:par_fid}.
 
-\begin{figure}[ht]
-\centering  
-\includegraphics[width=1.\textwidth]{./triangle_marg_tests}
-\caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green}, we show the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from Ref.~\cite{1912.08209}; in \textit{red}, we display the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally the \textit{blue} contours show the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 1$.}
-\label{fig:triangle_marg_tests}
-\end{figure}
+      To study the effect on the covariance matrix of the marginalization procedure described in detail in Section~\ref{ssec:theory.nz_new}, we visualize the original correlation matrix and its difference with the marginalized matrix in Fig.~\ref{fig:fid_marg_cov}.
+The main difference with the original matrix is a positive contribution to the off-diagonal entries for redshift samples in close proximity.
 
+      We set the free parameters of our model to the following values \da{these parameters are not defined anywhere.
+I think I know what they are, so I'll try to add it to section 2.}: $(A_{\rm smooth},\,A_{\rm noise},\,N_{\rm up},\,\Delta z_{\rm thr})=(1,\,4,\,3,\,0.06)$.
+These values were chosen so as to reduce the $\chi^2$ of the perturbed model using the marginalized covariance matrix when compared with the original covariance matrix \da{This sentence is imprecise.
+Presumably there's many parameter combinations that would achieve some level of reduction.
+As it is, the sentence reads as if you tweaked the parameters to get the result you want.
+I think I would remove the sentence}.
+The modest value of the smoothing parameter, $A_{\rm smooth}$ \da{So this is still something that confuses me: why can we say that $A_{\rm smooth}=1$ is ``small''?
+not sure what to compare it with, and in any case this value would depend on how the $N(z)$ is normalized.
+Should we say that the value was chosen to visually achieve some level of smoothing without massively changing the original $N(z)$?}, ensures that we are not oversmoothing the $N(z)$ distribution, while the choice of photometric noise as four times larger than the variance in the different photometric redshift codes provides a conservative estimate of the expected variation in $N(z)$ given the uncertainty \da{I feel like this discussion should have happened in Section 2, so I'll move it there.}.
       \begin{table}
@@
             $\mu_{1}$ & $13.10^{+0.30}_{-0.21}$, $^{+0.48}_{-0.54}$ & $12.99\pm 0.24$, $^{+0.41}_{-0.39}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ \\ [1ex]
- \hline
- \hline
+            \hline\hline
           \end{tabular}
         \end{center}
-\label{tab:chi2_tests}
-\caption{Table}
+        \caption{Minimum $\chi^2$ values (divided by the degrees of freedom) reached in each of the three MCMC runs and [68\%, 95\%] constraints on the 6 HOD parameters in the following scenarios: 1) adopting the original HSC covariance matrix with 14 parameters (6 HOD + 8 $N(z)$ ones), 2) adopting the original HSC covariance matrix with Gaussian priors on the $N(z)$ parameters determined by their statistical distributions due to cosmic variance, and 3) using the marginalized $N(z)$ covariance matrix developed in this work to constrain the 6 HOD parameters.}\label{tab:chi2_fid}
       \end{table}
+
+      To compare the results of the automatic marginalization procedure proposed in this work with those of the original shift-width parametrization in terms of the final constraints on the 6 HOD parameters, we explore the posterior distribution of these parameters in both cases using the MCMC ensemble sampler ${\tt emcee}$ \cite{2013PASP..125..306F}.
+To test the convergence of the parameter chains we use a simple Gelman-Rubin convergence diagnostic after analyzing the auto-correlation statistics.
+
+      \da{I stopped here today.}
+      In Fig.~\ref{fig:triangle_fid_marg} and Table~\ref{tab:chi2_fid}, we illustrate the key  comparison between the original HSC analysis, marginalising over 8 shift-width parameters, and the marginalisation scheme developed in this paper.
+One can notice that the full analysis (\textit{in gray}) leads to suboptimal constraints on the parameters when compared with the marginalized covariance (\textit{in blue}) in almost all cases except for those HOD parameters which are poorly constrained by the HSC dataset to begin width.
+Furthermore, the proximity of the \textit{blue} and \textit{red} contours \da{We haven't really explained what the red contours are.
+I'll rephrase} serves both as a sanity check that our method yields physical constraints of the 6 parameters (i.e.
+not exceeding the cosmic variance limits) while at the same time nearly optimal such (as our inferences are always limited by cosmic variance).
+
+      To check the robustness of our result, we furthermore run the following several tests:
+      \begin{itemize}
+        \item[\textbf{Test 1}:] We study the effect of the smoothing prior by turning it off, $A_{\rm smooth} = 0$.
+        \item[\textbf{Test 2}:] We vary the choice for the diagonal photometric noise parameter by increasing it ten-fold, $A_{\rm noise} = 42$ \da{A much better number than 40}.
+        \item[\textbf{Test 3}:] We compute the ${\sf T}$ matrix using a different set of HOD parameters perturbed by $1\sigma$ relative to the fiducial ones (listed in Section \ref{ssec:hsc.lin}) \da{We should say what the new values are and we could even put a dot in the triangle plot to show how far they are from the ML}.
+      \end{itemize}
+      We show the comparisons between these three tests and the fiducial marginalized $N(z)$ model in Fig.
+\ref{fig:triangle_marg_tests} and Table~\ref{tab:chi2_tests}.
+One can notice that the effect of the smoothing on the constraints is negligible, while increasing the diagonal noise weakens them.
+Neither of these tests changes significantly the mean of the posterior distribution.
+On the other hand, offsetting the HOD parameters used in the ${\sf T}$ matrix computation leads to a slight shift and broadening of the contours.
+This is somewhat expected, since the covariance matrix in this case is far from the favoured model parameters.
+This slight tension results in sub-par convergence of the chains and shifted mean posteriors \da{If convergence is really an issue we should have just run them for longer.
+Otherwise, I'd just remove this sentence and end up by saying that the shift/puffing is small given how wrong the new parameters are.}.
+      \begin{figure}[ht]
+        \centering
+        \includegraphics[width=1.\textwidth]{./triangle_marg_tests}
+        \caption{Triangle plot with constraints of the 6 HOD parameters obtained using different choices for the marginalized covariance matrix: in \textit{green}, we show the result where the diagonal noise is set to a very large value ($A_{\rm noise} = 42$); in \textit{gray} is shown the case where the marginalization is performed assuming fiducial HOD-model values which are $1\sigma$ away from their posterior mean values taken from Ref.~\cite{1912.08209}; in \textit{red}, we display the case where the smoothing has been removed ($A_{\rm smooth} = 0$); and finally the \textit{blue} contours show the fiducial case with $A_{\rm noise} = 4$ and $A_{\rm smooth} = 1$.} \label{fig:triangle_marg_tests}
+      \end{figure}
 
@@
 in this work to constrain the 6 HOD parameters.}
-\label{tab:chi2_fid}
+\label{tab:chi2_fidd}
 \end{table}
@@
           $\chi^2/\nu$ & 88.54/88 & 87.67/88 & 88.56/88 & 78.61/88 \\
+          $\mu_{\rm min}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.32}_{-0.36}$ & $11.85^{+0.22}_{-0.13}$, $^{+0.31}_{-0.37}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.33}_{-0.36}$ & $11.76^{+0.26}_{-0.17}$, $^{+0.38}_{-0.43}$ \\ [1ex]
+          $\mu_{0}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.8^{+3.6}_{-5.2}$, $^{+5.6}_{-5.5}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.9\pm 3.4$, $^{+5.5}_{-5.6}$ \\ [1ex]
+          $\mu_{1}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ & $13.03^{+0.26}_{-0.20}$, $^{+0.41}_{-0.45}$ & $13.00^{+0.25}_{-0.22}$, $^{+0.42}_{-0.44}$ & $12.93^{+0.29}_{-0.24}$, $^{+0.47}_{-0.50}$ \\ [1ex]
           $\mu_{{\rm min},p}$ & $-0.3^{+1.5}_{-1.8}$, $^{+3.2}_{-2.8}$ & $0.0\pm 1.7$, $^{+3.2}_{-3.3}$ & $-0.3^{+1.4}_{-1.8}$, $^{+3.2}_{-2.8}$ & $-0.8^{+1.8}_{-2.1}$, $^{+3.7}_{-3.5}$ \\ [1ex]
-$\mu_{\rm min}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.32}_{-0.36}$ & $11.85^{+0.22}_{-0.13}$, $^{+0.31}_{-0.37}$ & $11.83^{+0.21}_{-0.15}$, $^{+0.33}_{-0.36}$ & $11.76^{+0.26}_{-0.17}$, $^{+0.38}_{-0.43}$ \\ [1ex]
           $\mu_{0,p}$ & $2.7^{+6.7}_{-4.4}$, $^{+6.9}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ & $2.5\pm 4.3$, $^{+7.1}_{-7.1}$ \\ [1ex]
-$\mu_{0}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.8^{+3.6}_{-5.2}$, $^{+5.6}_{-5.5}$ & $5.8\pm 3.4$, $^{+5.5}_{-5.5}$ & $5.9\pm 3.4$, $^{+5.5}_{-5.6}$ \\ [1ex]
           $\mu_{1,p}$ & $1.0^{+1.6}_{-2.6}$, $^{+4.1}_{-3.5}$ & $1.5^{+2.1}_{-2.6}$, $^{+4.3}_{-3.9}$ & $0.98^{+1.5}_{-2.5}$, $^{+4.2}_{-3.5}$ & $0.6^{+1.9}_{-2.9}$, $^{+4.6}_{-4.0}$ \\ [1ex]
-$\mu_{1}$ & $13.00^{+0.25}_{-0.21}$, $^{+0.42}_{-0.44}$ & $13.03^{+0.26}_{-0.20}$, $^{+0.41}_{-0.45}$ & $13.00^{+0.25}_{-0.22}$, $^{+0.42}_{-0.44}$ & $12.93^{+0.29}_{-0.24}$, $^{+0.47}_{-0.50}$ \\ [1ex]
           \hline
@@
       \end{center}
-\caption{Minimum $\chi^2$ values (divided by the degrees
-of freedom) attained in each of the four MCMC runs
-and [68\%, 95\%] constraints on the 6 HOD parameters. The
-``fiducial'' column displays the results for the marginalized
-$N(z)$ covariance matrix adopting the fiducial parameter
-choices described in the text, while the three tests
-refer to the marginalized $N(z)$ covariance matrix
-obtained with no smoothing, large diagonal noise and
-$1\sigma$-offset HOD parameters, respectively.}
-\label{tab:chi2_tests}
+      \caption{Minimum $\chi^2$ values (divided by the degrees of freedom) attained in each of the four MCMC runs and [68\%, 95\%] \da{I think quoting 95\% makes this too cluttered.} constraints on the 6 HOD parameters.
+The ``fiducial'' column displays the results for the marginalized $N(z)$ covariance matrix adopting the fiducial parameter choices described in the text, while the three tests refer to the marginalized $N(z)$ covariance matrix obtained with no smoothing, large diagonal noise and $1\sigma$-offset HOD parameters, respectively.}\label{tab:chi2_tests}
     \end{table}
 
-\section{Conclusions}
-\label{sec:conclusions}
+  \section{Conclusions}\label{sec:conclusions}
+    In this paper we have presented a new method to marginalize over uncertainties in the redshift distribution in tomographic large-scale structure analyses.
+This method relies by modifying the data covariance matrix by commensurately increasing the variance of the modes that are most sensitive to these uncertainties.
+This is akin to mode-deprojection methods used in other areas of cosmological data analysis, such as power spectrum estimation or component separation.
 
-In this method we have presented a new method for marginalizing of redshift distribution uncertainties.
-This method relies by modifying the data covariance matrix by projecting the uncertainties in the redshift distribution to the appropriate data space basis.
+    We have demonstrated the performance of this method by applying it to the analysis of photometric galaxy clustering data from HSC, comparing it with previous approaches.
+The method performs as expected, propagating the uncertainties in $N(z)$ without including any additional nuisance parameters.
+In general this method is applicable for any 3$\times$2-point analysis combining galaxy clustering and cosmic shear.
+Since galaxy clustering is considerably more sensitive to redshift distribution uncertainties than cosmic shear (given the radially cumulative nature of the latter), the method will likely perform even better in that case.
 
-We have demonstrate this method on the photometric galaxy clustering data from HSC and compared to previous approaches.
-Method performs as expected without having to include many new parameters in the chain.
-In general this method is applicable for any 3x2pt analysis and will likely perform even better, because galaxy clustering is considerably more sensitive to redsfhit distributions than weak lensing (since the later has a much wider kernel).
+    This method relies on three approximations.
+The first approximation is that a first order expansion of the theory data vector with respect to the change in redshift distribution is sufficient.
+The second one is that the gradient of the data vector with respect to the redshift distribution amplitudes can be approximated to be constant over the parameter space of interest.
+We have explicitly shown that the first approximation is good enough for existing data from the COSMOS 30-band sample.
+With improving data, the $N(z)$ uncertainties will shrink, making this approximation more reliable.
+We have also shown that the second approximation is valid by re-evaluating our parameter constraints for a ${\sf T}$ matrix calculated away from the best-fit parameters.
+Note however, that the main reason to avoid recomputing ${\sf T}$ at every point in parameter space is that its calculation can take significantly longer than the calculation of the theory prediction ${\bf t}$.
+This could however be sped up by incorporating the calculation of ${\sf T}$ as a product of the Limber integrator used to compute ${\bf t}$.
+Alternatively, this approximation could be improved by expanding ${\sf T}$ itself to linear order in the cosmological and astrophysical parameters:
+    \begin{equation}
+      {\sf T} ({{\bf q},\bar{\vN}}) = {\sf T}_{\rm fid} + \left. \frac{\partial \sf T}{\partial {\bf q}}\right|_{{\bf q_{\rm fid}}} \left({\bf q} - {\bf q_{\rm fid}}\right). \label{eq:Tmatd}
+    \end{equation}
+    We leave this study for the future work. In short, these approximations are sufficient and can be further improved if necessary.
 
-This method relies on two approximations.
-The first approximation is that first order expansion of theory with respect to the change in redshift distribution is sufficient and second that the matrix of first derivatives can be approximated to be constant over the parameter space of interest.
-We have explicitly shown that the first approximation is good enough.
-With improving data, the contours will shrink and this will become even better approximation.
-The same is true for the second approximation.
-Note however, that the second approximation could be improved by considering a second order expansion, that would allow us to calculate $T$ as a first order expansion from the fiducial model, i.e.
-\begin{equation}
-        {\sf T} ({{\bf q},\bar{\vN}}) = {\sf T}_{\rm fid} + \left. \frac{\partial \sf T}{\partial {\bf q}}\right|_{{\bf q_{\rm fid}}} \left({\bf q} - {\bf q_{\rm fid}}\right).
-        \label{eq:Tmat}
-\end{equation}
-We leave this for the future work. But in short, these approximations are sufficient and can be further improved if necessary.
+    The third approximation used here is the assumption that the redshift distribution uncertainties follow a multivariate Gaussian distribution.
+This is likely not true in detail for direct calibration methods, where Poisson or Dirichlet distributions may be more appropriate in certain regimes \da{cite}.
+On the other hand, the $N(z)$ uncertainties for clustering redshifts measurements \da{cite} would likely be well described by Gaussian statistics.
+We argue that, regardless of the specifics of the underlying distribution, the Gaussian assumption allows us to quantify the allowed statistical variations in the $N(z)$ (potentially overestimating them) and to easily propagate them into the relevant modes of the theory data vector, leading to reliable parameter constraints marginalized over these uncertainties.
+A detailed validation of this assumption is also left for future work.
+With the development of new photometric redshift methods, the characterization of these distributions will most certainly become an integral part of the photometric redshift and $N(z)$ estimation \da{cite sanchez}.
 
-The most difficult issue is this determination of the covariance matrix $C_N$.
-In this paper we have considered contributions from the sample variance of COSMOS-30 field and an additional diagonal component given by the scatter between diffierent redshift codes and mutliplied by a ``safety'' factor.
-With the development of new methods for  photmetric redshifit, characterization of uncertainties will most certainly become an integral part of the photometric redshift estimation.
-These uncertainties might be characterised either by a covariance matrix (like in this paper) or perhaps with a set of samples drawn from the probability distribution.
-Even if these samples are not necessarily normal distributed to a good approximation, the implied covariance matrix might still be sufficient to propagate these uncertainties.
-Again this is something that we leave for the future work.
+    The central problem in implementing this method is the determination of the covariance matrix of the $N(z)$ uncertainties, ${\sf C}_N$, as well as any physical priors on e.g.
+the smoothness of the underlying distribution.
+In this paper we have considered contributions from the sample variance of COSMOS-30 field, and an additional diagonal component given by the scatter between the redshift distributions estimated from the stacking of individual source pdfs from four different photo-$z$ codes mutliplied by a ``safety'' factor.
+We also implemented ``smoothness'' prior following the intuition that the true $N(z)$ must be smooth and hence it does not make sense to needlessly marginalize over non-physical $N(z)$s.
+As we have shown, for the mild prior used here, the effect is largely negligible.
 
-Finally, in this paper we went through some pain to implement a smoothness prior following the intuition that the true $N(z)$ must be smooth and hence it does not make sense to needlessly  marginalize over non-physical $N(z)$s.
-As we have shown, the effect is largerly negligible.
-This is because the high frequency oscillations in $N(z)$ cancel out to a good approximation.
+    \da{Final inspirational paragraph goes here.}
 
 \appendix
-\section{Oversampling}
-We may want to sub-sample the originally measured $N_{\alpha}$ ($\alpha\in[0,N_z^{\rm coarse})$) into a finer grid $n_{\mu}$ ($\mu\in[0,N_z^{\rm fine})$).
-This is essentially interpolation, which we can write as a linear operator $n_\mu = O_{\mu\alpha}N_{\alpha}$.
-The CV covariance of the $n_\mu$ is then given by ${\rm Cov}^{\rm fine}_{\mu\nu}=O_{\mu\alpha}O_{\nu\beta}{\rm Cov}^{\rm coarse}_{\alpha\beta}$.
-
-For nearest-neighbor interpolation, the kernel $O_{\mu\alpha}$ is simply $O_{\mu\alpha}=\delta_{\alpha\alpha_\mu}$, where $\alpha_\mu$ is the index of the $N_\alpha$ that lies closest to $n_\mu$.
-In practice, the easiest thing to do will be to just use the same linear interpolation function (e.g.
-scipy's {\tt interp1d}) on all the rows of ${\rm Cov}^{\rm coarse}$ and then all the columns of the resulting matrix.
-
 \section{Precision of the linear expansion for auto- and cross-correlations}\label{app:autos}

### 2020-07-26 chain_analysis/CV_triangle_plot.py
latest plot settings

@@ -9,17 +9,21 @@ import sys
 
 # names and choices
 name_fid = "fid_newcov_april28"
-lab_fid = 'Full 14-parameter fit (HSC paper)'#'fiducial, new covmat'
+lab_fid = '14-parameter fit (HSC paper)'#'fiducial, new covmat'
 name_CV = "CV_newcov_july17"
-lab_CV = 'Cosmic variance constraint'
+lab_CV = '14-parameter fit with Cosmic Variance prior'#'Cosmic variance constraint'
 name_marg = "marg_newcov_june19"
-lab_marg = 'Marginalized N(z)'#, new covmat'
+lab_marg = '6-parameter fit using ${\sf C}_M$'#'Marginalized N(z)'
 dir_chains = "/users/boryanah/desclss_chains/"
 ups = ''#"_upsample4"
 
 # what are we plotting
 HOD_pars = ["lmminp","lmmin","m0p","m0","m1p","m1"]
 filename = "triangle_fid_marg"+ups+".pdf"
+
+ranges_dic = {"m1":[12,14],#"lmminp":,
+              "lmmin":[11,12.5]} #"m0p":,#"m0":,#"m1p":[],
+              
 
 def get_par_names(name):
     if 'april29' in name or 'june19' in name:
@@ -52,8 +56,8 @@ CV_chains = np.loadtxt(dir_chains+CV_outfile)
 CV_chains = CV_chains[w_rat*n_par*b_iter:]
 
 # feed samples to getdist
-CV_hsc = MCSamples(samples=CV_chains,names=par_names,labels=lab_names,name_tag='CV')
-fiducial_hsc = MCSamples(samples=fiducial_chains,names=par_names,labels=lab_names,name_tag='Fid')
+CV_hsc = MCSamples(samples=CV_chains,names=par_names,labels=lab_names,ranges=ranges_dic,name_tag='CV')
+fiducial_hsc = MCSamples(samples=fiducial_chains,names=par_names,labels=lab_names,ranges=ranges_dic,name_tag='Fid')
 
 # load the marginalized dataset
 marg_outfile = name_marg+"/"+name_marg+ups+".txt"
@@ -64,7 +68,7 @@ marg_chains = np.loadtxt(dir_chains+marg_outfile)
 marg_chains = marg_chains[w_rat*n_par*b_iter:]
 
[…]

### 2020-07-26 create_sacc.py

@@ -42,7 +42,6 @@ noi_fac, dz_thr, pert, A_smooth, upsample, cov_cv = 4., 0.06, 'cov', 1., int(sys
 #noi_fac, A_smooth, pert = 42., 0.0, 'shiftone'
 #noi_fac, A_smooth, pert = 420., 0.0, 'shiftall'
 
-
 # Cosmological parameters
 # Checked cosmo_params are the same as what the MCMC chain fixes for them
 cosmo_params = {'n_s': 0.9649,
@@ -221,6 +220,7 @@ if want_mini:
     res = minimize(smooth_chi2, x0, method='powell',\
                    options={'xtol': xtol, 'disp': True})
     m1, m1p = res.x
+
     #m1, m1p = 13.041574051253585, 0.7820088517924304
     print("m1, m1p = ",m1,m1p)
     new_hod_params = hod_params.copy()
@@ -233,8 +233,6 @@ chi2o = get_chi2(di,prec_o)
 chi2n = get_chi2(di,prec_n)
 print ("Chi2 original model, smoothed Nz, original cov:", chi2o, dof)
 print ("Chi2 original model, smoothed Nz, new cov ", chi2n)
-
-
 
 
 #Lets perturb dNz, draw from prior_cov
@@ -263,11 +261,11 @@ else:
 print ("Sanity: ",np.dot(dNz,np.dot(np.linalg.inv(prior_cov),dNz)))
 
 
-np.save("z_arr_up.npy",zaru)
-np.save("z_arr.npy",zar)
-np.save("Nz.npy",Nz)
-np.save("Nz_s.npy",Nz_s)
-np.save("Nz_p.npy",NzP)
+np.save("data_figs/z_arr_up.npy",zaru)
+np.save("data_figs/z_arr.npy",zar)
+np.save("data_figs/Nz.npy",Nz)
+np.save("data_figs/Nz_s.npy",Nz_s)
+np.save("data_figs/Nz_p.npy",NzP)
 if False:
     fig, ax = plt.subplots(4,1, facecolor="w",
             gridspec_kw={"hspace": 0.0},
@@ -322,10 +320,10 @@ ratio_det_CVnoismo = log_det_prec-log_det_prec_CVnoismo
 ratio_det_CVnoismo = np.exp(ratio_det_CVnoismo)
 print("Ratio of determinants of original precision to CV+noise+smooth precision per dof = ", ratio_det_CVnoismo**(1./dof))
 
[…]

### 2020-07-26 plot_cl.py

@@ -22,6 +22,7 @@ Nz = np.load("data_figs/Nz.npy")
 Nz_s = np.load("data_figs/Nz_s.npy")
 NzP = np.load("data_figs/Nz_p.npy")
 
+
 fig, ax = plt.subplots(4,1, facecolor="w",
           gridspec_kw={"hspace": 0.0},
           figsize=(14, 10))
@@ -51,7 +52,7 @@ cl_theory_perturbed = np.load("data_figs/cl_theory_perturbed.npy")-shot_noise
 cl_theory_taylor = np.load("data_figs/cl_theory_taylor.npy")-shot_noise
 
 fig, ax = plt.subplots(4,4, facecolor="w",
-        gridspec_kw={"hspace": 0.0, "wspace":0},
+        gridspec_kw={"hspace": 0., "wspace":0.4},#0.
         figsize=(23, 16))
 data = s_d.mean.vector-shot_noise
 err = np.sqrt(cov.diagonal())
@@ -68,13 +69,22 @@ for i in range(4):
             cax.plot(ell,well*cl_theory_s[ndx],c=cs[1])
             cax.plot(ell,well*cl_theory_perturbed[ndx],c=cs[2])
             cax.plot(ell,well*cl_theory_taylor[ndx],c=cs[3],ls=':')
-            cax.errorbar(ell,data[ndx]*well,yerr=err[ndx]*well,fmt='ko-')
+            # HSC data
+            #cax.errorbar(ell,data[ndx]*well,yerr=err[ndx]*well,fmt='ko-')
+            '''
+            # If no space between subplots, i.e. wspace = 0.
             if i > 0:
                 cax.yaxis.set_ticklabels([])
+                cax.set_ylim(ax[j,0].get_ylim()) # lol, makes things empty -- too different
             else:
                 cax.set_ylabel(r'$C_{\ell} \ \ell^{1.2}$')
+            '''
+            # if including wspace
+            if i == 0: cax.set_ylabel(r'$C_{\ell} \ \ell^{1.2}$')
             if j == 3:
                 cax.set_xlabel(r'$\ell$')
+            else:
+                cax.xaxis.set_ticklabels([])
             cax.legend(frameon=False,loc='upper right')
         else:
             cax.yaxis.set_ticklabels([])
[…]

### 2020-07-26 plot_cov.py

@@ -12,7 +12,7 @@ def cov2corr(covmat,nodiag=True):
         corrmat = covmat/np.sqrt(np.diag(covmat)[None,:]*np.diag(covmat)[:,None])
     return corrmat
     
-def plot2d(data,which,logscale=False,lim=None,aspect=1,save=True,levels=None,clip=0,clbar=True,cm=None,label=None,labsize=14,extent=None,ticksize=12,*args):
+def plot2d(data,which,want_axis=True,logscale=False,lim=None,aspect=1,save=True,levels=None,clip=0,clbar=True,cm=None,label=None,labsize=14,extent=None,ticksize=12,*args):
     # origin at upper left
     if label is not None:
         xlabel, ylabel = label
@@ -34,6 +34,9 @@ def plot2d(data,which,logscale=False,lim=None,aspect=1,save=True,levels=None,cli
     plt.xlabel(xlabel)
     plt.ylabel(ylabel)
     plt.gca().set_aspect(aspect)
+    if not want_axis:
+        plt.gca().xaxis.set_ticklabels([])
+        plt.gca().yaxis.set_ticklabels([])
     plt.colorbar(im,fraction=0.1, pad=0.02)
     if save: plt.savefig("Paper/"+which+".pdf"); plt.close()
 
@@ -58,6 +61,10 @@ s_m = sacc.SACC.loadFromHDF("data/"+dir_write+"/power_spectra_wdpj.sacc")
 cov_d = s_d.precision.getCovarianceMatrix()
 cov_m = s_m.precision.getCovarianceMatrix()
 
+
+dof = np.trace(np.dot((cov_d),np.linalg.inv(cov_m)))
+print(dof)
+
 # get the cls
 Cl_fid = s_d.mean.vector
 Cl_marg = s_m.mean.vector
@@ -75,6 +82,7 @@ ndx = ndx[0]
 # labels
 xlabel = r'$z$'
 ylabel = r'$\ell$'
+cov_label = r'$C_{\ell}^{i,j}$'
 cm = 'Reds'
 
 want_evals = 0
@@ -86,18 +94,21 @@ if want_evals:
     plt.savefig("eigs.png")
     plt.close()
 
-want_cov = 0
+want_cov = 1
 if want_cov:
     # Get the corr matrices
[…]

### 2020-07-26 chain_analysis/tests_triangle.py
added triangle plot for tests

@@ -1,0 +1,119 @@
+import numpy as np
+import matplotlib
+matplotlib.use("Agg")
+import matplotlib.pyplot as plt
+from getdist import plots,MCSamples
+import sys
+
+hod_dic = {'lmmin':11.88, 'lmminp':-0.5,
+              'm0':5.7, 'm0p':2.5,
+              'm1':13.08, 'm1p':0.9,}
+              
+new_hod_dic = {'lmmin': 12.1,'lmminp': -2.5,
+               'm0': 9.7, 'm0p': -2.5,
+               'm1': 13.35, 'm1p': -1.7}
+
+# parameters
+n_iter = 12000; w_rat = 6; n_par = 6; b_iter = 2000
+par_names = ["lmminp","lmmin","m0p","m0","m1p","m1"]
+lab_names = ["\\mu_{{\\rm min},p}","\\mu_{\\rm min}","\\mu_{0,p}","\\mu_{0}","\\mu_{1,p}","\\mu_{1}"]
+
+# what are we plotting
+HOD_pars = ["lmminp","lmmin","m0p","m0","m1p","m1"]
+filename = "triangle_marg_tests.pdf"#.png"
+dir_chains = "/users/boryanah/desclss_chains/"
+
+# names and choices
+name_marg = "marg_newcov_june19"#"marg_newcov_april29"
+name_test1 = "marg_newcov_1sigma_june21"#"marg_newcov_1sigma_may1"
+name_test2 = "marg_newcov_nosmooth_june22"#"marg_newcov_nosmooth_may2"
+name_test3 = "marg_newcov_largenoise_june23"#"marg_newcov_largenoise_may3"
+
+lab_marg = '6-parameter fit using ${\sf C}_M$, fiducial'#'Marginalized N(z), fiducial'#'marg Nz, new covmat'
+lab_test1 = '6-parameter fit using ${\sf C}_M$, $1\sigma$ off'#'Marginalized N(z), $1\sigma$ off'#'marg Nz 1 sigma, new covmat'
+lab_test2 = '6-parameter fit using ${\sf C}_M$, no smoothing'#'Marginalized N(z), no smoothing'#'marg Nz no smoothing, new covmat'
+lab_test3 = '6-parameter fit using ${\sf C}_M$, large diagonal noise'#'Marginalized N(z), large diagonal noise'#'marg Nz large noise, new covmat'
+
+# walkers ratio, number of params and burn in iterations
[…]

### 2020-07-26 chain_analysis/CV_triangle_plot.py
colors of triangle plot

@@ -109,6 +109,6 @@ print("\\end{table}")
 g = plots.getSubplotPlotter()
 g.settings.legend_fontsize = 20
 g.settings.scaling_factor = 0.1
-g.triangle_plot([fiducial_hsc,marg_hsc,CV_hsc],params=HOD_pars,legend_labels=[lab_fid,lab_marg,lab_CV],filled=True)
+g.triangle_plot([fiducial_hsc,marg_hsc,CV_hsc],params=HOD_pars,legend_labels=[lab_fid,lab_marg,lab_CV],colors=['gray','#006FED','#E03424'],filled=True)
 plt.savefig("../Paper/"+filename)
 plt.close()
